### 2016-03-24 additional_files/tex/cluster1_BP_Mm_EarlyUp.tex
cleaned up status but as on poster

% latex table generated in R 3.2.3 by xtable 1.8-2 package
% Thu Mar 24 06:00:18 2016
\begin{table}[ht]
\centering
\begin{tabular}{rllrrrrr}
  \hline
 & GO.ID & Term & Annotated & Significant & Expected & p.value & adj.p \\ 
  \hline
1 & GO:0045087 & innate immune response & 340 &  61 & 7.54 & 0.00 & 0.00 \\ 
  2 & GO:0006955 & immune response & 674 &  79 & 14.95 & 0.00 & 0.00 \\ 
  3 & GO:0006952 & defense response & 724 &  75 & 16.06 & 0.00 & 0.00 \\ 
  4 & GO:0002376 & immune system process & 1326 &  98 & 29.41 & 0.00 & 0.00 \\ 
  5 & GO:0098542 & defense response to other organism & 282 &  48 & 6.25 & 0.00 & 0.00 \\ 
  6 & GO:0043207 & response to external biotic stimulus & 438 &  55 & 9.71 & 0.00 & 0.00 \\ 
  7 & GO:0051707 & response to other organism & 438 &  55 & 9.71 & 0.00 & 0.00 \\ 
  8 & GO:0009607 & response to biotic stimulus & 460 &  55 & 10.20 & 0.00 & 0.00 \\ 
  9 & GO:0002252 & immune effector process & 440 &  48 & 9.76 & 0.00 & 0.00 \\ 
  10 & GO:0009615 & response to virus & 205 &  34 & 4.55 & 0.00 & 0.00 \\ 
  11 & GO:0051607 & defense response to virus & 181 &  32 & 4.01 & 0.00 & 0.00 \\ 
  12 & GO:0034097 & response to cytokine & 346 &  42 & 7.67 & 0.00 & 0.00 \\ 
  13 & GO:0034341 & response to interferon-gamma &  47 &  18 & 1.04 & 0.00 & 0.00 \\ 
  14 & GO:0006950 & response to stress & 2042 &  98 & 45.29 & 0.00 & 0.00 \\ 
  15 & GO:0035456 & response to interferon-beta &  33 &  14 & 0.73 & 0.00 & 0.00 \\ 
  16 & GO:0071345 & cellular response to cytokine stimulus & 276 &  32 & 6.12 & 0.00 & 0.00 \\ 
  17 & GO:0045088 & regulation of innate immune response & 146 &  24 & 3.24 & 0.00 & 0.00 \\ 
  18 & GO:0051704 & multi-organism process & 959 &  61 & 21.27 & 0.00 & 0.00 \\ 
  19 & GO:0009605 & response to external stimulus & 1135 &  67 & 25.17 & 0.00 & 0.00 \\ 
  20 & GO:0002682 & regulation of immune system process & 732 &  52 & 16.23 & 0.00 & 0.00 \\ 
  21 & GO:0048525 & negative regulation of viral process &  66 &  17 & 1.46 & 0.00 & 0.00 \\ 
  22 & GO:0019882 & antigen processing and presentation &  69 &  17 & 1.53 & 0.00 & 0.00 \\ 
  23 & GO:0035458 & cellular response to interferon-beta &  27 &  12 & 0.60 & 0.00 & 0.00 \\ 
  24 & GO:1903901 & negative regulation of viral life cycle &  63 &  16 & 1.40 & 0.00 & 0.00 \\ 
  25 & GO:0043901 & negative regulation of multi-organism pr... & 103 &  19 & 2.28 & 0.00 & 0.00 \\ 
  26 & GO:0050776 & regulation of immune response & 366 &  34 & 8.12 & 0.00 & 0.00 \\ 
  27 & GO:0050792 & regulation of viral process & 118 &  20 & 2.62 & 0.00 & 0.00 \\ 
  28 & GO:0043900 & regulation of multi-organism process & 271 &  29 & 6.01 & 0.00 & 0.00 \\ 
  29 & GO:0071346 & cellular response to interferon-gamma &  32 &  12 & 0.71 & 0.00 & 0.00 \\ 
  30 & GO:0019884 & antigen processing and presentation of e... &  25 &  11 & 0.55 & 0.00 & 0.00 \\ 
  31 & GO:0044764 & multi-organism cellular process & 201 &  25 & 4.46 & 0.00 & 0.00 \\ 
  32 & GO:0043903 & regulation of symbiosis, encompassing mu... & 137 &  21 & 3.04 & 0.00 & 0.00 \\ 
  33 & GO:0045071 & negative regulation of viral genome repl... &  33 &  12 & 0.73 & 0.00 & 0.00 \\ 
  34 & GO:0044403 & symbiosis, encompassing mutualism throug... & 225 &  26 & 4.99 & 0.00 & 0.00 \\ 
  35 & GO:0044419 & interspecies interaction between organis... & 225 &  26 & 4.99 & 0.00 & 0.00 \\ 
  36 & GO:0031347 & regulation of defense response & 352 &  32 & 7.81 & 0.00 & 0.00 \\ 
  37 & GO:0048002 & antigen processing and presentation of p... &  46 &  13 & 1.02 & 0.00 & 0.00 \\ 
  38 & GO:0009617 & response to bacterium & 240 &  26 & 5.32 & 0.00 & 0.00 \\ 
  39 & GO:0045089 & positive regulation of innate immune res... & 120 &  18 & 2.66 & 0.00 & 0.00 \\ 
  40 & GO:0002478 & antigen processing and presentation of e... &  20 &   9 & 0.44 & 0.00 & 0.00 \\ 
  41 & GO:0050778 & positive regulation of immune response & 285 &  27 & 6.32 & 0.00 & 0.00 \\ 
  42 & GO:1903900 & regulation of viral life cycle & 109 &  17 & 2.42 & 0.00 & 0.00 \\ 
  43 & GO:0034340 & response to type I interferon &  22 &   9 & 0.49 & 0.00 & 0.00 \\ 
  44 & GO:0042742 & defense response to bacterium & 102 &  16 & 2.26 & 0.00 & 0.00 \\ 
  45 & GO:0002684 & positive regulation of immune system pro... & 471 &  34 & 10.45 & 0.00 & 0.00 \\ 
  46 & GO:0045069 & regulation of viral genome replication &  52 &  12 & 1.15 & 0.00 & 0.00 \\ 
  47 & GO:0031349 & positive regulation of defense response & 170 &  20 & 3.77 & 0.00 & 0.00 \\ 
  48 & GO:0019058 & viral life cycle & 153 &  19 & 3.39 & 0.00 & 0.00 \\ 
  49 & GO:0002221 & pattern recognition receptor signaling p... &  77 &  14 & 1.71 & 0.00 & 0.00 \\ 
  50 & GO:0016032 & viral process & 191 &  21 & 4.24 & 0.00 & 0.00 \\ 
  51 & GO:0002758 & innate immune response-activating signal... &  80 &  14 & 1.77 & 0.00 & 0.00 \\ 
  52 & GO:0002683 & negative regulation of immune system pro... & 257 &  24 & 5.70 & 0.00 & 0.00 \\ 
  53 & GO:0042832 & defense response to protozoan &  20 &   8 & 0.44 & 0.00 & 0.00 \\ 
  54 & GO:0002218 & activation of innate immune response &  86 &  14 & 1.91 & 0.00 & 0.00 \\ 
  55 & GO:0002474 & antigen processing and presentation of p... &  28 &   9 & 0.62 & 0.00 & 0.00 \\ 
  56 & GO:0019079 & viral genome replication &  60 &  12 & 1.33 & 0.00 & 0.00 \\ 
  57 & GO:0050896 & response to stimulus & 4167 & 136 & 92.42 & 0.00 & 0.00 \\ 
  58 & GO:0002483 & antigen processing and presentation of e... &   9 &   6 & 0.20 & 0.00 & 0.00 \\ 
  59 & GO:0010033 & response to organic substance & 1314 &  61 & 29.14 & 0.00 & 0.00 \\ 
  60 & GO:0001562 & response to protozoan &  22 &   8 & 0.49 & 0.00 & 0.00 \\ 
  61 & GO:0098586 & cellular response to virus &  22 &   8 & 0.49 & 0.00 & 0.00 \\ 
  62 & GO:0019883 & antigen processing and presentation of e... &  10 &   6 & 0.22 & 0.00 & 0.00 \\ 
  63 & GO:0001816 & cytokine production & 381 &  28 & 8.45 & 0.00 & 0.00 \\ 
  64 & GO:0035455 & response to interferon-alpha &  16 &   7 & 0.35 & 0.00 & 0.00 \\ 
  65 & GO:0001817 & regulation of cytokine production & 334 &  26 & 7.41 & 0.00 & 0.00 \\ 
  66 & GO:0002237 & response to molecule of bacterial origin & 154 &  17 & 3.42 & 0.00 & 0.00 \\ 
  67 & GO:0060337 & type I interferon signaling pathway &  18 &   7 & 0.40 & 0.00 & 0.00 \\ 
  68 & GO:0071357 & cellular response to type I interferon &  18 &   7 & 0.40 & 0.00 & 0.00 \\ 
  69 & GO:0051250 & negative regulation of lymphocyte activa... &  89 &  13 & 1.97 & 0.00 & 0.00 \\ 
  70 & GO:0080134 & regulation of response to stress & 790 &  42 & 17.52 & 0.00 & 0.00 \\ 
  71 & GO:0019885 & antigen processing and presentation of e... &   7 &   5 & 0.16 & 0.00 & 0.00 \\ 
  72 & GO:0039528 & cytoplasmic pattern recognition receptor... &  13 &   6 & 0.29 & 0.00 & 0.00 \\ 
  73 & GO:0002695 & negative regulation of leukocyte activat... &  99 &  13 & 2.20 & 0.00 & 0.00 \\ 
  74 & GO:0002753 & cytoplasmic pattern recognition receptor... &  22 &   7 & 0.49 & 0.00 & 0.00 \\ 
  75 & GO:0001819 & positive regulation of cytokine producti... & 222 &  19 & 4.92 & 0.00 & 0.00 \\ 
  76 & GO:0032606 & type I interferon production &  45 &   9 & 1.00 & 0.00 & 0.00 \\ 
  77 & GO:0002764 & immune response-regulating signaling pat... & 181 &  17 & 4.01 & 0.00 & 0.00 \\ 
  78 & GO:0048584 & positive regulation of response to stimu... & 1076 &  49 & 23.86 & 0.00 & 0.00 \\ 
  79 & GO:0050866 & negative regulation of cell activation & 108 &  13 & 2.40 & 0.00 & 0.00 \\ 
  80 & GO:0002821 & positive regulation of adaptive immune r... &  61 &  10 & 1.35 & 0.00 & 0.00 \\ 
  81 & GO:0002697 & regulation of immune effector process & 258 &  20 & 5.72 & 0.00 & 0.00 \\ 
  82 & GO:0045341 & MHC class I biosynthetic process &   5 &   4 & 0.11 & 0.00 & 0.00 \\ 
  83 & GO:0045343 & regulation of MHC class I biosynthetic p... &   5 &   4 & 0.11 & 0.00 & 0.00 \\ 
  84 & GO:0002757 & immune response-activating signal transd... & 171 &  16 & 3.79 & 0.00 & 0.00 \\ 
  85 & GO:0002699 & positive regulation of immune effector p... & 115 &  13 & 2.55 & 0.00 & 0.00 \\ 
  86 & GO:0019221 & cytokine-mediated signaling pathway & 175 &  16 & 3.88 & 0.00 & 0.00 \\ 
  87 & GO:0050868 & negative regulation of T cell activation &  66 &  10 & 1.46 & 0.00 & 0.00 \\ 
  88 & GO:0071310 & cellular response to organic substance & 981 &  45 & 21.76 & 0.00 & 0.00 \\ 
  89 & GO:0044406 & adhesion of symbiont to host &  11 &   5 & 0.24 & 0.00 & 0.00 \\ 
  90 & GO:0051249 & regulation of lymphocyte activation & 246 &  19 & 5.46 & 0.00 & 0.00 \\ 
  91 & GO:0032496 & response to lipopolysaccharide & 141 &  14 & 3.13 & 0.00 & 0.00 \\ 
  92 & GO:1903038 & negative regulation of leukocyte cell-ce... &  70 &  10 & 1.55 & 0.00 & 0.00 \\ 
  93 & GO:0002824 & positive regulation of adaptive immune r... &  56 &   9 & 1.24 & 0.00 & 0.00 \\ 
  94 & GO:0039529 & RIG-I signaling pathway &  12 &   5 & 0.27 & 0.00 & 0.00 \\ 
  95 & GO:0045824 & negative regulation of innate immune res... &  31 &   7 & 0.69 & 0.00 & 0.00 \\ 
  96 & GO:0050830 & defense response to Gram-positive bacter... &  45 &   8 & 1.00 & 0.00 & 0.00 \\ 
  97 & GO:0002819 & regulation of adaptive immune response &  93 &  11 & 2.06 & 0.00 & 0.00 \\ 
  98 & GO:0002253 & activation of immune response & 194 &  16 & 4.30 & 0.00 & 0.00 \\ 
  99 & GO:0002250 & adaptive immune response & 217 &  17 & 4.81 & 0.00 & 0.00 \\ 
  100 & GO:1900246 & positive regulation of RIG-I signaling p... &   7 &   4 & 0.16 & 0.00 & 0.00 \\ 
   \hline
\end{tabular}
\end{table}