### 2016-03-24 additional_files/tex/cluster3_BP_Mm_7dpiDown.tex

% latex table generated in R 3.2.3 by xtable 1.8-2 package
% Thu Mar 24 05:59:04 2016
\begin{table}[ht]
\centering
\begin{tabular}{rllrrrrr}
  \hline
 & GO.ID & Term & Annotated & Significant & Expected & p.value & adj.p \\ 
  \hline
1 & GO:0055114 & oxidation-reduction process & 597 &  30 & 8.37 & 0.00 & 0.00 \\ 
  2 & GO:0032787 & monocarboxylic acid metabolic process & 350 &  21 & 4.90 & 0.00 & 0.00 \\ 
  3 & GO:0019752 & carboxylic acid metabolic process & 578 &  27 & 8.10 & 0.00 & 0.00 \\ 
  4 & GO:0006631 & fatty acid metabolic process & 234 &  17 & 3.28 & 0.00 & 0.00 \\ 
  5 & GO:0043436 & oxoacid metabolic process & 617 &  27 & 8.65 & 0.00 & 0.00 \\ 
  6 & GO:0006082 & organic acid metabolic process & 625 &  27 & 8.76 & 0.00 & 0.00 \\ 
  7 & GO:0044281 & small molecule metabolic process & 1187 &  39 & 16.63 & 0.00 & 0.00 \\ 
  8 & GO:0006629 & lipid metabolic process & 767 &  29 & 10.75 & 0.00 & 0.00 \\ 
  9 & GO:0044710 & single-organism metabolic process & 3037 &  69 & 42.56 & 0.00 & 0.00 \\ 
  10 & GO:0001676 & long-chain fatty acid metabolic process &  47 &   7 & 0.66 & 0.00 & 0.00 \\ 
  11 & GO:0006690 & icosanoid metabolic process &  56 &   7 & 0.78 & 0.00 & 0.00 \\ 
  12 & GO:1901568 & fatty acid derivative metabolic process &  56 &   7 & 0.78 & 0.00 & 0.00 \\ 
  13 & GO:0006805 & xenobiotic metabolic process &  38 &   6 & 0.53 & 0.00 & 0.00 \\ 
  14 & GO:0019369 & arachidonic acid metabolic process &  25 &   5 & 0.35 & 0.00 & 0.00 \\ 
  15 & GO:0071466 & cellular response to xenobiotic stimulus &  43 &   6 & 0.60 & 0.00 & 0.00 \\ 
  16 & GO:0019395 & fatty acid oxidation &  65 &   7 & 0.91 & 0.00 & 0.00 \\ 
  17 & GO:0034440 & lipid oxidation &  66 &   7 & 0.92 & 0.00 & 0.00 \\ 
  18 & GO:0072329 & monocarboxylic acid catabolic process &  66 &   7 & 0.92 & 0.00 & 0.00 \\ 
  19 & GO:0042738 & exogenous drug catabolic process &  15 &   4 & 0.21 & 0.00 & 0.00 \\ 
  20 & GO:0009410 & response to xenobiotic stimulus &  47 &   6 & 0.66 & 0.00 & 0.00 \\ 
  21 & GO:0019373 & epoxygenase P450 pathway &   6 &   3 & 0.08 & 0.00 & 0.00 \\ 
  22 & GO:0044255 & cellular lipid metabolic process & 579 &  21 & 8.11 & 0.00 & 0.00 \\ 
  23 & GO:0042737 & drug catabolic process &  16 &   4 & 0.22 & 0.00 & 0.00 \\ 
  24 & GO:0042493 & response to drug & 157 &  10 & 2.20 & 0.00 & 0.00 \\ 
  25 & GO:0032099 & negative regulation of appetite &   8 &   3 & 0.11 & 0.00 & 0.00 \\ 
  26 & GO:0009062 & fatty acid catabolic process &  59 &   6 & 0.83 & 0.00 & 0.00 \\ 
  27 & GO:0033559 & unsaturated fatty acid metabolic process &  59 &   6 & 0.83 & 0.00 & 0.00 \\ 
  28 & GO:0097089 & methyl-branched fatty acid metabolic pro... &   2 &   2 & 0.03 & 0.00 & 0.00 \\ 
  29 & GO:0032096 & negative regulation of response to food &   9 &   3 & 0.13 & 0.00 & 0.00 \\ 
  30 & GO:0060979 & vasculogenesis involved in coronary vasc... &   9 &   3 & 0.13 & 0.00 & 0.00 \\ 
  31 & GO:0016054 & organic acid catabolic process & 122 &   8 & 1.71 & 0.00 & 0.00 \\ 
  32 & GO:0046395 & carboxylic acid catabolic process & 122 &   8 & 1.71 & 0.00 & 0.00 \\ 
  33 & GO:0017144 & drug metabolic process &  24 &   4 & 0.34 & 0.00 & 0.00 \\ 
  34 & GO:0006635 & fatty acid beta-oxidation &  45 &   5 & 0.63 & 0.00 & 0.00 \\ 
  35 & GO:0030258 & lipid modification & 127 &   8 & 1.78 & 0.00 & 0.00 \\ 
  36 & GO:0032095 & regulation of response to food &  11 &   3 & 0.15 & 0.00 & 0.00 \\ 
  37 & GO:0032098 & regulation of appetite &  11 &   3 & 0.15 & 0.00 & 0.00 \\ 
  38 & GO:2001023 & regulation of response to drug &   3 &   2 & 0.04 & 0.00 & 0.00 \\ 
  39 & GO:1901615 & organic hydroxy compound metabolic proce... & 287 &  12 & 4.02 & 0.00 & 0.00 \\ 
  40 & GO:0015701 & bicarbonate transport &  15 &   3 & 0.21 & 0.00 & 0.00 \\ 
  41 & GO:0001561 & fatty acid alpha-oxidation &   4 &   2 & 0.06 & 0.00 & 0.00 \\ 
  42 & GO:0030916 & otic vesicle formation &   4 &   2 & 0.06 & 0.00 & 0.00 \\ 
  43 & GO:0071600 & otic vesicle morphogenesis &   4 &   2 & 0.06 & 0.00 & 0.00 \\ 
  44 & GO:0032094 & response to food &  16 &   3 & 0.22 & 0.00 & 0.00 \\ 
  45 & GO:0048566 & embryonic digestive tract development &  18 &   3 & 0.25 & 0.00 & 0.00 \\ 
  46 & GO:0060977 & coronary vasculature morphogenesis &  18 &   3 & 0.25 & 0.00 & 0.00 \\ 
  47 & GO:0042760 & very long-chain fatty acid catabolic pro... &   5 &   2 & 0.07 & 0.00 & 0.00 \\ 
  48 & GO:0044539 & long-chain fatty acid import &   5 &   2 & 0.07 & 0.00 & 0.00 \\ 
  49 & GO:0046135 & pyrimidine nucleoside catabolic process &   5 &   2 & 0.07 & 0.00 & 0.00 \\ 
  50 & GO:0072366 & regulation of cellular ketone metabolic ... &   5 &   2 & 0.07 & 0.00 & 0.00 \\ 
  51 & GO:0097164 & ammonium ion metabolic process &  94 &   6 & 1.32 & 0.00 & 0.00 \\ 
  52 & GO:0015711 & organic anion transport & 202 &   9 & 2.83 & 0.00 & 0.00 \\ 
  53 & GO:0001822 & kidney development & 132 &   7 & 1.85 & 0.00 & 0.00 \\ 
  54 & GO:0050892 & intestinal absorption &  20 &   3 & 0.28 & 0.00 & 0.00 \\ 
  55 & GO:0016042 & lipid catabolic process & 171 &   8 & 2.40 & 0.00 & 0.00 \\ 
  56 & GO:0019532 & oxalate transport &   6 &   2 & 0.08 & 0.00 & 0.00 \\ 
  57 & GO:0060484 & lung-associated mesenchyme development &   6 &   2 & 0.08 & 0.00 & 0.00 \\ 
  58 & GO:0071599 & otic vesicle development &   6 &   2 & 0.08 & 0.00 & 0.00 \\ 
  59 & GO:0043648 & dicarboxylic acid metabolic process &  70 &   5 & 0.98 & 0.00 & 0.01 \\ 
  60 & GO:0015718 & monocarboxylic acid transport &  71 &   5 & 0.99 & 0.00 & 0.01 \\ 
  61 & GO:0044282 & small molecule catabolic process & 178 &   8 & 2.49 & 0.00 & 0.01 \\ 
  62 & GO:0048754 & branching morphogenesis of an epithelial... & 106 &   6 & 1.49 & 0.00 & 0.01 \\ 
  63 & GO:0010817 & regulation of hormone levels & 262 &  10 & 3.67 & 0.00 & 0.01 \\ 
  64 & GO:0048002 & antigen processing and presentation of p... &  46 &   4 & 0.64 & 0.00 & 0.01 \\ 
  65 & GO:0032105 & negative regulation of response to extra... &  23 &   3 & 0.32 & 0.00 & 0.01 \\ 
  66 & GO:0032108 & negative regulation of response to nutri... &  23 &   3 & 0.32 & 0.00 & 0.01 \\ 
  67 & GO:0006560 & proline metabolic process &   7 &   2 & 0.10 & 0.00 & 0.01 \\ 
  68 & GO:0090331 & negative regulation of platelet aggregat... &   7 &   2 & 0.10 & 0.00 & 0.01 \\ 
  69 & GO:1902358 & sulfate transmembrane transport &   7 &   2 & 0.10 & 0.00 & 0.01 \\ 
  70 & GO:0072001 & renal system development & 143 &   7 & 2.00 & 0.00 & 0.01 \\ 
  71 & GO:0018958 & phenol-containing compound metabolic pro... &  48 &   4 & 0.67 & 0.00 & 0.01 \\ 
  72 & GO:0015849 & organic acid transport & 147 &   7 & 2.06 & 0.00 & 0.01 \\ 
  73 & GO:0046942 & carboxylic acid transport & 147 &   7 & 2.06 & 0.00 & 0.01 \\ 
  74 & GO:0070374 & positive regulation of ERK1 and ERK2 cas... &  78 &   5 & 1.09 & 0.00 & 0.01 \\ 
  75 & GO:0006584 & catecholamine metabolic process &  25 &   3 & 0.35 & 0.00 & 0.01 \\ 
  76 & GO:0007431 & salivary gland development &  25 &   3 & 0.35 & 0.00 & 0.01 \\ 
  77 & GO:0009712 & catechol-containing compound metabolic p... &  25 &   3 & 0.35 & 0.00 & 0.01 \\ 
  78 & GO:0044242 & cellular lipid catabolic process & 114 &   6 & 1.60 & 0.01 & 0.01 \\ 
  79 & GO:0009948 & anterior/posterior axis specification &  26 &   3 & 0.36 & 0.01 & 0.01 \\ 
  80 & GO:0060603 & mammary gland duct morphogenesis &  26 &   3 & 0.36 & 0.01 & 0.01 \\ 
  81 & GO:0060976 & coronary vasculature development &  26 &   3 & 0.36 & 0.01 & 0.01 \\ 
  82 & GO:0030004 & cellular monovalent inorganic cation hom... &  51 &   4 & 0.71 & 0.01 & 0.01 \\ 
  83 & GO:0006790 & sulfur compound metabolic process & 194 &   8 & 2.72 & 0.01 & 0.01 \\ 
  84 & GO:0008272 & sulfate transport &   9 &   2 & 0.13 & 0.01 & 0.01 \\ 
  85 & GO:0010544 & negative regulation of platelet activati... &   9 &   2 & 0.13 & 0.01 & 0.01 \\ 
  86 & GO:0072364 & regulation of cellular ketone metabolic ... &   9 &   2 & 0.13 & 0.01 & 0.01 \\ 
  87 & GO:0002474 & antigen processing and presentation of p... &  28 &   3 & 0.39 & 0.01 & 0.01 \\ 
  88 & GO:0051196 & regulation of coenzyme metabolic process &  28 &   3 & 0.39 & 0.01 & 0.01 \\ 
  89 & GO:0006820 & anion transport & 288 &  10 & 4.04 & 0.01 & 0.01 \\ 
  90 & GO:0015909 & long-chain fatty acid transport &  29 &   3 & 0.41 & 0.01 & 0.01 \\ 
  91 & GO:0045844 & positive regulation of striated muscle t... &  29 &   3 & 0.41 & 0.01 & 0.01 \\ 
  92 & GO:0048636 & positive regulation of muscle organ deve... &  29 &   3 & 0.41 & 0.01 & 0.01 \\ 
  93 & GO:0051193 & regulation of cofactor metabolic process &  29 &   3 & 0.41 & 0.01 & 0.01 \\ 
  94 & GO:0009056 & catabolic process & 1369 &  30 & 19.18 & 0.01 & 0.01 \\ 
  95 & GO:0051923 & sulfation &  10 &   2 & 0.14 & 0.01 & 0.01 \\ 
  96 & GO:0060081 & membrane hyperpolarization &  10 &   2 & 0.14 & 0.01 & 0.01 \\ 
  97 & GO:0072529 & pyrimidine-containing compound catabolic... &  10 &   2 & 0.14 & 0.01 & 0.01 \\ 
  98 & GO:0009880 & embryonic pattern specification &  30 &   3 & 0.42 & 0.01 & 0.01 \\ 
  99 & GO:1901863 & positive regulation of muscle tissue dev... &  30 &   3 & 0.42 & 0.01 & 0.01 \\ 
  100 & GO:0042060 & wound healing & 210 &   8 & 2.94 & 0.01 & 0.01 \\ 
   \hline
\end{tabular}
\end{table}