### 2016-03-24 additional_files/tex/cluster4_BP_Mm_allInfDown.tex

% latex table generated in R 3.2.3 by xtable 1.8-2 package
% Thu Mar 24 06:01:34 2016
\begin{table}[ht]
\centering
\begin{tabular}{rllrrrrr}
  \hline
 & GO.ID & Term & Annotated & Significant & Expected & p.value & adj.p \\ 
  \hline
1 & GO:0044281 & small molecule metabolic process & 1187 &  94 & 44.61 & 0.00 & 0.00 \\ 
  2 & GO:0032787 & monocarboxylic acid metabolic process & 350 &  40 & 13.15 & 0.00 & 0.00 \\ 
  3 & GO:0006629 & lipid metabolic process & 767 &  63 & 28.83 & 0.00 & 0.00 \\ 
  4 & GO:0006082 & organic acid metabolic process & 625 &  55 & 23.49 & 0.00 & 0.00 \\ 
  5 & GO:0043436 & oxoacid metabolic process & 617 &  53 & 23.19 & 0.00 & 0.00 \\ 
  6 & GO:0019752 & carboxylic acid metabolic process & 578 &  50 & 21.72 & 0.00 & 0.00 \\ 
  7 & GO:0042178 & xenobiotic catabolic process &   9 &   6 & 0.34 & 0.00 & 0.00 \\ 
  8 & GO:0006805 & xenobiotic metabolic process &  38 &  10 & 1.43 & 0.00 & 0.00 \\ 
  9 & GO:0044710 & single-organism metabolic process & 3037 & 158 & 114.14 & 0.00 & 0.00 \\ 
  10 & GO:0044255 & cellular lipid metabolic process & 579 &  46 & 21.76 & 0.00 & 0.00 \\ 
  11 & GO:0006631 & fatty acid metabolic process & 234 &  25 & 8.79 & 0.00 & 0.00 \\ 
  12 & GO:0071466 & cellular response to xenobiotic stimulus &  43 &  10 & 1.62 & 0.00 & 0.00 \\ 
  13 & GO:0006790 & sulfur compound metabolic process & 194 &  22 & 7.29 & 0.00 & 0.00 \\ 
  14 & GO:0034754 & cellular hormone metabolic process &  45 &  10 & 1.69 & 0.00 & 0.00 \\ 
  15 & GO:0042445 & hormone metabolic process &  89 &  14 & 3.34 & 0.00 & 0.00 \\ 
  16 & GO:0009410 & response to xenobiotic stimulus &  47 &  10 & 1.77 & 0.00 & 0.00 \\ 
  17 & GO:0009719 & response to endogenous stimulus & 701 &  50 & 26.35 & 0.00 & 0.00 \\ 
  18 & GO:0042537 & benzene-containing compound metabolic pr... &  16 &   6 & 0.60 & 0.00 & 0.00 \\ 
  19 & GO:0016042 & lipid catabolic process & 171 &  19 & 6.43 & 0.00 & 0.00 \\ 
  20 & GO:0009813 & flavonoid biosynthetic process &   6 &   4 & 0.23 & 0.00 & 0.00 \\ 
  21 & GO:0052696 & flavonoid glucuronidation &   6 &   4 & 0.23 & 0.00 & 0.00 \\ 
  22 & GO:0055114 & oxidation-reduction process & 597 &  43 & 22.44 & 0.00 & 0.00 \\ 
  23 & GO:0071495 & cellular response to endogenous stimulus & 540 &  40 & 20.30 & 0.00 & 0.00 \\ 
  24 & GO:0072329 & monocarboxylic acid catabolic process &  66 &  11 & 2.48 & 0.00 & 0.00 \\ 
  25 & GO:0019626 & short-chain fatty acid catabolic process &   3 &   3 & 0.11 & 0.00 & 0.00 \\ 
  26 & GO:0051410 & detoxification of nitrogen compound &   3 &   3 & 0.11 & 0.00 & 0.00 \\ 
  27 & GO:0070458 & cellular detoxification of nitrogen comp... &   3 &   3 & 0.11 & 0.00 & 0.00 \\ 
  28 & GO:0006063 & uronic acid metabolic process &   7 &   4 & 0.26 & 0.00 & 0.00 \\ 
  29 & GO:0009812 & flavonoid metabolic process &   7 &   4 & 0.26 & 0.00 & 0.00 \\ 
  30 & GO:0019585 & glucuronate metabolic process &   7 &   4 & 0.26 & 0.00 & 0.00 \\ 
  31 & GO:0052695 & cellular glucuronidation &   7 &   4 & 0.26 & 0.00 & 0.00 \\ 
  32 & GO:0044242 & cellular lipid catabolic process & 114 &  14 & 4.28 & 0.00 & 0.00 \\ 
  33 & GO:0009725 & response to hormone & 365 &  29 & 13.72 & 0.00 & 0.00 \\ 
  34 & GO:0008202 & steroid metabolic process & 162 &  17 & 6.09 & 0.00 & 0.00 \\ 
  35 & GO:0044282 & small molecule catabolic process & 178 &  18 & 6.69 & 0.00 & 0.00 \\ 
  36 & GO:0006749 & glutathione metabolic process &  42 &   8 & 1.58 & 0.00 & 0.00 \\ 
  37 & GO:0009636 & response to toxic substance &  78 &  11 & 2.93 & 0.00 & 0.00 \\ 
  38 & GO:0085029 & extracellular matrix assembly &  15 &   5 & 0.56 & 0.00 & 0.00 \\ 
  39 & GO:0018916 & nitrobenzene metabolic process &   4 &   3 & 0.15 & 0.00 & 0.00 \\ 
  40 & GO:0051387 & negative regulation of neurotrophin TRK ... &   4 &   3 & 0.15 & 0.00 & 0.00 \\ 
  41 & GO:0046459 & short-chain fatty acid metabolic process &   9 &   4 & 0.34 & 0.00 & 0.00 \\ 
  42 & GO:1901201 & regulation of extracellular matrix assem... &   9 &   4 & 0.34 & 0.00 & 0.00 \\ 
  43 & GO:0006164 & purine nucleotide biosynthetic process & 124 &  14 & 4.66 & 0.00 & 0.00 \\ 
  44 & GO:0009062 & fatty acid catabolic process &  59 &   9 & 2.22 & 0.00 & 0.00 \\ 
  45 & GO:1903055 & positive regulation of extracellular mat... &  10 &   4 & 0.38 & 0.00 & 0.00 \\ 
  46 & GO:0010817 & regulation of hormone levels & 262 &  22 & 9.85 & 0.00 & 0.00 \\ 
  47 & GO:0072522 & purine-containing compound biosynthetic ... & 130 &  14 & 4.89 & 0.00 & 0.00 \\ 
  48 & GO:0044699 & single-organism process & 7377 & 308 & 277.25 & 0.00 & 0.00 \\ 
  49 & GO:0035637 & multicellular organismal signaling &  38 &   7 & 1.43 & 0.00 & 0.00 \\ 
  50 & GO:0009165 & nucleotide biosynthetic process & 165 &  16 & 6.20 & 0.00 & 0.00 \\ 
  51 & GO:0042572 & retinol metabolic process &  11 &   4 & 0.41 & 0.00 & 0.00 \\ 
  52 & GO:0072521 & purine-containing compound metabolic pro... & 289 &  23 & 10.86 & 0.00 & 0.00 \\ 
  53 & GO:1900542 & regulation of purine nucleotide metaboli... &  92 &  11 & 3.46 & 0.00 & 0.00 \\ 
  54 & GO:0016054 & organic acid catabolic process & 122 &  13 & 4.59 & 0.00 & 0.00 \\ 
  55 & GO:0046395 & carboxylic acid catabolic process & 122 &  13 & 4.59 & 0.00 & 0.00 \\ 
  56 & GO:1901293 & nucleoside phosphate biosynthetic proces... & 170 &  16 & 6.39 & 0.00 & 0.00 \\ 
  57 & GO:0019395 & fatty acid oxidation &  65 &   9 & 2.44 & 0.00 & 0.00 \\ 
  58 & GO:0006081 & cellular aldehyde metabolic process &  41 &   7 & 1.54 & 0.00 & 0.00 \\ 
  59 & GO:0034440 & lipid oxidation &  66 &   9 & 2.48 & 0.00 & 0.00 \\ 
  60 & GO:0006000 & fructose metabolic process &   6 &   3 & 0.23 & 0.00 & 0.00 \\ 
  61 & GO:0006140 & regulation of nucleotide metabolic proce... &  97 &  11 & 3.65 & 0.00 & 0.00 \\ 
  62 & GO:0060688 & regulation of morphogenesis of a branchi... &  32 &   6 & 1.20 & 0.00 & 0.00 \\ 
  63 & GO:0051186 & cofactor metabolic process & 249 &  20 & 9.36 & 0.00 & 0.00 \\ 
  64 & GO:0007267 & cell-cell signaling & 440 &  30 & 16.54 & 0.00 & 0.00 \\ 
  65 & GO:0009117 & nucleotide metabolic process & 362 &  26 & 13.61 & 0.00 & 0.00 \\ 
  66 & GO:0034220 & ion transmembrane transport & 383 &  27 & 14.39 & 0.00 & 0.00 \\ 
  67 & GO:0006635 & fatty acid beta-oxidation &  45 &   7 & 1.69 & 0.00 & 0.00 \\ 
  68 & GO:0006811 & ion transport & 756 &  45 & 28.41 & 0.00 & 0.00 \\ 
  69 & GO:0055085 & transmembrane transport & 546 &  35 & 20.52 & 0.00 & 0.00 \\ 
  70 & GO:0006001 & fructose catabolic process &   2 &   2 & 0.08 & 0.00 & 0.00 \\ 
  71 & GO:0018879 & biphenyl metabolic process &   2 &   2 & 0.08 & 0.00 & 0.00 \\ 
  72 & GO:0034587 & piRNA metabolic process &   2 &   2 & 0.08 & 0.00 & 0.00 \\ 
  73 & GO:0061624 & fructose catabolic process to hydroxyace... &   2 &   2 & 0.08 & 0.00 & 0.00 \\ 
  74 & GO:0061625 & glycolytic process through fructose-1-ph... &   2 &   2 & 0.08 & 0.00 & 0.00 \\ 
  75 & GO:1901135 & carbohydrate derivative metabolic proces... & 631 &  39 & 23.72 & 0.00 & 0.00 \\ 
  76 & GO:0009152 & purine ribonucleotide biosynthetic proce... & 117 &  12 & 4.40 & 0.00 & 0.00 \\ 
  77 & GO:0006753 & nucleoside phosphate metabolic process & 370 &  26 & 13.91 & 0.00 & 0.00 \\ 
  78 & GO:0003071 & renal system process involved in regulat... &   7 &   3 & 0.26 & 0.00 & 0.00 \\ 
  79 & GO:0019755 & one-carbon compound transport &   7 &   3 & 0.26 & 0.00 & 0.00 \\ 
  80 & GO:0051386 & regulation of neurotrophin TRK receptor ... &   7 &   3 & 0.26 & 0.00 & 0.00 \\ 
  81 & GO:1901203 & positive regulation of extracellular mat... &   7 &   3 & 0.26 & 0.00 & 0.00 \\ 
  82 & GO:0006163 & purine nucleotide metabolic process & 257 &  20 & 9.66 & 0.00 & 0.00 \\ 
  83 & GO:0044711 & single-organism biosynthetic process & 851 &  49 & 31.98 & 0.00 & 0.00 \\ 
  84 & GO:0006040 & amino sugar metabolic process &  24 &   5 & 0.90 & 0.00 & 0.00 \\ 
  85 & GO:0017144 & drug metabolic process &  24 &   5 & 0.90 & 0.00 & 0.00 \\ 
  86 & GO:0009118 & regulation of nucleoside metabolic proce... &  35 &   6 & 1.32 & 0.00 & 0.00 \\ 
  87 & GO:1903578 & regulation of ATP metabolic process &  35 &   6 & 1.32 & 0.00 & 0.00 \\ 
  88 & GO:0016266 & O-glycan processing &  15 &   4 & 0.56 & 0.00 & 0.00 \\ 
  89 & GO:0001523 & retinoid metabolic process &  25 &   5 & 0.94 & 0.00 & 0.00 \\ 
  90 & GO:0061098 & positive regulation of protein tyrosine ... &  25 &   5 & 0.94 & 0.00 & 0.00 \\ 
  91 & GO:0009260 & ribonucleotide biosynthetic process & 122 &  12 & 4.59 & 0.00 & 0.00 \\ 
  92 & GO:1901564 & organonitrogen compound metabolic proces... & 1198 &  64 & 45.03 & 0.00 & 0.00 \\ 
  93 & GO:0008285 & negative regulation of cell proliferatio... & 400 &  27 & 15.03 & 0.00 & 0.00 \\ 
  94 & GO:0035690 & cellular response to drug &  37 &   6 & 1.39 & 0.00 & 0.00 \\ 
  95 & GO:0009150 & purine ribonucleotide metabolic process & 246 &  19 & 9.25 & 0.00 & 0.00 \\ 
  96 & GO:0009755 & hormone-mediated signaling pathway &  63 &   8 & 2.37 & 0.00 & 0.00 \\ 
  97 & GO:0043467 & regulation of generation of precursor me... &  63 &   8 & 2.37 & 0.00 & 0.00 \\ 
  98 & GO:0046390 & ribose phosphate biosynthetic process & 124 &  12 & 4.66 & 0.00 & 0.00 \\ 
  99 & GO:1900371 & regulation of purine nucleotide biosynth... &  50 &   7 & 1.88 & 0.00 & 0.00 \\ 
  100 & GO:1990748 & cellular detoxification &  16 &   4 & 0.60 & 0.00 & 0.00 \\ 
   \hline
\end{tabular}
\end{table}