### 2016-04-08 additional_files/tex/cluster1_BP_Mm_EarlyUp.tex
Large changes in annotation script and lots of small adjustments in others (paths changed, typos, ggpair command adjustment... etc)

% latex table generated in R 3.2.2 by xtable 1.8-2 package
% Fri Apr  8 11:20:33 2016
\begin{table}[ht]
\centering
\begin{tabular}{rllrrrrr}
  \hline
 & GO.ID & Term & Annotated & Significant & Expected & p.value & adj.p \\ 
  \hline
1 & GO:0045087 & innate immune response & 362 &  51 & 7.18 & 0.00 & 0.00 \\ 
  2 & GO:0006955 & immune response & 713 &  65 & 14.14 & 0.00 & 0.00 \\ 
  3 & GO:0002376 & immune system process & 1406 &  85 & 27.88 & 0.00 & 0.00 \\ 
  4 & GO:0006952 & defense response & 755 &  61 & 14.97 & 0.00 & 0.00 \\ 
  5 & GO:0098542 & defense response to other organism & 295 &  38 & 5.85 & 0.00 & 0.00 \\ 
  6 & GO:0043207 & response to external biotic stimulus & 464 &  45 & 9.20 & 0.00 & 0.00 \\ 
  7 & GO:0051707 & response to other organism & 464 &  45 & 9.20 & 0.00 & 0.00 \\ 
  8 & GO:0009607 & response to biotic stimulus & 488 &  45 & 9.68 & 0.00 & 0.00 \\ 
  9 & GO:0034097 & response to cytokine & 360 &  38 & 7.14 & 0.00 & 0.00 \\ 
  10 & GO:0034341 & response to interferon-gamma &  47 &  16 & 0.93 & 0.00 & 0.00 \\ 
  11 & GO:0002252 & immune effector process & 461 &  39 & 9.14 & 0.00 & 0.00 \\ 
  12 & GO:0009615 & response to virus & 220 &  27 & 4.36 & 0.00 & 0.00 \\ 
  13 & GO:0051607 & defense response to virus & 192 &  25 & 3.81 & 0.00 & 0.00 \\ 
  14 & GO:0071345 & cellular response to cytokine stimulus & 292 &  29 & 5.79 & 0.00 & 0.00 \\ 
  15 & GO:0035456 & response to interferon-beta &  35 &  12 & 0.69 & 0.00 & 0.00 \\ 
  16 & GO:0002682 & regulation of immune system process & 774 &  47 & 15.35 & 0.00 & 0.00 \\ 
  17 & GO:0045088 & regulation of innate immune response & 152 &  20 & 3.01 & 0.00 & 0.00 \\ 
  18 & GO:0071346 & cellular response to interferon-gamma &  34 &  11 & 0.67 & 0.00 & 0.00 \\ 
  19 & GO:0035458 & cellular response to interferon-beta &  29 &  10 & 0.58 & 0.00 & 0.00 \\ 
  20 & GO:0051704 & multi-organism process & 1006 &  51 & 19.95 & 0.00 & 0.00 \\ 
  21 & GO:0050776 & regulation of immune response & 385 &  29 & 7.63 & 0.00 & 0.00 \\ 
  22 & GO:0031347 & regulation of defense response & 367 &  28 & 7.28 & 0.00 & 0.00 \\ 
  23 & GO:0034340 & response to type I interferon &  26 &   9 & 0.52 & 0.00 & 0.00 \\ 
  24 & GO:0006950 & response to stress & 2191 &  81 & 43.44 & 0.00 & 0.00 \\ 
  25 & GO:0042832 & defense response to protozoan &  21 &   8 & 0.42 & 0.00 & 0.00 \\ 
  26 & GO:0009617 & response to bacterium & 253 &  22 & 5.02 & 0.00 & 0.00 \\ 
  27 & GO:0001562 & response to protozoan &  23 &   8 & 0.46 & 0.00 & 0.00 \\ 
  28 & GO:0002683 & negative regulation of immune system pro... & 261 &  22 & 5.18 & 0.00 & 0.00 \\ 
  29 & GO:0019882 & antigen processing and presentation &  72 &  12 & 1.43 & 0.00 & 0.00 \\ 
  30 & GO:0019884 & antigen processing and presentation of e... &  25 &   8 & 0.50 & 0.00 & 0.00 \\ 
  31 & GO:0010033 & response to organic substance & 1390 &  58 & 27.56 & 0.00 & 0.00 \\ 
  32 & GO:0009605 & response to external stimulus & 1176 &  52 & 23.32 & 0.00 & 0.00 \\ 
  33 & GO:0043900 & regulation of multi-organism process & 281 &  22 & 5.57 & 0.00 & 0.00 \\ 
  34 & GO:0044403 & symbiosis, encompassing mutualism throug... & 234 &  20 & 4.64 & 0.00 & 0.00 \\ 
  35 & GO:0044419 & interspecies interaction between organis... & 234 &  20 & 4.64 & 0.00 & 0.00 \\ 
  36 & GO:0060337 & type I interferon signaling pathway &  22 &   7 & 0.44 & 0.00 & 0.00 \\ 
  37 & GO:0071357 & cellular response to type I interferon &  22 &   7 & 0.44 & 0.00 & 0.00 \\ 
  38 & GO:0042742 & defense response to bacterium & 105 &  13 & 2.08 & 0.00 & 0.00 \\ 
  39 & GO:0044764 & multi-organism cellular process & 212 &  18 & 4.20 & 0.00 & 0.00 \\ 
  40 & GO:0035455 & response to interferon-alpha &  15 &   6 & 0.30 & 0.00 & 0.00 \\ 
  41 & GO:0001816 & cytokine production & 397 &  25 & 7.87 & 0.00 & 0.00 \\ 
  42 & GO:0043903 & regulation of symbiosis, encompassing mu... & 139 &  14 & 2.76 & 0.00 & 0.00 \\ 
  43 & GO:0002237 & response to molecule of bacterial origin & 162 &  15 & 3.21 & 0.00 & 0.00 \\ 
  44 & GO:0050792 & regulation of viral process & 122 &  13 & 2.42 & 0.00 & 0.00 \\ 
  45 & GO:0048525 & negative regulation of viral process &  69 &  10 & 1.37 & 0.00 & 0.00 \\ 
  46 & GO:0045089 & positive regulation of innate immune res... & 127 &  13 & 2.52 & 0.00 & 0.00 \\ 
  47 & GO:0051250 & negative regulation of lymphocyte activa... &  90 &  11 & 1.78 & 0.00 & 0.00 \\ 
  48 & GO:0002684 & positive regulation of immune system pro... & 494 &  27 & 9.79 & 0.00 & 0.00 \\ 
  49 & GO:0002478 & antigen processing and presentation of e... &  20 &   6 & 0.40 & 0.00 & 0.00 \\ 
  50 & GO:0045824 & negative regulation of innate immune res... &  31 &   7 & 0.61 & 0.00 & 0.00 \\ 
  51 & GO:0050778 & positive regulation of immune response & 298 &  20 & 5.91 & 0.00 & 0.00 \\ 
  52 & GO:0032606 & type I interferon production &  45 &   8 & 0.89 & 0.00 & 0.00 \\ 
  53 & GO:0080134 & regulation of response to stress & 830 &  37 & 16.46 & 0.00 & 0.00 \\ 
  54 & GO:0031349 & positive regulation of defense response & 180 &  15 & 3.57 & 0.00 & 0.00 \\ 
  55 & GO:0045071 & negative regulation of viral genome repl... &  33 &   7 & 0.65 & 0.00 & 0.00 \\ 
  56 & GO:0048002 & antigen processing and presentation of p... &  47 &   8 & 0.93 & 0.00 & 0.00 \\ 
  57 & GO:0002695 & negative regulation of leukocyte activat... & 100 &  11 & 1.98 & 0.00 & 0.00 \\ 
  58 & GO:0001817 & regulation of cytokine production & 344 &  21 & 6.82 & 0.00 & 0.00 \\ 
  59 & GO:0019221 & cytokine-mediated signaling pathway & 188 &  15 & 3.73 & 0.00 & 0.00 \\ 
  60 & GO:0032607 & interferon-alpha production &  14 &   5 & 0.28 & 0.00 & 0.00 \\ 
  61 & GO:1903901 & negative regulation of viral life cycle &  66 &   9 & 1.31 & 0.00 & 0.00 \\ 
  62 & GO:0098586 & cellular response to virus &  24 &   6 & 0.48 & 0.00 & 0.00 \\ 
  63 & GO:0071310 & cellular response to organic substance & 1039 &  42 & 20.60 & 0.00 & 0.00 \\ 
  64 & GO:0002477 & antigen processing and presentation of e... &   3 &   3 & 0.06 & 0.00 & 0.00 \\ 
  65 & GO:0002481 & antigen processing and presentation of e... &   3 &   3 & 0.06 & 0.00 & 0.00 \\ 
  66 & GO:0043901 & negative regulation of multi-organism pr... & 108 &  11 & 2.14 & 0.00 & 0.00 \\ 
  67 & GO:0050777 & negative regulation of immune response &  89 &  10 & 1.76 & 0.00 & 0.00 \\ 
  68 & GO:0050866 & negative regulation of cell activation & 109 &  11 & 2.16 & 0.00 & 0.00 \\ 
  69 & GO:0016032 & viral process & 200 &  15 & 3.97 & 0.00 & 0.00 \\ 
  70 & GO:0002483 & antigen processing and presentation of e... &   9 &   4 & 0.18 & 0.00 & 0.00 \\ 
  71 & GO:0002697 & regulation of immune effector process & 271 &  17 & 5.37 & 0.00 & 0.00 \\ 
  72 & GO:0019883 & antigen processing and presentation of e... &  10 &   4 & 0.20 & 0.00 & 0.00 \\ 
  73 & GO:0002428 & antigen processing and presentation of p... &   4 &   3 & 0.08 & 0.00 & 0.00 \\ 
  74 & GO:0045341 & MHC class I biosynthetic process &   4 &   3 & 0.08 & 0.00 & 0.00 \\ 
  75 & GO:0045343 & regulation of MHC class I biosynthetic p... &   4 &   3 & 0.08 & 0.00 & 0.00 \\ 
  76 & GO:0002221 & pattern recognition receptor signaling p... &  81 &   9 & 1.61 & 0.00 & 0.00 \\ 
  77 & GO:0042221 & response to chemical & 1868 &  61 & 37.04 & 0.00 & 0.00 \\ 
  78 & GO:0032496 & response to lipopolysaccharide & 148 &  12 & 2.93 & 0.00 & 0.00 \\ 
  79 & GO:0050830 & defense response to Gram-positive bacter... &  48 &   7 & 0.95 & 0.00 & 0.00 \\ 
  80 & GO:0002758 & innate immune response-activating signal... &  84 &   9 & 1.67 & 0.00 & 0.00 \\ 
  81 & GO:0002449 & lymphocyte mediated immunity & 150 &  12 & 2.97 & 0.00 & 0.00 \\ 
  82 & GO:0050868 & negative regulation of T cell activation &  66 &   8 & 1.31 & 0.00 & 0.00 \\ 
  83 & GO:0048584 & positive regulation of response to stimu... & 1130 &  42 & 22.41 & 0.00 & 0.00 \\ 
  84 & GO:0070887 & cellular response to chemical stimulus & 1330 &  47 & 26.37 & 0.00 & 0.00 \\ 
  85 & GO:0051249 & regulation of lymphocyte activation & 259 &  16 & 5.14 & 0.00 & 0.00 \\ 
  86 & GO:0001819 & positive regulation of cytokine producti... & 232 &  15 & 4.60 & 0.00 & 0.00 \\ 
  87 & GO:0002716 & negative regulation of natural killer ce... &  12 &   4 & 0.24 & 0.00 & 0.00 \\ 
  88 & GO:0045953 & negative regulation of natural killer ce... &  12 &   4 & 0.24 & 0.00 & 0.00 \\ 
  89 & GO:0050896 & response to stimulus & 4441 & 117 & 88.06 & 0.00 & 0.00 \\ 
  90 & GO:1903038 & negative regulation of leukocyte cell-ce... &  71 &   8 & 1.41 & 0.00 & 0.00 \\ 
  91 & GO:0002218 & activation of innate immune response &  91 &   9 & 1.80 & 0.00 & 0.00 \\ 
  92 & GO:0031341 & regulation of cell killing &  53 &   7 & 1.05 & 0.00 & 0.00 \\ 
  93 & GO:0045069 & regulation of viral genome replication &  53 &   7 & 1.05 & 0.00 & 0.00 \\ 
  94 & GO:1903900 & regulation of viral life cycle & 113 &  10 & 2.24 & 0.00 & 0.00 \\ 
  95 & GO:0002753 & cytoplasmic pattern recognition receptor... &  24 &   5 & 0.48 & 0.00 & 0.00 \\ 
  96 & GO:0032647 & regulation of interferon-alpha productio... &  13 &   4 & 0.26 & 0.00 & 0.00 \\ 
  97 & GO:0044406 & adhesion of symbiont to host &  13 &   4 & 0.26 & 0.00 & 0.00 \\ 
  98 & GO:0001906 & cell killing &  73 &   8 & 1.45 & 0.00 & 0.00 \\ 
  99 & GO:0019058 & viral life cycle & 163 &  12 & 3.23 & 0.00 & 0.00 \\ 
  100 & GO:0042267 & natural killer cell mediated cytotoxicit... &  39 &   6 & 0.77 & 0.00 & 0.00 \\ 
   \hline
\end{tabular}
\end{table}