### 2016-04-08 additional_files/tex/cluster2_BP_Mm_7dpiUp.tex

% latex table generated in R 3.2.2 by xtable 1.8-2 package
% Fri Apr  8 11:17:23 2016
\begin{table}[ht]
\centering
\begin{tabular}{rllrrrrr}
  \hline
 & GO.ID & Term & Annotated & Significant & Expected & p.value & adj.p \\ 
  \hline
1 & GO:0055114 & oxidation-reduction process & 650 &  28 & 9.76 & 0.00 & 0.00 \\ 
  2 & GO:0044281 & small molecule metabolic process & 1261 &  42 & 18.94 & 0.00 & 0.00 \\ 
  3 & GO:0006629 & lipid metabolic process & 796 &  29 & 11.95 & 0.00 & 0.00 \\ 
  4 & GO:0044710 & single-organism metabolic process & 3313 &  76 & 49.76 & 0.00 & 0.00 \\ 
  5 & GO:0006805 & xenobiotic metabolic process &  42 &   6 & 0.63 & 0.00 & 0.00 \\ 
  6 & GO:0071466 & cellular response to xenobiotic stimulus &  46 &   6 & 0.69 & 0.00 & 0.00 \\ 
  7 & GO:0009410 & response to xenobiotic stimulus &  51 &   6 & 0.77 & 0.00 & 0.00 \\ 
  8 & GO:0006140 & regulation of nucleotide metabolic proce... & 103 &   8 & 1.55 & 0.00 & 0.00 \\ 
  9 & GO:1901387 & positive regulation of voltage-gated cal... &   2 &   2 & 0.03 & 0.00 & 0.00 \\ 
  10 & GO:0009117 & nucleotide metabolic process & 400 &  16 & 6.01 & 0.00 & 0.00 \\ 
  11 & GO:0017144 & drug metabolic process &  23 &   4 & 0.35 & 0.00 & 0.00 \\ 
  12 & GO:0019369 & arachidonic acid metabolic process &  24 &   4 & 0.36 & 0.00 & 0.00 \\ 
  13 & GO:0006753 & nucleoside phosphate metabolic process & 408 &  16 & 6.13 & 0.00 & 0.00 \\ 
  14 & GO:0008202 & steroid metabolic process & 169 &   9 & 2.54 & 0.00 & 0.01 \\ 
  15 & GO:0055086 & nucleobase-containing small molecule met... & 443 &  16 & 6.65 & 0.00 & 0.01 \\ 
  16 & GO:0015701 & bicarbonate transport &  14 &   3 & 0.21 & 0.00 & 0.01 \\ 
  17 & GO:0042738 & exogenous drug catabolic process &  14 &   3 & 0.21 & 0.00 & 0.01 \\ 
  18 & GO:0006631 & fatty acid metabolic process & 243 &  11 & 3.65 & 0.00 & 0.01 \\ 
  19 & GO:0051289 & protein homotetramerization &  53 &   5 & 0.80 & 0.00 & 0.01 \\ 
  20 & GO:0009992 & cellular water homeostasis &   4 &   2 & 0.06 & 0.00 & 0.01 \\ 
  21 & GO:0032782 & bile acid secretion &   4 &   2 & 0.06 & 0.00 & 0.01 \\ 
  22 & GO:0042737 & drug catabolic process &  15 &   3 & 0.23 & 0.00 & 0.01 \\ 
  23 & GO:0006690 & icosanoid metabolic process &  55 &   5 & 0.83 & 0.00 & 0.01 \\ 
  24 & GO:1901568 & fatty acid derivative metabolic process &  55 &   5 & 0.83 & 0.00 & 0.01 \\ 
  25 & GO:0015850 & organic hydroxy compound transport & 116 &   7 & 1.74 & 0.00 & 0.01 \\ 
  26 & GO:0043583 & ear development & 116 &   7 & 1.74 & 0.00 & 0.01 \\ 
  27 & GO:0051262 & protein tetramerization &  87 &   6 & 1.31 & 0.00 & 0.01 \\ 
  28 & GO:0055085 & transmembrane transport & 563 &  18 & 8.46 & 0.00 & 0.01 \\ 
  29 & GO:0045989 & positive regulation of striated muscle c... &   5 &   2 & 0.08 & 0.00 & 0.01 \\ 
  30 & GO:0042445 & hormone metabolic process &  89 &   6 & 1.34 & 0.00 & 0.01 \\ 
  31 & GO:0051186 & cofactor metabolic process & 265 &  11 & 3.98 & 0.00 & 0.01 \\ 
  32 & GO:0030104 & water homeostasis &  18 &   3 & 0.27 & 0.00 & 0.01 \\ 
  33 & GO:0071333 & cellular response to glucose stimulus &  62 &   5 & 0.93 & 0.00 & 0.01 \\ 
  34 & GO:0019637 & organophosphate metabolic process & 668 &  20 & 10.03 & 0.00 & 0.01 \\ 
  35 & GO:0046903 & secretion & 625 &  19 & 9.39 & 0.00 & 0.01 \\ 
  36 & GO:0071326 & cellular response to monosaccharide stim... &  64 &   5 & 0.96 & 0.00 & 0.01 \\ 
  37 & GO:0071331 & cellular response to hexose stimulus &  64 &   5 & 0.96 & 0.00 & 0.01 \\ 
  38 & GO:0019373 & epoxygenase P450 pathway &   6 &   2 & 0.09 & 0.00 & 0.01 \\ 
  39 & GO:0019532 & oxalate transport &   6 &   2 & 0.09 & 0.00 & 0.01 \\ 
  40 & GO:0048505 & regulation of timing of cell differentia... &   6 &   2 & 0.09 & 0.00 & 0.01 \\ 
  41 & GO:0071322 & cellular response to carbohydrate stimul... &  68 &   5 & 1.02 & 0.00 & 0.01 \\ 
  42 & GO:1900542 & regulation of purine nucleotide metaboli... &  98 &   6 & 1.47 & 0.00 & 0.01 \\ 
  43 & GO:0050433 & regulation of catecholamine secretion &  21 &   3 & 0.32 & 0.00 & 0.01 \\ 
  44 & GO:0015711 & organic anion transport & 207 &   9 & 3.11 & 0.00 & 0.01 \\ 
  45 & GO:0016042 & lipid catabolic process & 172 &   8 & 2.58 & 0.00 & 0.01 \\ 
  46 & GO:0003214 & cardiac left ventricle morphogenesis &   7 &   2 & 0.11 & 0.00 & 0.01 \\ 
  47 & GO:0006833 & water transport &   7 &   2 & 0.11 & 0.00 & 0.01 \\ 
  48 & GO:0032099 & negative regulation of appetite &   7 &   2 & 0.11 & 0.00 & 0.01 \\ 
  49 & GO:1902358 & sulfate transmembrane transport &   7 &   2 & 0.11 & 0.00 & 0.01 \\ 
  50 & GO:0001678 & cellular glucose homeostasis &  72 &   5 & 1.08 & 0.00 & 0.01 \\ 
  51 & GO:0032787 & monocarboxylic acid metabolic process & 377 &  13 & 5.66 & 0.00 & 0.01 \\ 
  52 & GO:0007200 & phospholipase C-activating G-protein cou... &  23 &   3 & 0.35 & 0.00 & 0.01 \\ 
  53 & GO:0050432 & catecholamine secretion &  23 &   3 & 0.35 & 0.00 & 0.01 \\ 
  54 & GO:0001676 & long-chain fatty acid metabolic process &  47 &   4 & 0.71 & 0.01 & 0.01 \\ 
  55 & GO:0032105 & negative regulation of response to extra... &  24 &   3 & 0.36 & 0.01 & 0.01 \\ 
  56 & GO:0032108 & negative regulation of response to nutri... &  24 &   3 & 0.36 & 0.01 & 0.01 \\ 
  57 & GO:0055082 & cellular chemical homeostasis & 342 &  12 & 5.14 & 0.01 & 0.01 \\ 
  58 & GO:0032612 & interleukin-1 production &  48 &   4 & 0.72 & 0.01 & 0.01 \\ 
  59 & GO:0006820 & anion transport & 302 &  11 & 4.54 & 0.01 & 0.01 \\ 
  60 & GO:0032096 & negative regulation of response to food &   8 &   2 & 0.12 & 0.01 & 0.01 \\ 
  61 & GO:0033604 & negative regulation of catecholamine sec... &   8 &   2 & 0.12 & 0.01 & 0.01 \\ 
  62 & GO:0040034 & regulation of development, heterochronic &   8 &   2 & 0.12 & 0.01 & 0.01 \\ 
  63 & GO:0060081 & membrane hyperpolarization &   8 &   2 & 0.12 & 0.01 & 0.01 \\ 
  64 & GO:0060979 & vasculogenesis involved in coronary vasc... &   8 &   2 & 0.12 & 0.01 & 0.01 \\ 
  65 & GO:0006730 & one-carbon metabolic process &  25 &   3 & 0.38 & 0.01 & 0.01 \\ 
  66 & GO:1901661 & quinone metabolic process &  25 &   3 & 0.38 & 0.01 & 0.01 \\ 
  67 & GO:1901135 & carbohydrate derivative metabolic proces... & 677 &  19 & 10.17 & 0.01 & 0.01 \\ 
  68 & GO:0034220 & ion transmembrane transport & 393 &  13 & 5.90 & 0.01 & 0.01 \\ 
  69 & GO:0030004 & cellular monovalent inorganic cation hom... &  50 &   4 & 0.75 & 0.01 & 0.01 \\ 
  70 & GO:0009187 & cyclic nucleotide metabolic process &  79 &   5 & 1.19 & 0.01 & 0.01 \\ 
  71 & GO:0010817 & regulation of hormone levels & 267 &  10 & 4.01 & 0.01 & 0.01 \\ 
  72 & GO:0008272 & sulfate transport &   9 &   2 & 0.14 & 0.01 & 0.01 \\ 
  73 & GO:0032148 & activation of protein kinase B activity &   9 &   2 & 0.14 & 0.01 & 0.01 \\ 
  74 & GO:0071875 & adrenergic receptor signaling pathway &   9 &   2 & 0.14 & 0.01 & 0.01 \\ 
  75 & GO:0052652 & cyclic purine nucleotide metabolic proce... &  52 &   4 & 0.78 & 0.01 & 0.01 \\ 
  76 & GO:0072521 & purine-containing compound metabolic pro... & 315 &  11 & 4.73 & 0.01 & 0.01 \\ 
  77 & GO:0009190 & cyclic nucleotide biosynthetic process &  53 &   4 & 0.80 & 0.01 & 0.01 \\ 
  78 & GO:0051937 & catecholamine transport &  28 &   3 & 0.42 & 0.01 & 0.01 \\ 
  79 & GO:0019318 & hexose metabolic process & 154 &   7 & 2.31 & 0.01 & 0.01 \\ 
  80 & GO:0045981 & positive regulation of nucleotide metabo... &  54 &   4 & 0.81 & 0.01 & 0.01 \\ 
  81 & GO:1900544 & positive regulation of purine nucleotide... &  54 &   4 & 0.81 & 0.01 & 0.01 \\ 
  82 & GO:0045980 & negative regulation of nucleotide metabo... &  29 &   3 & 0.44 & 0.01 & 0.01 \\ 
  83 & GO:0002275 & myeloid cell activation involved in immu... &  55 &   4 & 0.83 & 0.01 & 0.01 \\ 
  84 & GO:0032098 & regulation of appetite &  10 &   2 & 0.15 & 0.01 & 0.01 \\ 
  85 & GO:0051923 & sulfation &  10 &   2 & 0.15 & 0.01 & 0.01 \\ 
  86 & GO:0042493 & response to drug & 158 &   7 & 2.37 & 0.01 & 0.01 \\ 
   \hline
\end{tabular}
\end{table}