### 2016-04-14 additional_files/tex/SI_clusters/clusterTablesCombined.tex
Changes to heatmaps (more cluster in Ef) and enrichment, to include new clusters. Also started a .tex to summarize all GO-tables with legends. This is work in progress

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% University/School Laboratory Report
% LaTeX Template
% Version 3.1 (25/3/14)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Linux and Unix Users Group at Virginia Tech Wiki 
% (https://vtluug.org/wiki/Example_LaTeX_chem_lab_report)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%
%% Totta:
%% This collection aims to provide an overview of the enriched GO-terms in the
%% different clusters in the heatmaps. 
%% We include both mouse and Eimeria data

%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass{article}

%\usepackage[version=3]{mhchem} % Package for chemical equation typesetting
%\usepackage{siunitx} % Provides the \SI{}{} and \si{} command for typesetting SI units
\usepackage{graphicx} % Required for the inclusion of images
\usepackage[section]{placeins}
\usepackage[hscale=0.8,vscale=0.8]{geometry}
%\usepackage{natbib} % Required to change bibliography style to APA
%\usepackage{amsmath} % Required for some math elements

\setlength\parindent{0pt} % Removes all indentation from paragraphs

\renewcommand{\labelenumi}{\alph{enumi}.} % Make numbering in the enumerate environment by letter rather than number (e.g. section 6)

%\usepackage{times} % Uncomment to use the Times New Roman font

%----------------------------------------------------------------------------------------
%	DOCUMENT INFORMATION
%----------------------------------------------------------------------------------------

\title{GO-terms enriched in heatmap gene clusters for mouse and \textit{Eimeria falciformis}} % Title

\author{Totta \textsc{Kasemo}} % Author name

\date{\today} % Date for the report

\begin{document}

\maketitle % Insert the title, author and date

\begin{center}
\begin{tabular}{l r}
%Date Performed: & January 1, 2012 \\ % Date the experiment was performed
%Partners: & James Smith \\ % Partner names
%& Mary Smith \\
%Instructor: & Professor Smith % Instructor/supervisor
\end{tabular}
\end{center}

% If you wish to include an abstract, uncomment the lines below
% \begin{abstract}
% Abstract text
% \end{abstract}

%----------------------------------------------------------------------------------------
%	SECTION 1
%----------------------------------------------------------------------------------------

\section{Objective}
bla bla bla
\newpage
%\begin{center}\ce{}\end{center}

%----------------------------------------------------------------------------------------
%	SECTION 2 
%----------------------------------------------------------------------------------------

\section{Results for \textit{E. falciformis}}
  Cluster 1, BP, Ef.
  Overrepresented GO biological process terms among parasite genes which are 
  highly abundant on day 7 p.i.. Compared to cluster 2, which also have a high
  abundance on day 7, these genes cluster more closely with genes in cluster 4, 
  which are highly expressed in oocysts.\\
  %\newpage
  \input{../cluster1_BP_Ef_Day7Up_1.tex}

  Cluster 2, BP, Ef.
  Overrepresented GO biological process terms among parasite genes which are 
  highly abundant on day 7 p.i.. Compared to cluster 1, which is also upregulated 
  on day 7 p.i., these genes cluster more closely with cluster 6, which contains
  genes highly abundant in sporozoites.\\
  \input{../cluster2_BP_Ef_Day7Up_2.tex}

  Cluster 1, MF, Ef.\\
  Overrepresented GO molecular function terms among parasite genes which are 
  highly abundant on day 7 p.i.. Compared to cluster 2, which also have a high
  abundance on day 7, these genes cluster more closely with genes in cluster 4, 
  which are highly expressed in oocysts.\\
  \input{../cluster1_MF_Ef_Day7Up_1.tex}

  Cluster 2, MF, Ef.
  Overrepresented GO molecular function terms among parasite genes which are 
  highly abundant on day 7 p.i.. Compared to cluster 1, which is also upregulated 
  on day 7 p.i., these genes cluster more closely with cluster 6, which contains
  genes highly abundant in sporozoites.\\
  \input{../cluster2_MF_Ef_Day7Up_2.tex}

  Cluster 3, BP, Ef.
  Overrepresented GO biological process terms among parasite genes which have a low
  abundance in sporulated oocysts and sporozoites. In mouse stages, i.e., day 3, 5, and 7 samples
  there is a tendency to upregulation among these genes.\\
  \input{../cluster3_BP_Ef_OocSporDown.tex}

  Cluster 3, MF, Ef.
  Overrepresented GO molecular function terms among parasite genes which have a low
  abundance in sporulated oocysts and sporozoites. In mouse stages, i.e., day 3, 5, and 7 samples
  there is a tendency to upregulation among these genes.\\
  \input{../cluster3_MF_Ef_OocSporDown.tex}

  Cluster 4, BP, Ef.
  Overrepresented GO biological process terms among parasite genes which are highly abundant in
  sporulated oocysts. These genes cluster most closely with cluster 1, in which genes are highly
  abundant on day 7 p.i.. In cluster 4, genes have below average abundance in all day 3 and 5 samples.\\
  \input{../cluster4_BP_Ef_OocystsUp.tex}

  Cluster 4, MF, Ef.
  Overrepresented GO molecular function terms among parasite genes which are highly abundant in
  sporulated oocysts. These genes cluster most closely with cluster 1, in which genes are highly
  abundant on day 7 p.i.. In cluster 4, genes have below average abundance in all day 3 and 5 samples.\\
  \input{../cluster4_MF_Ef_OocystsUp.tex}

  Cluster 5, BP, Ef.
  Overrepresented GO biological process terms among parasite genes which have low abundance
  in sporulated oocysts, sporozoites and on day 7 p.i.. These genes seem to increase in abundance
  upon infection and decrease late in infection, i.e., on day 7.\\
  \input{../cluster5_BP_Ef_OoSpoDown_day5Up.tex}

  Cluster 5, MF, Ef.
  Overrepresented GO molecular function terms among parasite genes which have low abundance
  in sporulated oocysts, sporozoites and on day 7 p.i.. These genes seem to increase in abundance
  upon infection and decrease late in infection, i.e., on day 7.\\
  \input{../cluster5_MF_Ef_OoSpoDown_day5Up.tex}

  Cluster 6, BP, Ef.
  Overrepresented GO biological process terms among parasite genes which are highly abundant in sporozoites
  but mainly downregulated in sporulated oocysts, apart from a few genes in the cluster.
  This cluster seems to distinguish sporozoites from sporulated oocysts and have an average mRNA abundance in
  all other samples.\\
  \input{../cluster6_BP_Ef_SporozoitesUp.tex}

  Cluster 6, MF, Ef.
  Overrepresented GO molecular function terms among parasite genes which are highly abundant in sporozoites
  but mainly downregulated in sporulated oocysts, apart from a few genes in the cluster.
  This cluster seems to distinguish sporozoites from sporulated oocysts and have an average mRNA abundance in
  all other samples.\\
  \input{../cluster6_MF_Ef_SporozoitesUp.tex}

  Cluster 7, BP, Ef.
  Overrepresented GO biological process terms among parasite genes which have low abundance in sporulated oocysts.
  In sporozoites and early in infection, i.e., day 3 and 5 p.i., some of these genes are highly abundant. On day
  7 p.i. these mRNAs have a below average abundance.\\
  \input{../cluster7_BP_Ef_Day7OocyDown_SporoUP.tex}

  Cluster 7, MF, Ef.
  Overrepresented GO molecular function terms among parasite genes which have low abundance in sporulated oocysts.
  In sporozoites and early in infection, i.e., day 3 and 5 p.i., some of these genes are highly abundant. On day
  7 p.i. these mRNAs have a below average abundance.\\
  \input{../cluster7_MF_Ef_Day7OocyDown_SporoUP.tex}

\newpage

\section{Results for mouse}
  Cluster 1, BP, Mm
  Overrepresented GO biological process terms among mouse genes which have low abundance pre infection and 
  become more abundant in most day 3 and 5 samples from immunecompetent mice, but less in immune deficient Rag1-/- mice. 
  The upregulation trend is most clear in three first infection day 5 samples, which also form a separate
  sample cluster. On day 7, most of these genes are not differentially abundant, or below average, i.e., they
  seem to peak on day 5 p.i..\\
  \input{../cluster1_BP_Mm_EarlyUp.tex}

  Cluster 1, MF, Mm.
  Overrepresented GO molecular function terms among mouse genes which have low abundance pre infection and 
  become more abundant in most day 3 and 5 samples from immunecompetent mice, but less in immune deficient Rag1-/- mice. 
  The upregulation trend is most clear in three first infection day 5 samples, which also form a separate
  sample cluster. On day 7, most of these genes are not differentially abundant, or below average, i.e., they
  seem to peak on day 5 p.i..\\
  \input{../cluster1_MF_Mm_EarlyUp.tex}

  Cluster 2, BP, Mm.
  Overrepresented GO biological process terms among mouse genes which are downregulated on day 7 p.i. and
  upregulated pre infection, i.e., these genes become more and more downregulated during infection.\\
  \input{../cluster2_BP_Mm_7dpiUp.tex}

  Cluster 2, MF, Mm.
  Overrepresented GO molecular function terms among mouse genes which are downregulated on day 7 p.i. and
  upregulated pre infection, i.e., these genes become more and more downregulated during infection.\\
  \input{../cluster2_MF_Mm_7dpiUp.tex}

  Cluster 3, BP, Mm.
  Overrepresented GO biological process terms among mouse genes which are upregulated pre infection and
  downregulated on day 7 p.i., except in one NMRI second infection sample, possibly indicating no or
  poor infection in this sample. As in cluster 2, these mRNAs are most highly abundant before infection and
  become less abundant over the course of infection.\\
  \input{../cluster3_BP_Mm_7dpiDown.tex}

  Cluster 3, MF, Mm.
  Overrepresented GO molecular function terms among mouse genes which are upregulated pre infection and
  downregulated on day 7 p.i., except in one NMRI second infection sample, possibly indicating no or
  poor infection in this sample. As in cluster 2, these mRNAs are most highly abundant before infection and
  become less abundant over the course of infection.\\
  \input{../cluster3_MF_Mm_7dpiDown.tex}

  Cluster 4, BP, Mm.
  Overrepresented GO biological process terms among mouse genes which are highly abundant on day 7 p.i.. 
  There is overall no clear difference between pre infection and days 3 or 5 among these genesi. A slight
  tendency to higher abundance in second infection samples can be seen among these genes.\\
  \input{../cluster4_BP_Mm_allInfDown.tex}

  Cluster 4, MF, Mm.
  Overrepresented GO molecular function terms among mouse genes which are highly abundant on day 7 p.i.. 
  There is overall no clear difference between pre infection and days 3 or 5 among these genesi. A slight
  tendency to higher abundance in second infection samples can be seen among these genes.\\
  \input{../cluster4_MF_Mm_allInfDown.tex}
% figure 1
%\begin{figure}[h]
%\begin{center}
%\includegraphics[width=0.8\textwidth]{distributions_mouseNocutoff} % Include the image placeholder.png
%\caption{Density distribution of raw transcript counts (log10) for all
%  mouse samples in analysis. All samples share the same bimodal
%  distribution trend by visual inspection. The first density peak
%  occur at $1 - 10^2$ transcripts, \textit{i.e.} transcripts which are
%  detected between 1 and 100 times in that sample. The second density
%  peak occurs in the range of $10^3 - 10^4$.}
%\end{center}
%\end{figure}

%----------------------------------------------------------------------------------------
%	SECTION 5
%----------------------------------------------------------------------------------------

\section{Conclusions}

Mouse\\

Parasite\\

%----------------------------------------------------------------------------------------
%	SECTION 6
%----------------------------------------------------------------------------------------

%\section{Answers to Definitions}

%\begin{enumerate}
%\begin{item}
%  The \emph{atomic weight of an element} is the relative weight of
%  one of its atoms compared to C-12 with a weight of
%  12.0000000$\ldots$, hydrogen with a weight of 1.008, to oxygen with
%  a weight of 16.00. Atomic weight is also the average weight of all
%  the atoms of that element as they occur in nature.
%\end{item}
%\begin{item}
%  The \emph{units of atomic weight} are two-fold, with an identical
%  numerical value. They are g/mole of atoms (or just g/mol) or
%  amu/atom.
%\end{item}
%\begin{item}
%  \emph{Percentage discrepancy} between an accepted (literature)
%  value and an experimental value is
%\begin{equation*}
%\frac{\mathrm{experimental\;result} - \mathrm{accepted\;result}}{\mathrm{accepted\;result}}
%\end{equation*}
%\end{item}
%\end{enumerate}

%----------------------------------------------------------------------------------------
%	BIBLIOGRAPHY
%----------------------------------------------------------------------------------------

%\bibliographystyle{apalike}

%\bibliography{sample}

%----------------------------------------------------------------------------------------

\end{document}