### 2016-04-14 additional_files/tex/cluster3_BP_Mm_7dpiDown.tex

% latex table generated in R 3.2.3 by xtable 1.8-2 package
% Thu Apr 14 15:01:22 2016
\begin{table}[ht]
\centering
\begin{tabular}{llrrrrr}
  \hline
GO.ID & Term & Annotated & Significant & Expected & p.value & adj.p \\ 
  \hline
GO:0006805 & xenobiotic metabolic process &  42 &  18 & 1.51 & 0.00 & 0.00 \\ 
  GO:0006082 & organic acid metabolic process & 661 &  68 & 23.81 & 0.00 & 0.00 \\ 
  GO:0044281 & small molecule metabolic process & 1261 & 101 & 45.43 & 0.00 & 0.00 \\ 
  GO:0032787 & monocarboxylic acid metabolic process & 377 &  49 & 13.58 & 0.00 & 0.00 \\ 
  GO:0071466 & cellular response to xenobiotic stimulus &  46 &  18 & 1.66 & 0.00 & 0.00 \\ 
  GO:0009813 & flavonoid biosynthetic process &  13 &  11 & 0.47 & 0.00 & 0.00 \\ 
  GO:0052696 & flavonoid glucuronidation &  13 &  11 & 0.47 & 0.00 & 0.00 \\ 
  GO:0043436 & oxoacid metabolic process & 652 &  66 & 23.49 & 0.00 & 0.00 \\ 
  GO:0019752 & carboxylic acid metabolic process & 612 &  63 & 22.05 & 0.00 & 0.00 \\ 
  GO:0009812 & flavonoid metabolic process &  14 &  11 & 0.50 & 0.00 & 0.00 \\ 
  GO:0052695 & cellular glucuronidation &  14 &  11 & 0.50 & 0.00 & 0.00 \\ 
  GO:0009410 & response to xenobiotic stimulus &  51 &  18 & 1.84 & 0.00 & 0.00 \\ 
  GO:0006063 & uronic acid metabolic process &  15 &  11 & 0.54 & 0.00 & 0.00 \\ 
  GO:0019585 & glucuronate metabolic process &  15 &  11 & 0.54 & 0.00 & 0.00 \\ 
  GO:0052697 & xenobiotic glucuronidation &   8 &   8 & 0.29 & 0.00 & 0.00 \\ 
  GO:0006629 & lipid metabolic process & 796 &  66 & 28.68 & 0.00 & 0.00 \\ 
  GO:0009719 & response to endogenous stimulus & 729 &  60 & 26.26 & 0.00 & 0.00 \\ 
  GO:0071495 & cellular response to endogenous stimulus & 561 &  50 & 20.21 & 0.00 & 0.00 \\ 
  GO:0009725 & response to hormone & 385 &  37 & 13.87 & 0.00 & 0.00 \\ 
  GO:0006631 & fatty acid metabolic process & 243 &  28 & 8.75 & 0.00 & 0.00 \\ 
  GO:0044255 & cellular lipid metabolic process & 604 &  49 & 21.76 & 0.00 & 0.00 \\ 
  GO:0055114 & oxidation-reduction process & 650 &  50 & 23.42 & 0.00 & 0.00 \\ 
  GO:0042445 & hormone metabolic process &  89 &  15 & 3.21 & 0.00 & 0.00 \\ 
  GO:0032870 & cellular response to hormone stimulus & 278 &  28 & 10.02 & 0.00 & 0.00 \\ 
  GO:0044710 & single-organism metabolic process & 3313 & 163 & 119.36 & 0.00 & 0.00 \\ 
  GO:0034754 & cellular hormone metabolic process &  43 &  10 & 1.55 & 0.00 & 0.00 \\ 
  GO:0005996 & monosaccharide metabolic process & 177 &  20 & 6.38 & 0.00 & 0.00 \\ 
  GO:0042178 & xenobiotic catabolic process &   9 &   5 & 0.32 & 0.00 & 0.00 \\ 
  GO:0042537 & benzene-containing compound metabolic pr... &  15 &   6 & 0.54 & 0.00 & 0.00 \\ 
  GO:0006790 & sulfur compound metabolic process & 200 &  21 & 7.21 & 0.00 & 0.00 \\ 
  GO:0019395 & fatty acid oxidation &  66 &  11 & 2.38 & 0.00 & 0.00 \\ 
  GO:0034440 & lipid oxidation &  67 &  11 & 2.41 & 0.00 & 0.00 \\ 
  GO:0044699 & single-organism process & 7965 & 325 & 286.96 & 0.00 & 0.00 \\ 
  GO:0042221 & response to chemical & 1868 &  99 & 67.30 & 0.00 & 0.00 \\ 
  GO:1901615 & organic hydroxy compound metabolic proce... & 299 &  26 & 10.77 & 0.00 & 0.00 \\ 
  GO:0010817 & regulation of hormone levels & 267 &  24 & 9.62 & 0.00 & 0.00 \\ 
  GO:0016042 & lipid catabolic process & 172 &  18 & 6.20 & 0.00 & 0.00 \\ 
  GO:0006811 & ion transport & 785 &  50 & 28.28 & 0.00 & 0.00 \\ 
  GO:0022600 & digestive system process &  51 &   9 & 1.84 & 0.00 & 0.00 \\ 
  GO:0044242 & cellular lipid catabolic process & 117 &  14 & 4.22 & 0.00 & 0.00 \\ 
  GO:0042738 & exogenous drug catabolic process &  14 &   5 & 0.50 & 0.00 & 0.00 \\ 
  GO:0051386 & regulation of neurotrophin TRK receptor ... &   8 &   4 & 0.29 & 0.00 & 0.00 \\ 
  GO:0044344 & cellular response to fibroblast growth f... &  42 &   8 & 1.51 & 0.00 & 0.00 \\ 
  GO:0008202 & steroid metabolic process & 169 &  17 & 6.09 & 0.00 & 0.00 \\ 
  GO:0017144 & drug metabolic process &  23 &   6 & 0.83 & 0.00 & 0.00 \\ 
  GO:0042737 & drug catabolic process &  15 &   5 & 0.54 & 0.00 & 0.00 \\ 
  GO:0005975 & carbohydrate metabolic process & 481 &  34 & 17.33 & 0.00 & 0.00 \\ 
  GO:0072329 & monocarboxylic acid catabolic process &  68 &  10 & 2.45 & 0.00 & 0.00 \\ 
  GO:0071774 & response to fibroblast growth factor &  44 &   8 & 1.59 & 0.00 & 0.00 \\ 
  GO:0007586 & digestion &  56 &   9 & 2.02 & 0.00 & 0.00 \\ 
  GO:0072001 & renal system development & 141 &  15 & 5.08 & 0.00 & 0.00 \\ 
  GO:0044539 & long-chain fatty acid import &   4 &   3 & 0.14 & 0.00 & 0.00 \\ 
  GO:0051387 & negative regulation of neurotrophin TRK ... &   4 &   3 & 0.14 & 0.00 & 0.00 \\ 
  GO:0042493 & response to drug & 158 &  16 & 5.69 & 0.00 & 0.00 \\ 
  GO:0006672 & ceramide metabolic process &  58 &   9 & 2.09 & 0.00 & 0.00 \\ 
  GO:0008543 & fibroblast growth factor receptor signal... &  35 &   7 & 1.26 & 0.00 & 0.00 \\ 
  GO:0044723 & single-organism carbohydrate metabolic p... & 435 &  31 & 15.67 & 0.00 & 0.00 \\ 
  GO:0007267 & cell-cell signaling & 456 &  32 & 16.43 & 0.00 & 0.00 \\ 
  GO:0070887 & cellular response to chemical stimulus & 1330 &  72 & 47.92 & 0.00 & 0.00 \\ 
  GO:0001822 & kidney development & 131 &  14 & 4.72 & 0.00 & 0.00 \\ 
  GO:0006040 & amino sugar metabolic process &  26 &   6 & 0.94 & 0.00 & 0.00 \\ 
  GO:0034762 & regulation of transmembrane transport & 180 &  17 & 6.48 & 0.00 & 0.00 \\ 
  GO:0042572 & retinol metabolic process &  10 &   4 & 0.36 & 0.00 & 0.00 \\ 
  GO:0043269 & regulation of ion transport & 306 &  24 & 11.02 & 0.00 & 0.00 \\ 
  GO:0009062 & fatty acid catabolic process &  61 &   9 & 2.20 & 0.00 & 0.00 \\ 
  GO:0044282 & small molecule catabolic process & 183 &  17 & 6.59 & 0.00 & 0.00 \\ 
  GO:0001655 & urogenital system development & 168 &  16 & 6.05 & 0.00 & 0.00 \\ 
  GO:0022603 & regulation of anatomical structure morph... & 572 &  37 & 20.61 & 0.00 & 0.00 \\ 
  GO:0044763 & single-organism cellular process & 7155 & 291 & 257.78 & 0.00 & 0.00 \\ 
  GO:0006637 & acyl-CoA metabolic process &  52 &   8 & 1.87 & 0.00 & 0.00 \\ 
  GO:0035383 & thioester metabolic process &  52 &   8 & 1.87 & 0.00 & 0.00 \\ 
  GO:0006732 & coenzyme metabolic process & 227 &  19 & 8.18 & 0.00 & 0.00 \\ 
  GO:0016054 & organic acid catabolic process & 126 &  13 & 4.54 & 0.00 & 0.00 \\ 
  GO:0046395 & carboxylic acid catabolic process & 126 &  13 & 4.54 & 0.00 & 0.00 \\ 
  GO:0015698 & inorganic anion transport &  67 &   9 & 2.41 & 0.00 & 0.00 \\ 
  GO:0007169 & transmembrane receptor protein tyrosine ... & 324 &  24 & 11.67 & 0.00 & 0.00 \\ 
  GO:0009308 & amine metabolic process &  82 &  10 & 2.95 & 0.00 & 0.00 \\ 
  GO:0048011 & neurotrophin TRK receptor signaling path... &  22 &   5 & 0.79 & 0.00 & 0.00 \\ 
  GO:0030258 & lipid modification & 134 &  13 & 4.83 & 0.00 & 0.00 \\ 
  GO:0034765 & regulation of ion transmembrane transpor... & 168 &  15 & 6.05 & 0.00 & 0.00 \\ 
  GO:0055085 & transmembrane transport & 563 &  35 & 20.28 & 0.00 & 0.00 \\ 
  GO:0006576 & cellular biogenic amine metabolic proces... &  59 &   8 & 2.13 & 0.00 & 0.00 \\ 
  GO:0032941 & secretion by tissue &  46 &   7 & 1.66 & 0.00 & 0.00 \\ 
  GO:0043537 & negative regulation of blood vessel endo... &  14 &   4 & 0.50 & 0.00 & 0.00 \\ 
  GO:0060602 & branch elongation of an epithelium &  14 &   4 & 0.50 & 0.00 & 0.00 \\ 
  GO:0018879 & biphenyl metabolic process &   2 &   2 & 0.07 & 0.00 & 0.00 \\ 
  GO:0046168 & glycerol-3-phosphate catabolic process &   2 &   2 & 0.07 & 0.00 & 0.00 \\ 
  GO:0061031 & endodermal digestive tract morphogenesis &   2 &   2 & 0.07 & 0.00 & 0.00 \\ 
  GO:0097089 & methyl-branched fatty acid metabolic pro... &   2 &   2 & 0.07 & 0.00 & 0.00 \\ 
  GO:0097623 & potassium ion export across plasma membr... &   2 &   2 & 0.07 & 0.00 & 0.00 \\ 
  GO:0001676 & long-chain fatty acid metabolic process &  47 &   7 & 1.69 & 0.00 & 0.00 \\ 
  GO:0007268 & synaptic transmission & 245 &  19 & 8.83 & 0.00 & 0.00 \\ 
  GO:0071310 & cellular response to organic substance & 1039 &  56 & 37.43 & 0.00 & 0.00 \\ 
  GO:0019755 & one-carbon compound transport &   7 &   3 & 0.25 & 0.00 & 0.00 \\ 
  GO:0060736 & prostate gland growth &   7 &   3 & 0.25 & 0.00 & 0.00 \\ 
  GO:0090331 & negative regulation of platelet aggregat... &   7 &   3 & 0.25 & 0.00 & 0.00 \\ 
  GO:0008610 & lipid biosynthetic process & 383 &  26 & 13.80 & 0.00 & 0.00 \\ 
  GO:0009755 & hormone-mediated signaling pathway &  61 &   8 & 2.20 & 0.00 & 0.00 \\ 
  GO:0007167 & enzyme linked receptor protein signaling... & 487 &  31 & 17.55 & 0.00 & 0.00 \\ 
  GO:0016266 & O-glycan processing &  15 &   4 & 0.54 & 0.00 & 0.00 \\ 
   \hline
\end{tabular}
\end{table}