### 2016-04-14 additional_files/tex/cluster3_MF_Ef_OocSporDown.tex

% latex table generated in R 3.2.3 by xtable 1.8-2 package
% Thu Apr 14 14:57:26 2016
\begin{table}[ht]
\centering
\begin{tabular}{llrrrrr}
  \hline
GO.ID & Term & Annotated & Significant & Expected & p.value & adj.p \\ 
  \hline
GO:0004312 & fatty acid synthase activity &   2 &   2 & 0.09 & 0.00 & 0.07 \\ 
  GO:0051920 & peroxiredoxin activity &   2 &   2 & 0.09 & 0.00 & 0.07 \\ 
  GO:0003824 & catalytic activity & 1318 &  74 & 58.58 & 0.00 & 0.07 \\ 
  GO:0016746 & transferase activity, transferring acyl ... &  46 &   7 & 2.04 & 0.00 & 0.08 \\ 
  GO:0019205 & nucleobase-containing compound kinase ac... &   8 &   3 & 0.36 & 0.00 & 0.08 \\ 
  GO:0016615 & malate dehydrogenase activity &   3 &   2 & 0.13 & 0.01 & 0.09 \\ 
  GO:0003746 & translation elongation factor activity &  10 &   3 & 0.44 & 0.01 & 0.09 \\ 
  GO:0008135 & translation factor activity, RNA binding &  30 &   5 & 1.33 & 0.01 & 0.09 \\ 
   \hline
\end{tabular}
\end{table}