### 2016-04-14 additional_files/tex/cluster4_BP_Mm_allInfDown.tex

% latex table generated in R 3.2.3 by xtable 1.8-2 package
% Thu Apr 14 15:04:34 2016
\begin{table}[ht]
\centering
\begin{tabular}{llrrrrr}
  \hline
GO.ID & Term & Annotated & Significant & Expected & p.value & adj.p \\ 
  \hline
GO:0002376 & immune system process & 1406 & 126 & 41.64 & 0.00 & 0.00 \\ 
  GO:0006955 & immune response & 713 &  85 & 21.12 & 0.00 & 0.00 \\ 
  GO:0050896 & response to stimulus & 4441 & 229 & 131.52 & 0.00 & 0.00 \\ 
  GO:0006954 & inflammatory response & 327 &  53 & 9.68 & 0.00 & 0.00 \\ 
  GO:0044700 & single organism signaling & 2975 & 176 & 88.11 & 0.00 & 0.00 \\ 
  GO:0006952 & defense response & 755 &  80 & 22.36 & 0.00 & 0.00 \\ 
  GO:0023052 & signaling & 2981 & 176 & 88.28 & 0.00 & 0.00 \\ 
  GO:0007165 & signal transduction & 2773 & 167 & 82.12 & 0.00 & 0.00 \\ 
  GO:0001775 & cell activation & 589 &  69 & 17.44 & 0.00 & 0.00 \\ 
  GO:0007154 & cell communication & 3137 & 178 & 92.90 & 0.00 & 0.00 \\ 
  GO:0022610 & biological adhesion & 800 &  79 & 23.69 & 0.00 & 0.00 \\ 
  GO:0002684 & positive regulation of immune system pro... & 494 &  61 & 14.63 & 0.00 & 0.00 \\ 
  GO:0007155 & cell adhesion & 791 &  78 & 23.43 & 0.00 & 0.00 \\ 
  GO:0002682 & regulation of immune system process & 774 &  76 & 22.92 & 0.00 & 0.00 \\ 
  GO:0001816 & cytokine production & 397 &  53 & 11.76 & 0.00 & 0.00 \\ 
  GO:0065007 & biological regulation & 6468 & 274 & 191.55 & 0.00 & 0.00 \\ 
  GO:0048518 & positive regulation of biological proces... & 3404 & 182 & 100.81 & 0.00 & 0.00 \\ 
  GO:0045321 & leukocyte activation & 515 &  60 & 15.25 & 0.00 & 0.00 \\ 
  GO:0040011 & locomotion & 903 &  79 & 26.74 & 0.00 & 0.00 \\ 
  GO:0098602 & single organism cell adhesion & 516 &  58 & 15.28 & 0.00 & 0.00 \\ 
  GO:0051239 & regulation of multicellular organismal p... & 1538 & 108 & 45.55 & 0.00 & 0.00 \\ 
  GO:0007159 & leukocyte cell-cell adhesion & 332 &  46 & 9.83 & 0.00 & 0.00 \\ 
  GO:0016337 & single organismal cell-cell adhesion & 474 &  55 & 14.04 & 0.00 & 0.00 \\ 
  GO:0051716 & cellular response to stimulus & 3741 & 189 & 110.79 & 0.00 & 0.00 \\ 
  GO:0098609 & cell-cell adhesion & 507 &  56 & 15.01 & 0.00 & 0.00 \\ 
  GO:0044707 & single-multicellular organism process & 3409 & 176 & 100.96 & 0.00 & 0.00 \\ 
  GO:0050789 & regulation of biological process & 6161 & 259 & 182.46 & 0.00 & 0.00 \\ 
  GO:0048870 & cell motility & 784 &  70 & 23.22 & 0.00 & 0.00 \\ 
  GO:0051674 & localization of cell & 784 &  70 & 23.22 & 0.00 & 0.00 \\ 
  GO:0032501 & multicellular organismal process & 3487 & 177 & 103.27 & 0.00 & 0.00 \\ 
  GO:0050794 & regulation of cellular process & 5853 & 249 & 173.34 & 0.00 & 0.00 \\ 
  GO:0016477 & cell migration & 740 &  67 & 21.92 & 0.00 & 0.00 \\ 
  GO:0001817 & regulation of cytokine production & 344 &  44 & 10.19 & 0.00 & 0.00 \\ 
  GO:0034109 & homotypic cell-cell adhesion & 361 &  45 & 10.69 & 0.00 & 0.00 \\ 
  GO:0048583 & regulation of response to stimulus & 2109 & 126 & 62.46 & 0.00 & 0.00 \\ 
  GO:0048584 & positive regulation of response to stimu... & 1130 &  85 & 33.47 & 0.00 & 0.00 \\ 
  GO:0050865 & regulation of cell activation & 319 &  42 & 9.45 & 0.00 & 0.00 \\ 
  GO:0048522 & positive regulation of cellular process & 3048 & 160 & 90.27 & 0.00 & 0.00 \\ 
  GO:0046649 & lymphocyte activation & 435 &  49 & 12.88 & 0.00 & 0.00 \\ 
  GO:0035556 & intracellular signal transduction & 1570 & 103 & 46.50 & 0.00 & 0.00 \\ 
  GO:0002694 & regulation of leukocyte activation & 301 &  40 & 8.91 & 0.00 & 0.00 \\ 
  GO:0070486 & leukocyte aggregation & 314 &  40 & 9.30 & 0.00 & 0.00 \\ 
  GO:1902531 & regulation of intracellular signal trans... & 991 &  76 & 29.35 & 0.00 & 0.00 \\ 
  GO:0050900 & leukocyte migration & 164 &  29 & 4.86 & 0.00 & 0.00 \\ 
  GO:0009605 & response to external stimulus & 1176 &  83 & 34.83 & 0.00 & 0.00 \\ 
  GO:0030155 & regulation of cell adhesion & 414 &  45 & 12.26 & 0.00 & 0.00 \\ 
  GO:0051249 & regulation of lymphocyte activation & 259 &  35 & 7.67 & 0.00 & 0.00 \\ 
  GO:0006928 & movement of cell or subcellular componen... & 992 &  74 & 29.38 & 0.00 & 0.00 \\ 
  GO:0001819 & positive regulation of cytokine producti... & 232 &  33 & 6.87 & 0.00 & 0.00 \\ 
  GO:0006950 & response to stress & 2191 & 123 & 64.89 & 0.00 & 0.00 \\ 
  GO:0060326 & cell chemotaxis & 145 &  26 & 4.29 & 0.00 & 0.00 \\ 
  GO:0050793 & regulation of developmental process & 1367 &  89 & 40.48 & 0.00 & 0.00 \\ 
  GO:0023051 & regulation of signaling & 1746 & 104 & 51.71 & 0.00 & 0.00 \\ 
  GO:0044763 & single-organism cellular process & 7155 & 274 & 211.90 & 0.00 & 0.00 \\ 
  GO:0006935 & chemotaxis & 270 &  34 & 8.00 & 0.00 & 0.00 \\ 
  GO:0008219 & cell death & 1372 &  88 & 40.63 & 0.00 & 0.00 \\ 
  GO:0042330 & taxis & 271 &  34 & 8.03 & 0.00 & 0.00 \\ 
  GO:0065008 & regulation of biological quality & 1978 & 112 & 58.58 & 0.00 & 0.00 \\ 
  GO:0016265 & death & 1376 &  88 & 40.75 & 0.00 & 0.00 \\ 
  GO:0070887 & cellular response to chemical stimulus & 1330 &  86 & 39.39 & 0.00 & 0.00 \\ 
  GO:0042110 & T cell activation & 305 &  36 & 9.03 & 0.00 & 0.00 \\ 
  GO:0070489 & T cell aggregation & 305 &  36 & 9.03 & 0.00 & 0.00 \\ 
  GO:0010646 & regulation of cell communication & 1781 & 104 & 52.74 & 0.00 & 0.00 \\ 
  GO:0071593 & lymphocyte aggregation & 307 &  36 & 9.09 & 0.00 & 0.00 \\ 
  GO:0009966 & regulation of signal transduction & 1557 &  95 & 46.11 & 0.00 & 0.00 \\ 
  GO:0051240 & positive regulation of multicellular org... & 890 &  66 & 26.36 & 0.00 & 0.00 \\ 
  GO:0048731 & system development & 2399 & 126 & 71.05 & 0.00 & 0.00 \\ 
  GO:0048856 & anatomical structure development & 2919 & 144 & 86.45 & 0.00 & 0.00 \\ 
  GO:0007166 & cell surface receptor signaling pathway & 1300 &  83 & 38.50 & 0.00 & 0.00 \\ 
  GO:0044699 & single-organism process & 7965 & 292 & 235.89 & 0.00 & 0.00 \\ 
  GO:0040017 & positive regulation of locomotion & 276 &  33 & 8.17 & 0.00 & 0.00 \\ 
  GO:0030154 & cell differentiation & 2151 & 116 & 63.70 & 0.00 & 0.00 \\ 
  GO:0002520 & immune system development & 623 &  52 & 18.45 & 0.00 & 0.00 \\ 
  GO:0002440 & production of molecular mediator of immu... & 111 &  21 & 3.29 & 0.00 & 0.00 \\ 
  GO:0048534 & hematopoietic or lymphoid organ developm... & 586 &  50 & 17.35 & 0.00 & 0.00 \\ 
  GO:0002521 & leukocyte differentiation & 347 &  37 & 10.28 & 0.00 & 0.00 \\ 
  GO:0050663 & cytokine secretion & 101 &  20 & 2.99 & 0.00 & 0.00 \\ 
  GO:0030595 & leukocyte chemotaxis & 103 &  20 & 3.05 & 0.00 & 0.00 \\ 
  GO:1902533 & positive regulation of intracellular sig... & 556 &  48 & 16.47 & 0.00 & 0.00 \\ 
  GO:0050867 & positive regulation of cell activation & 195 &  27 & 5.77 & 0.00 & 0.00 \\ 
  GO:0042221 & response to chemical & 1868 & 104 & 55.32 & 0.00 & 0.00 \\ 
  GO:0002250 & adaptive immune response & 226 &  29 & 6.69 & 0.00 & 0.00 \\ 
  GO:0097529 & myeloid leukocyte migration &  94 &  19 & 2.78 & 0.00 & 0.00 \\ 
  GO:0030335 & positive regulation of cell migration & 258 &  31 & 7.64 & 0.00 & 0.00 \\ 
  GO:0044767 & single-organism developmental process & 3254 & 153 & 96.37 & 0.00 & 0.00 \\ 
  GO:0032879 & regulation of localization & 1499 &  89 & 44.39 & 0.00 & 0.00 \\ 
  GO:0009306 & protein secretion & 311 &  34 & 9.21 & 0.00 & 0.00 \\ 
  GO:0051270 & regulation of cellular component movemen... & 494 &  44 & 14.63 & 0.00 & 0.00 \\ 
  GO:2000147 & positive regulation of cell motility & 264 &  31 & 7.82 & 0.00 & 0.00 \\ 
  GO:0022407 & regulation of cell-cell adhesion & 249 &  30 & 7.37 & 0.00 & 0.00 \\ 
  GO:0097530 & granulocyte migration &  66 &  16 & 1.95 & 0.00 & 0.00 \\ 
  GO:0012501 & programmed cell death & 1290 &  80 & 38.20 & 0.00 & 0.00 \\ 
  GO:1990266 & neutrophil migration &  57 &  15 & 1.69 & 0.00 & 0.00 \\ 
  GO:0002696 & positive regulation of leukocyte activat... & 190 &  26 & 5.63 & 0.00 & 0.00 \\ 
  GO:0006915 & apoptotic process & 1267 &  79 & 37.52 & 0.00 & 0.00 \\ 
  GO:0002252 & immune effector process & 461 &  42 & 13.65 & 0.00 & 0.00 \\ 
  GO:0032502 & developmental process & 3284 & 153 & 97.26 & 0.00 & 0.00 \\ 
  GO:0032940 & secretion by cell & 559 &  47 & 16.55 & 0.00 & 0.00 \\ 
  GO:0009617 & response to bacterium & 253 &  30 & 7.49 & 0.00 & 0.00 \\ 
  GO:0051272 & positive regulation of cellular componen... & 270 &  31 & 8.00 & 0.00 & 0.00 \\ 
   \hline
\end{tabular}
\end{table}