### 2016-04-14 additional_files/tex/cluster6_BP_Ef_SporozoitesUp.tex

% latex table generated in R 3.2.3 by xtable 1.8-2 package
% Thu Apr 14 14:57:52 2016
\begin{table}[ht]
\centering
\begin{tabular}{llrrrrr}
  \hline
GO.ID & Term & Annotated & Significant & Expected & p.value & adj.p \\ 
  \hline
GO:0007154 & cell communication &  63 &   8 & 1.68 & 0.00 & 0.02 \\ 
  GO:0007165 & signal transduction &  58 &   7 & 1.54 & 0.00 & 0.02 \\ 
  GO:0023052 & signaling &  58 &   7 & 1.54 & 0.00 & 0.02 \\ 
  GO:0044700 & single organism signaling &  58 &   7 & 1.54 & 0.00 & 0.02 \\ 
  GO:0007186 & G-protein coupled receptor signaling pat... &   3 &   2 & 0.08 & 0.00 & 0.03 \\ 
  GO:0007205 & protein kinase C-activating G-protein co... &   3 &   2 & 0.08 & 0.00 & 0.03 \\ 
  GO:0051716 & cellular response to stimulus &  99 &   8 & 2.64 & 0.00 & 0.05 \\ 
  GO:0009405 & pathogenesis &   5 &   2 & 0.13 & 0.01 & 0.07 \\ 
  GO:0051704 & multi-organism process &   5 &   2 & 0.13 & 0.01 & 0.07 \\ 
   \hline
\end{tabular}
\end{table}