### 2016-04-14 additional_files/tex/cluster7_MF_Ef_Day7OocyDown_SporoUP.tex

% latex table generated in R 3.2.3 by xtable 1.8-2 package
% Thu Apr 14 14:57:55 2016
\begin{table}[ht]
\centering
\begin{tabular}{llrrrrr}
  \hline
GO.ID & Term & Annotated & Significant & Expected & p.value & adj.p \\ 
  \hline
GO:0052689 & carboxylic ester hydrolase activity &  12 &   5 & 0.56 & 0.00 & 0.01 \\ 
  GO:0016307 & phosphatidylinositol phosphate kinase ac... &   4 &   3 & 0.19 & 0.00 & 0.02 \\ 
  GO:0004725 & protein tyrosine phosphatase activity &   6 &   3 & 0.28 & 0.00 & 0.05 \\ 
  GO:0004812 & aminoacyl-tRNA ligase activity &  41 &   7 & 1.92 & 0.00 & 0.05 \\ 
  GO:0016875 & ligase activity, forming carbon-oxygen b... &  42 &   7 & 1.96 & 0.00 & 0.05 \\ 
  GO:0016876 & ligase activity, forming aminoacyl-tRNA ... &  42 &   7 & 1.96 & 0.00 & 0.05 \\ 
  GO:0008026 & ATP-dependent helicase activity &  60 &   8 & 2.81 & 0.01 & 0.08 \\ 
  GO:0070035 & purine NTP-dependent helicase activity &  60 &   8 & 2.81 & 0.01 & 0.08 \\ 
   \hline
\end{tabular}
\end{table}