### 2016-05-19 manuscript/tex_bmc_format/manuscript_rnaseq.tex
working on GO clusters and genes and some text in draft manuscript

%% BioMed_Central_Tex_Template_v1.06
%%                                      %
%  bmc_article.tex            ver: 1.06 %
%                                       %

%%IMPORTANT: do not delete the first line of this template
%%It must be present to enable the BMC Submission system to
%%recognise this template!!

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                     %%
%%  LaTeX template for BioMed Central  %%
%%     journal article submissions     %%
%%                                     %%
%%          <8 June 2012>              %%
%%                                     %%
%%                                     %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                                                 %%
%% For instructions on how to fill out this Tex template           %%
%% document please refer to Readme.html and the instructions for   %%
%% authors page on the biomed central website                      %%
%% http://www.biomedcentral.com/info/authors/                      %%
%%                                                                 %%
%% Please do not use \input{...} to include other tex files.       %%
%% Submit your LaTeX manuscript as one .tex document.              %%
%%                                                                 %%
%% All additional figures and files should be attached             %%
%% separately and not embedded in the \TeX\ document itself.       %%
%%                                                                 %%
%% BioMed Central currently use the MikTex distribution of         %%
%% TeX for Windows) of TeX and LaTeX.  This is available from      %%
%% http://www.miktex.org                                           %%
%%                                                                 %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%% additional documentclass options:
%  [doublespacing]
%  [linenumbers]   - put the line numbers on margins

%%% loading packages, author definitions

%\documentclass[twocolumn]{bmcart}% uncomment this for twocolumn layout and comment line below
\documentclass{bmcart}

%%% Load packages
%\usepackage{amsthm,amsmath}
%\RequirePackage{natbib}
%\RequirePackage[authoryear]{natbib}% uncomment this for author-year bibliography
%\RequirePackage{hyperref}
%\usepackage[utf8]{inputenc} %unicode support
%\usepackage[applemac]{inputenc} %applemac support if unicode package fails
%%\usepackage[latin1]{inputenc} %UNIX support if unicode package fails
%\usepackage{siunitx}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                             %%
%%  If you wish to display your graphics for   %%
%%  your own use using includegraphic or       %%
%%  includegraphics, then comment out the      %%
%%  following two lines of code.               %%
%%  NB: These line *must* be included when     %%
%%  submitting to BMC.                         %%
%%  All figure files must be submitted as      %%
%%  separate graphics through the BMC          %%
%%  submission process, not included in the    %%
%%  submitted article.                         %%
%%                                             %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{graphicx}
\graphicspath{ {../figures/} }
%\def\includegraphic{}
%\def\includegraphics{}

%%% Put your definitions there:
\startlocaldefs
%\affi1=Department of Molecular Parasitology, Humboldt University, Berlin, Germany
%\affi2=Max Planck Institute for Biology of Ageing, Joseph-Stelzmann-Strasse 9B, 50931 Cologne, Germany
%\affi3=Department Reproduction Biology and Evolutionary Ecology, Leibniz Institute of Zoo and Wildlife Research, Forschungsverbund Berlin e.V., Berlin, Germany
\endlocaldefs

%%% Begin ...
\begin{document}

%%% Start of article front matter
\begin{frontmatter}

\begin{fmbox}
\dochead{Research}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter the title of your article here     %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{A sample article title}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter the authors here                   %%
%%                                          %%
%% Specify information, if available,       %%
%% in the form:                             %%
%%   <key>={<id1>,<id2>}                    %%
%%   <key>=                                 %%
%% Comment or delete the keys which are     %%
%% not used. Repeat \author command as much %%
%% as required.                             %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\author[
   addressref={aff1},                   % id's of addresses, e.g. {aff1,aff2}
   corref={aff1},                       % id of corresponding address, if any
   noteref={n1},                        % id's of article notes, if any
   email={jane.e.doe@cambridge.co.uk}   % email address
]{\inits{JE}\fnm{Jane E} \snm{Doe}}
\author[
   addressref={aff1,aff2},
   email={john.RS.Smith@cambridge.co.uk}
]{\inits{JRS}\fnm{John RS} \snm{Smith}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter the authors' addresses here        %%
%%                                          %%
%% Repeat \address commands as much as      %%
%% required.                                %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\address[id=aff1]{%                           % unique id
  \orgname{Department of Zoology, Cambridge}, % university, etc
  \street{Waterloo Road},                     %
  %\postcode{}                                % post or zip code
  \city{London},                              % city
  \cny{UK}                                    % country
}
\address[id=aff2]{%
  \orgname{Marine Ecology Department, Institute of Marine Sciences Kiel},
  \street{D\"{u}sternbrooker Weg 20},
  \postcode{24105}
  \city{Kiel},
  \cny{Germany}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\author[
%   addressref={affi1},                   % id's of addresses, e.g. {aff1,aff2}
%   corref={aff1},                       % id of corresponding address, if any
%  % noteref={n1},                        % id's of article notes, if any
%   email={kasemotq@hu-berlin.de}   % email address
%]{\inits{TK}\fnm{Totta} \snm{Kasemo}}
%\author[
%   addressref={affi1}                   % id's of addresses, e.g. {aff1,aff2}
%  % corref={aff1},                       % id of corresponding address, if any
%  % noteref={n1},                        % id's of article notes, if any
%  % email={}   % email address
%]{\inits{SS}\fnm{Simone} \snm{Spork}}
%%\author[
%   addressref={affi2},                   % id's of addresses, e.g. {aff1,aff2}
%   %corref={aff3},                       % id of corresponding address, if any
%   noteref={n2},                        % id's of article notes, if any
%   %email={emanuel.heitlinger@hu-berlin.de}   % email address
%]{\inits{CD}\fnm{Christoph} \snm{Dieterich}}
%\author[
%   addressref={affi1, affi3},                   % id's of addresses, e.g. {aff1,aff2}
%   %corref={aff3},                       % id of corresponding address, if any
%   noteref={n1},                        % id's of article notes, if any
%   email={emanuel.heitlinger@hu-berlin.de}   % email address
%]{\inits{EH}\fnm{Emanuel} \snm{Heitlinger}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter the authors' addresses here        %%
%%                                          %%
%% Repeat \address commands as much as      %%
%% required.                                %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\address[id=aff1]{%                           % unique id
%  \orgname{Department of Zoology, Cambridge}, % university, etc
%  \street{Waterloo Road},                     %
  %\postcode{}                                % post or zip code
%  \city{London},                              % city
%  \cny{UK}                                    % country
%}
%\address[id=affi1]{%
%  \orgname{Humboldt-Universitat zu Berlin, Department of Molecular Parasitology},
%  \street{Philipstr. 13, House 14},
%  \postcode{10115}
%  \city{Berlin},
%  \cny{Germany}
%}
%\address[id=affi2]{%  
%  \orgname{Max Planck Institute for Biology of Ageing},
%  \street{Joseph-Stelzmann-Strasse 9B},
%  \postcode{50931}
%  \city{Cologne},
%  \cny{Germany}
%}
%\address[id=affi3]{%  
%  \orgname{Department Reproduction Biology and Evolutionary Ecology, Leibniz Institute of Zoo and Wildlife Research},
%  \street{Alfred-Kowalke-Str. 17},
%  \postcode{10315}
%  \city{Berlin},
%  \cny{Germany}
%}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter short notes here                   %%
%%                                          %%
%% Short notes will be after addresses      %%
%% on first page.                           %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{artnotes}
%\note{Sample of title note}     % note to the article
%\note[id=n1]{Equal contributor} % note, connected to author
\end{artnotes}

\end{fmbox}% comment this for two column layout

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% The Abstract begins here                 %%
%%                                          %%
%% Please refer to the Instructions for     %%
%% authors on http://www.biomedcentral.com  %%
%% and include the section headings         %%
%% accordingly for your article type.       %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{abstractbox}

\begin{abstract} % abstract
\parttitle{First part title} %if any
Text for this section.

\parttitle{Second part title} %if any
Text for this section.
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% The keywords begin here                  %%
%%                                          %%
%% Put each keyword in separate \kwd{}.     %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{keyword}
\kwd{Parasite, apicomplexa, RNA-seq, transcriptome, life-cycle, interaction}
\kwd{article}
\kwd{author}
\end{keyword}

% MSC classifications codes, if any
%\begin{keyword}[class=AMS]
%\kwd[Primary ]{}
%\kwd{}
%\kwd[; secondary ]{}
%\end{keyword}

\end{abstractbox}
%
%\end{fmbox}% uncomment this for twcolumn layout

\end{frontmatter}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% The Main Body begins here                %%
%%                                          %%
%% Please refer to the instructions for     %%
%% authors on:                              %%
%% http://www.biomedcentral.com/info/authors%%
%% and include the section headings         %%
%% accordingly for your article type.       %%
%%                                          %%
%% See the Results and Discussion section   %%
%% for details on how to create sub-sections%%
%%                                          %%
%% use \cite{...} to cite references        %%
%%  \cite{koon} and                         %%
%%  \cite{oreg,khar,zvai,xjon,schn,pond}    %%
%%  \nocite{smith,marg,hunn,advi,koha,mouse}%%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%% start of article main body
% <put your article body there>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%	INTRODUCTION
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section*{Introduction}
Text and results for this section, as per the individual journal's instructions for authors. 
%\cite{koon,oreg,khar,zvai,xjon,schn,pond,smith,marg,hunn,advi,koha,mouse}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%	RESULTS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

  \section*{Results}
Text for this section \ldots
\subsection*{GO term enrichments in heatmap clusters}
The annotations referred to here are in many 
cases inferred from the existence of predicted protein domains and have in some cases been 
verified in e.g. \textit{E. tenella} or other \textit{Eimeria} spp.

\subsubsection*{Preparation for invasion in oocysts}
The mRNA profile in the oocyst stage is mainly determined by highly abundant genes in cluster 4.
Overrepresented GO-term in this cluster are largely determined by genes orthologous to 
peptidases, microneme localized proteins reported to be involved in invasion, 
genes associated with clotting and in protozoans, adhesion, and genes that are annotated 
to be involved in amino acid biosynthesis.
  Aminopeptidase N ('related' annotation) is the annotation for orthologs of three genes with 
abundant mRNAs in oocysts. In humans, this enzyme has been reported to cleave peptides bound to 
major histocompatibility complex, MHC, II (UniProt reference if we want to keep this... but does any
secretion happen from oocysts...? Or is this too far-fetched to be interesting?).

Three genes can be linked to adhesion and through localization the same genes
also are asociated with invasion in the literature. 
A protein domain which is associated with adhesion in protozoans, thrombospondin type 1 
domain-containing protein, was found in our oocyst cluster (cluster 1). 
Thrombospondin type 1 domains have been reported in \textit{E. tenella} microneme localizing 
proteins, MIC. Tomley et al. (Tomley01) describe MIC4 as a protein contatining thrombospondin 
type 1 repeats. In \textit{E. tenella} MIC4 is expressed in sporozoites where it localizes to the 
apical end, in late schizonts and late oocyst stages, when sporozoites are forming.
(Tomley01)
In the protozoan \textit{P. falciparum}, another domain occuring in our data, the sushi domain, 
is reported in the apical sushi protein, ASP. In \textit{P. falciparum} it localizes to the 
micronemes in merozoites but not other stages (OKeeffe05). To our knowledge, the role of sushi 
domain proteins in \textit{Eimeria} spp. oocysts has not yet been investigated. 
  Other protein domains detected via orthologs to mRNAs expressed in oocysts in our data are
Limulus clotting factor C, Coch-5b2 (Cochlin) and Lgl1, LCCL, also known as F5/8 domain. 
LCCL domains have  been associated with discoidin lectin domains, which in turn are involved 
in cell adhesion (Pfam entries for 'LCCL domain' and 'Discoidin domain', May 2016). 
In the slime mold \textit{Dictyostelium discoideum)} the LCCL domain
is part of the mold'd discoidin adhesion protein. In humans, the LCCL domain in annotated as a 
coagulation factor domain. (Pfam entry for 'discoidin domain', May 2016)

\subsubsection*{Amino acid biosynthesis in oocysts}
mRNA of a aminotransferases indicate amino acid biosynthesis in oocysts. 
We also identify D-3-phosphoglycerate dehydrogenase and alanine dehydrogenase, which are enzymes 
contributing to L-serine and L-alanine production, respectively. A putative \textit{Eimeria} spp.
cystathionine beta-synthase, CBS, in this cluster also indicates de novo cysteine production. 
Alkyl sulfatase mRNA is another gene which contributes to overrepresentation of a GO-term in 
mRNAs highly abundant in oocysts. Generally, this enzyme enables an organism to exploit organic 
sulfur to produce and incorporate inorganic sulfur into the amino acids cysteine and methionine, 
when no inorganoc sulfur is available.

'Embryonic development'
Nicalin 1, patched family protein (hedgehog)

Propionate is one of two most abundant small-chain fatty acids 
in the gut along with butyrate. Both fatty acids are largely produced as degradation products from food
by commensal bacteria (Sun13). MmgE/PrpD is important for propionate catabolism in the 
2-methylcitric acid cycle and has been shown to be used by the intestinal intracellular bacterium 
\textit{Salmonella typhimurium} to generate pyruvate (Horswill99).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%% DAY xx   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection*{cluster 3..........}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%% DAY 7  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection*{Motility-related mRNAs indicate gamete development on day 7}
Two clusters contain genes with mRNAs highly abundant on day 7 p.i; cluster 1 and 2.
Dynein, kinesin and tubulin are annotations highly represented among orthologs of genes 
in both these clusters. The annotations indicate an important role for motility at this
timepoint, probably reflecting development of microgametes.
In addition, in cluster 2, there are two 'EF-hand domain containing proteins' annotations as well 
as caltractin, centrin-1, and troponin annotations. Caltractin and 
centrin-1 are associated with the centrosome and structure and function of microtubuli in mammals, 
and troponin is linked to muscle function (UniProt). 
Also potentially linked to motility is the occurence of growth arrest specific protein 8,
Gas8, which in the mouse has been reported to be highly 
expressed in the testes and important for mouse sperm function (Yeh02).

Other genes among the 38 indicate carbon fixation (glycolysis/gluconeogenesis) or 
conversions of nucleoside phosphates. In addition, a Ras family protein, RNA 
polymerase II transcription initiation factor and Sec23 and Sec24 were among orthologs 
identified in \textit{E. falciformis} cluster 2.

In cluster 1, carbon metabolism genes are represented by 6-phosphogluconate dehydrogenase 
and glycogen phosphorylase family protein 1. UDP-glucose 4-epimerase and amiloride-sensitive 
amine oxidase are reported as upregulated in gametocytes in \textit{E. tenella} by 
RNA-seq (Walker15) and suggested by those authors to play a role in cyst wall synthesis.

\subsubsection*{Microneme proteins highly expressed on day 7 p.i.}
Unintuitively for a protozoan organism, seven out of eight GO biological process terms
in cluster 1 are associated with wound healing and blood coagulation.
An explanation is offered by some of the orthologs to the three \textit{E. falciformis} genes
responsible for these terms. In protozoa, e.g., other \textit{Eimeria} spp. and 
\textit{Toxoplasma gondii} orthologs are annotated as 'Micronemal protein MIC4,
related' (E. tenella) and more generally for several other protozoa, 'PAN domain containing proteins'.
The PAN domain is found in the plasminogen/hepatocyte growth factor family and in coagulation 
factor XI family (REF), explaining why terms related to blood coagulation are enriched by these genes.
Later publications on \textit{T. gondii} (Marchant12) also associate PAN domains and proteins in 
apicomplexan parasites with micronemes and therefore invasion. In our case, this is peculiar, since 
the enrichment appears on day 7 p.i.. A possible role at this timepoint is suggested by work on 
the fungi \textit{Sclerotinia sclerotiorum} where Yu et al. reported an important role for PAN domain
proteins in cell wall integrity (Yu12). This role for MIC proteins has to our knowledge 
not been investigated in apicomplexan parasites. 
The PAN domain domain has also been reported to be common in nematodes such as 
\textit{Caenorhabditis elegans}, however the function is not understood. (Thordai99)
The other two GO terms in the cluster of day seven upregulated genes are DNA replication and DNA 
replication initiation, which most likely reflects late stage schizogony or gamete formation.
Six genes contribute to this enrichment and orthologs are either annotated as DNA replcation
licencing factors, DNA polymerases or minichromosome maintenance proteins 2/3/5/7, Mcm2/3/5/7.

\subsubsection*{Gene and sample patterns by hierarchical clustering}
Samples (columns) cluster into two major clusters where day 7 p.i. samples form one group distinct from
other samples. In the second group, oocsts and sporozoites are distinct and sporozoites cluster
most closely with day 3 and 5 p.i. samples. Day 3 and 5 p.i. samples also cluster into two groups,
of which one contains all NMRI day 5 p.i. samples. Apart from this, the two day 3 and 5 p.i. sample
clusters have no obvious patters.

For gene clusters (rows), the two groups with high mRNA abundance on day 7 p.i. (cluster 1 and 2)
do not cluster most closely with each other, but with the cluster for high mRNA abundance in oocysts
(cluster 1 association) and with the cluster for high mRNA abundance in sporozoites (cluster 2 
association).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%	DISCUSSION
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Discussion}
In our analysis we demonstrate which biological processes are dominant in different
life cycle stages of E. falciformis in the mouse. The RNAseq transcriptome provided here 
allows for detailed analysis of genes involved in those processes, providing candidates for 
life stage specific marker in \textit{Eimeria} spp. research.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%	METHODS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Methods}
\subsection*{Mice and infection procedure}
Three strains of mice were used in our experiments: NMRI (Charles River Laboratories, Sulzfeld,
Germany), C57BL/6 (), and Rag1-/- on C57BL/6 background (gift from Susanne Hartmann, FU?). 
Animal procedures were performed according to the German Animal Protection Laws as 
directed and approved by the overseeing authority Landesamt fuer Gesundheit und Soziales 
(Berlin, Germany). Animals where infected as described by Schmid et al., (Schmid12), but
tapwater was used instead of PBS for administration of oocysts. Briefly, the indicated number of
sporulated oocysts were purified by flotation from feces stored in potassium dichromate and 
administered orally in 100 uL tapwater. One \textit{E. falciformis} isolate
was used for all infections and parasite samples. Oocysts were originally purchased from Bayer 
() and the strain is maintained through passage in mice  in our facilities as described 
elsewhere (Schmid12).
\ldots
\subsection*{RNA extraction}
 \ldots
\subsection*{Sequence quality assessment and alignment}
Fastq\_quality\_filter was applied to Illumina Hiseq 2000 sequenced samples. 
Since this is not easily applicable to pair-end sequencing data, a low threshold was 
used on the hiseq data. A phred score of 10 was used, i.e., the probability of false base 
calling is one in ten. We further set q = 60, i.e., nine out of ten bases
or more is required to be correct in at least 
60\% of the bases in each read for the read sequence to be kept for further analysis.
This resulted in.......
\subsubsection*{Alignment and reference genomes}
We used the published \textit{Mus musculus} mm10 assembly (Genome Reference Consortium Mouse 
Build 38, GCA\_000001635.2) as reference genome including annotations for mouse data. The
\textit{E. falciformis} genome (Heitlinger14) was downloaded from ToxoDB (Gajria07). For the
alignment, the mouse and parasite genome files were merged into a dual reference genome, and 
files including mRNA sequences from both species were aligned against the dual reference genome
using TopHat2 (version 2.0.14, Trapnell09)/ Bowtie2 (version 1.1.2, Langmead12). Single-end and 
pair-end sequence samples were aligned separately with library type 'fr-unstranded' specified 
for pair-end samples. Import into R was enabled by the R package Ballgown, which requires bam 
files to be processed by Tablemaker (Frazee15). Tablemaker in turn makes use of Cufflinks 
(version 2.1.1, Trapnell10).

\ldots

\subsection*{Sub-heading for section}
Text for this sub-heading \ldots
\subsubsection*{Sub-sub heading for section}
Text for this sub-sub-heading \ldots
\paragraph*{Sub-sub-sub heading for section}
Text for this sub-sub-sub-heading \ldots
text text text
text text text
text text text
text text text
text text text
text text text
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Backmatter begins here                   %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{backmatter}

\section*{Competing interests}
  The authors declare that they have no competing interests.

\section*{Author's contributions}
    Text for this section \ldots

\section*{Acknowledgements}
  Text for this section \ldots
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                  The Bibliography                       %%
%%                                                         %%
%%  Bmc_mathpys.bst  will be used to                       %%
%%  create a .BBL file for submission.                     %%
%%  After submission of the .TEX file,                     %%
%%  you will be prompted to submit your .BBL file.         %%
%%                                                         %%
%%                                                         %%
%%  Note that the displayed Bibliography will not          %%
%%  necessarily be rendered by Latex exactly as specified  %%
%%  in the online Instructions for Authors.                %%
%%                                                         %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% if your bibliography is in bibtex format, use those commands:
\bibliographystyle{bmc-mathphys} % Style BST file (bmc-mathphys, vancouver, spbasic).
\bibliography{bmc_article}      % Bibliography file (usually '*.bib' )
% for author-year bibliography (bmc-mathphys or spbasic)
% a) write to bib file (bmc-mathphys only)
% @settings{label, options="nameyear"}
% b) uncomment next line
%\nocite{label}

% or include bibliography directly:
% \begin{thebibliography}
% \bibitem{b1}
% \end{thebibliography}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                               %%
%% Figures                       %%
%%                               %%
%% NB: this is for captions and  %%
%% Titles. All graphics must be  %%
%% submitted separately and NOT  %%
%% included in the Tex document  %%
%%                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%
%% Do not use \listoffigures as most will included as separate files

%\begin{figure}
 %\begin{picture}
 % \includegraphics[width=0.5\linewidth]{ef-heatmap}
 %\end{picture}
%\end{figure}

\section*{Figures}
\begin{figure}[h!]
  \includegraphics[width=0.5\linewidth]{efmm-ratio}
  \caption{\csentence{Parasite read sequences per mouse read sequences for each experimental condition.}
  The more advanced the infection is, the higher the ratio of parasite reads is, reflecting parasite replication between day zero and day seven. ......compare 1st and 2nd ....... patterns between Rag and NMRI/C57BL/6?  ........ Values are mean of replicates (n=2, if * n=1) on log10 scale. Each sample (replicate) consists of mRNA from three different mice.}
  \end{figure}

  \begin{figure}[h!]
  \includegraphics[width=0.7\linewidth]{ef-heatmap}
  \caption{\csentence{Parasite genes with different mRNA abundance between samples, sorted into five clusters by hierarchical clustering (method = complete, distance = euclidean).}
      \textit{E. falciformis} samples from seven days p.i. cluster together (NMRI mice only). These samples have distinct mRNA abundance patterns in all gene clusters, although more pronounced in clusters 1 (up) and 4 (down). Distinct groups of genes also define sporozoites (cluster 5, up) and oocysts (cluster 2, up). mRNA profiles on days three and five p.i. from all three mouse strains cluster together. These samples are distinct from oocysts, NMRI day 7 p.i., and sporozoites, however closest to the latter. On scale bar, 0 is mean mRNA abundance for each gene. Up and downregulation is standard deviations from mean.}
      \end{figure}

\begin{figure}[h!]
  \includegraphics[width=0.7\linewidth]{mm-heatmap}
  \caption{\csentence{Mouse genes with different mRNA abundance between samples, sorted into four clusters by hierarchical clustering (method = complete, distance = euclidean).}
Three out of four samples from day 7 p.i. (NMRI only) cluster together. These samples are characterized by genes in cluster 2 (up), 3, and 4 (down). NMRI and C57BL/6 uninfected samples cluster together (left), defined by clusters 3, 4 (up), and 1 (down). Rag1-/- samples cluster together. These samples share a weak downregulation of most genes in cluster 2 as well as upregulation of a small group genes in the same cluster. Uninfected Rag1-/- samples are separated from infected ones with distinct profiles in clusters 1, 3, and 4. On scale bar, 0 is mean mRNA abundance for each gene. Up and downregulation is standard deviations from mean.}
      \end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                               %%
%% Tables                        %%
%%                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% Use of \listoftables is discouraged.
%%
\section*{Tables}
\begin{table}[h!]
\caption{Sample table title. This is where the description of the table should go.}
      \begin{tabular}{cccc}
        \hline
           & B1  &B2   & B3\\ \hline
        A1 & 0.1 & 0.2 & 0.3\\
        A2 & ... & ..  & .\\
        A3 & ..  & .   & .\\ \hline
      \end{tabular}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                               %%
%% Additional Files              %%
%%                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section*{Additional Files}
  \subsection*{Additional file 1 --- Sample additional file title}
    Additional file descriptions text (including details of how to
    view the file, if it is in a non-standard format or the file extension).  This might
    refer to a multi-page table or a figure.

  \subsection*{Additional file 2 --- Sample additional file title}
    Additional file descriptions text.

\end{backmatter}
\end{document}