### 2016-07-14 figuresANDmanuscript/table_rc.tex

% latex table generated in R 3.3.0 by xtable 1.8-2 package
% Wed Jun 22 17:40:30 2016
\begin{table}[ht]
\centering
\begin{tabular}{lllll}
  \hline
Sample & Mouse transcripts & E. falciformis transcripts & Percentage E. falciformis & \#E. falciformis genes \\ 
  \hline
NMRI\_oocysts\_rep1 & 11676.000 & 108477484.000 & 99.989 & 5734.000 \\ 
  NMRI\_oocysts\_rep2 & 19024.000 & 126543533.000 & 99.985 & 5774.000 \\ 
  NMRI\_sporozoites\_rep1 & 13800.000 & 92259539.000 & 99.985 & 5808.000 \\ 
  NMRI\_sporozoites\_rep2 & 8702.000 & 21508353.000 & 99.960 & 5564.000 \\ 
  NMRI\_1stInf\_7dpi\_rep1 & 12532238.000 & 79648900.000 & 86.405 & 5894.000 \\ 
  NMRI\_1stInf\_7dpi\_rep2 & 94310278.000 & 154343046.000 & 62.072 & 5897.000 \\ 
  NMRI\_1stInf\_5dpi\_rep3 & 334671421.000 & 54022504.000 & 13.899 & 5794.000 \\ 
  NMRI\_2ndInf\_7dpi\_rep1 & 97221189.000 & 9927803.000 & 9.265 & 5865.000 \\ 
  NMRI\_1stInf\_5dpi\_rep1 & 204647381.000 & 18549727.000 & 8.311 & 5739.000 \\ 
  C57BL6\_1stInf\_5dpi\_rep2 & 32887009.000 & 250954.000 & 0.757 & 3946.000 \\ 
  NMRI\_2ndInf\_5dpi\_rep1 & 589643389.000 & 3752923.000 & 0.632 & 5602.000 \\ 
  NMRI\_1stInf\_5dpi\_rep2 & 316721928.000 & 1609009.000 & 0.505 & 5439.000 \\ 
  C57BL6\_2ndInf\_5dpi\_rep1 & 62053975.000 & 311900.000 & 0.500 & 4610.000 \\ 
  NMRI\_1stInf\_3dpi\_rep1 & 209723287.000 & 815820.000 & 0.388 & 5466.000 \\ 
  C57BL6\_1stInf\_5dpi\_rep1 & 65523435.000 & 217606.000 & 0.331 & 4259.000 \\ 
  Rag\_2ndInf\_5dpi\_rep1 & 85288323.000 & 224273.000 & 0.262 & 4251.000 \\ 
  Rag\_1stInf\_5dpi\_rep2 & 57192380.000 & 113673.000 & 0.198 & 2969.000 \\ 
  NMRI\_1stInf\_3dpi\_rep2 & 515516785.000 & 903330.000 & 0.175 & 5101.000 \\ 
  Rag\_1stInf\_5dpi\_rep1 & 71173382.000 & 73445.000 & 0.103 & 2748.000 \\ 
  NMRI\_2ndInf\_7dpi\_rep2 & 122999109.000 & 46700.000 & 0.038 & 2174.000 \\ 
  NMRI\_2ndInf\_3dpi\_rep2 & 106446839.000 & 34699.000 & 0.033 & 1901.000 \\ 
  NMRI\_1stInf\_0dpi\_rep1 & 229165002.000 & 31459.000 & 0.014 & 1380.000 \\ 
  NMRI\_2ndInf\_5dpi\_rep2 & 113957667.000 & 12692.000 & 0.011 & 539.000 \\ 
  NMRI\_2ndInf\_3dpi\_rep1 & 91352242.000 & 5355.000 & 0.006 & 121.000 \\ 
  NMRI\_2ndInf\_0dpi\_rep2 & 90681785.000 & 3738.000 & 0.004 & 62.000 \\ 
  Rag\_0dpi\_rep1 & 43414004.000 & 474.000 & 0.001 & 2.000 \\ 
  C57BL6\_0dpi\_rep1 & 53877840.000 & 491.000 & 0.001 & 2.000 \\ 
  C57BL6\_0dpi\_rep2 & 76753491.000 & 657.000 & 0.001 & 2.000 \\ 
  Rag\_0dpi\_rep2 & 80702547.000 & 730.000 & 0.001 & 3.000 \\ 
  NMRI\_2ndInf\_0dpi\_rep1 & 285032128.000 & 326.000 & 0.000 & 2.000 \\ 
   \hline
\end{tabular}
\end{table}