### 2016-08-12 tables/Table1_ReadCounts.tex
Streamlined pData and RC-summary table -> Table1

% latex table generated in R 3.3.1 by xtable 1.8-2 package
% Fri Aug 12 11:39:34 2016
\begin{table}[ht]
\centering
\begin{tabular}{llllllll}
  \hline
Sample & Sequencing method & batch & total reads & reads mapping Mouse & reads mapping E. falciformis & Percentage E. falciformis & \# E. falciformis genes \\ 
  \hline
NMRI\_2ndInf\_0dpi\_rep1 & GAII & 2 & 108,937,797 & 70,489,674 & 247 & 0.0004 & 1 \\ 
  Rag\_1stInf\_0dpi\_rep1 & hiseq & 3 & 25,362,793 & 18,853,850 & 443 & 0.0023 & 2 \\ 
  C57BL6\_1stInf\_0dpi\_rep1 & hiseq & 3 & 35,731,249 & 25,119,348 & 457 & 0.0018 & 2 \\ 
  C57BL6\_1stInf\_0dpi\_rep2 & hiseq & 3 & 47,085,959 & 34,377,133 & 608 & 0.0018 & 2 \\ 
  Rag\_1stInf\_0dpi\_rep2 & hiseq & 3 & 46,556,156 & 35,233,327 & 676 & 0.0019 & 2 \\ 
  NMRI\_2ndInf\_0dpi\_rep2 & hiseq & 3 & 58,122,244 & 40,794,245 & 3,406 & 0.0083 & 51 \\ 
  NMRI\_2ndInf\_3dpi\_rep1 & hiseq & 3 & 57,934,016 & 40,544,287 & 4,803 & 0.0118 & 95 \\ 
  NMRI\_2ndInf\_5dpi\_rep2 & hiseq & 3 & 63,965,539 & 48,289,181 & 10,941 & 0.0227 & 407 \\ 
  NMRI\_1stInf\_0dpi\_rep1 & GAII & 1 & 82,364,585 & 55,176,243 & 17,954 & 0.0325 & 701 \\ 
  NMRI\_2ndInf\_3dpi\_rep2 & hiseq & 3 & 65,548,826 & 46,171,909 & 29,548 & 0.0640 & 1,580 \\ 
  NMRI\_2ndInf\_7dpi\_rep2 & hiseq & 3 & 67,487,466 & 51,722,265 & 40,091 & 0.0775 & 1,836 \\ 
  Rag\_1stInf\_5dpi\_rep1 & hiseq & 3 & 38,651,359 & 29,982,453 & 63,024 & 0.2098 & 2,548 \\ 
  Rag\_1stInf\_5dpi\_rep2 & hiseq & 3 & 34,779,832 & 25,297,803 & 99,000 & 0.3898 & 2,828 \\ 
  C57BL6\_1stInf\_5dpi\_rep1 & hiseq & 3 & 40,904,388 & 29,319,604 & 185,969 & 0.6303 & 4,173 \\ 
  Rag\_2ndInf\_5dpi\_rep1 & hiseq & 3 & 50,049,848 & 37,093,621 & 192,856 & 0.5172 & 4,167 \\ 
  C57BL6\_1stInf\_5dpi\_rep2 & hiseq & 3 & 29,511,368 & 18,062,349 & 215,696 & 1.1801 & 3,823 \\ 
  C57BL6\_2ndInf\_5dpi\_rep1 & hiseq & 3 & 35,148,432 & 25,660,184 & 262,909 & 1.0142 & 4,563 \\ 
  NMRI\_1stInf\_3dpi\_rep1 & GAII & 1 & 73,236,430 & 49,993,358 & 394,384 & 0.7827 & 5,220 \\ 
  NMRI\_1stInf\_3dpi\_rep2 & GAII & 2 & 160,709,694 & 117,791,044 & 413,051 & 0.3494 & 4,862 \\ 
  NMRI\_1stInf\_5dpi\_rep2 & GAII & 2 & 119,902,722 & 76,419,774 & 794,570 & 1.0290 & 5,333 \\ 
  NMRI\_2ndInf\_5dpi\_rep1 & GAII & 2 & 230,773,955 & 143,186,486 & 1,846,840 & 1.2734 & 5,533 \\ 
  NMRI\_2ndInf\_7dpi\_rep1 & hiseq & 3 & 70,366,762 & 41,467,146 & 8,634,201 & 17.2335 & 5,875 \\ 
  NMRI\_1stInf\_5dpi\_rep1 & GAII & 2 & 76,702,168 & 47,037,087 & 8,669,701 & 15.5631 & 5,700 \\ 
  NMRI\_sporozoites\_rep2 & GAII & 0 & 19,551,681 & 8,656 & 11,470,604 & 99.9246 & 5,513 \\ 
  NMRI\_1stInf\_5dpi\_rep3 & GAII & 0 & 191,099,180 & 83,735,624 & 27,839,458 & 24.9513 & 5,784 \\ 
  NMRI\_1stInf\_7dpi\_rep1 & GAII & 1 & 66,505,514 & 3,310,666 & 39,400,884 & 92.2488 & 5,932 \\ 
  NMRI\_sporozoites\_rep1 & GAII & 1 & 67,325,397 & 4,334 & 43,774,401 & 99.9901 & 5,825 \\ 
  NMRI\_oocysts\_rep1 & GAII & 1 & 68,859,802 & 3,805 & 49,653,065 & 99.9923 & 5,695 \\ 
  NMRI\_oocysts\_rep2 & GAII & 0 & 151,090,783 & 18,524 & 71,019,860 & 99.9739 & 5,777 \\ 
  NMRI\_1stInf\_7dpi\_rep2 & GAII & 1 & 139,749,046 & 21,699,324 & 73,539,445 & 77.2159 & 5,943 \\ 
   \hline
\end{tabular}
\end{table}