### 2016-08-17 Supplement/tex/cluster3_BP_Mm_allInfDown.tex

% latex table generated in R 3.3.1 by xtable 1.8-2 package
% Mon Aug 15 15:27:09 2016
\begin{table}[ht]
\centering
\begin{tabular}{llrrrrr}
  \hline
GO.ID & Term & Annotated & Significant & Expected & p.value & adj.p \\ 
  \hline
GO:0021527 & spinal cord association neuron different... &   2 &   2 & 0.04 & 0.00 & 0.02 \\ 
  GO:0071280 & cellular response to copper ion &   3 &   2 & 0.06 & 0.00 & 0.02 \\ 
  GO:0046854 & phosphatidylinositol phosphorylation &  25 &   4 & 0.50 & 0.00 & 0.02 \\ 
  GO:0061512 & protein localization to cilium &  25 &   4 & 0.50 & 0.00 & 0.02 \\ 
  GO:0030258 & lipid modification & 140 &   9 & 2.77 & 0.00 & 0.02 \\ 
  GO:0021516 & dorsal spinal cord development &   4 &   2 & 0.08 & 0.00 & 0.02 \\ 
  GO:0060830 & ciliary receptor clustering involved in ... &   4 &   2 & 0.08 & 0.00 & 0.02 \\ 
  GO:0060971 & embryonic heart tube left/right pattern ... &   4 &   2 & 0.08 & 0.00 & 0.02 \\ 
  GO:0046834 & lipid phosphorylation &  29 &   4 & 0.57 & 0.00 & 0.02 \\ 
  GO:0015936 & coenzyme A metabolic process &  16 &   3 & 0.32 & 0.00 & 0.02 \\ 
  GO:0021522 & spinal cord motor neuron differentiation &  16 &   3 & 0.32 & 0.00 & 0.02 \\ 
  GO:0019087 & transformation of host cell by virus &   5 &   2 & 0.10 & 0.00 & 0.02 \\ 
  GO:0032484 & Ral protein signal transduction &   5 &   2 & 0.10 & 0.00 & 0.02 \\ 
  GO:0048842 & positive regulation of axon extension in... &   5 &   2 & 0.10 & 0.00 & 0.02 \\ 
  GO:1902669 & positive regulation of axon guidance &   5 &   2 & 0.10 & 0.00 & 0.02 \\ 
  GO:0046486 & glycerolipid metabolic process & 218 &  11 & 4.32 & 0.00 & 0.02 \\ 
  GO:0007033 & vacuole organization & 130 &   8 & 2.58 & 0.00 & 0.02 \\ 
  GO:0051188 & cofactor biosynthetic process & 104 &   7 & 2.06 & 0.00 & 0.02 \\ 
  GO:0042073 & intraciliary transport &  35 &   4 & 0.69 & 0.00 & 0.02 \\ 
  GO:0044088 & regulation of vacuole organization &  35 &   4 & 0.69 & 0.00 & 0.02 \\ 
  GO:0098840 & protein transport along microtubule &  35 &   4 & 0.69 & 0.00 & 0.02 \\ 
  GO:0021514 & ventral spinal cord interneuron differen... &   6 &   2 & 0.12 & 0.01 & 0.02 \\ 
  GO:0021515 & cell differentiation in spinal cord &  19 &   3 & 0.38 & 0.01 & 0.02 \\ 
  GO:0021517 & ventral spinal cord development &  19 &   3 & 0.38 & 0.01 & 0.02 \\ 
  GO:0009410 & response to xenobiotic stimulus &  37 &   4 & 0.73 & 0.01 & 0.02 \\ 
  GO:0033865 & nucleoside bisphosphate metabolic proces... &  20 &   3 & 0.40 & 0.01 & 0.02 \\ 
  GO:0033875 & ribonucleoside bisphosphate metabolic pr... &  20 &   3 & 0.40 & 0.01 & 0.02 \\ 
  GO:0034032 & purine nucleoside bisphosphate metabolic... &  20 &   3 & 0.40 & 0.01 & 0.02 \\ 
  GO:0051186 & cofactor metabolic process & 269 &  12 & 5.33 & 0.01 & 0.02 \\ 
  GO:0044255 & cellular lipid metabolic process & 633 &  22 & 12.54 & 0.01 & 0.02 \\ 
  GO:0006729 & tetrahydrobiopterin biosynthetic process &   7 &   2 & 0.14 & 0.01 & 0.02 \\ 
  GO:0014049 & positive regulation of glutamate secreti... &   7 &   2 & 0.14 & 0.01 & 0.02 \\ 
  GO:0046146 & tetrahydrobiopterin metabolic process &   7 &   2 & 0.14 & 0.01 & 0.02 \\ 
  GO:0070050 & neuron cellular homeostasis &   7 &   2 & 0.14 & 0.01 & 0.02 \\ 
  GO:0072488 & ammonium transmembrane transport &   7 &   2 & 0.14 & 0.01 & 0.02 \\ 
  GO:0097094 & craniofacial suture morphogenesis &   7 &   2 & 0.14 & 0.01 & 0.02 \\ 
  GO:0032892 & positive regulation of organic acid tran... &  21 &   3 & 0.42 & 0.01 & 0.02 \\ 
  GO:0021510 & spinal cord development &  40 &   4 & 0.79 & 0.01 & 0.02 \\ 
  GO:1903793 & positive regulation of anion transport &  40 &   4 & 0.79 & 0.01 & 0.02 \\ 
  GO:0031503 & protein complex localization &  91 &   6 & 1.80 & 0.01 & 0.02 \\ 
   \hline
\end{tabular}
\end{table}