### 2016-08-17 Supplement/tex/cluster4_BP_Mm_CompInfDown.tex

% latex table generated in R 3.3.1 by xtable 1.8-2 package
% Mon Aug 15 15:28:15 2016
\begin{table}[ht]
\centering
\begin{tabular}{llrrrrr}
  \hline
GO.ID & Term & Annotated & Significant & Expected & p.value & adj.p \\ 
  \hline
GO:0070886 & positive regulation of calcineurin-NFAT ... &   5 &   3 & 0.11 & 0.00 & 0.01 \\ 
  GO:0048016 & inositol phosphate-mediated signaling &  23 &   5 & 0.52 & 0.00 & 0.01 \\ 
  GO:0050817 & coagulation & 120 &  10 & 2.72 & 0.00 & 0.01 \\ 
  GO:0032102 & negative regulation of response to exter... & 184 &  12 & 4.17 & 0.00 & 0.02 \\ 
  GO:0007596 & blood coagulation & 118 &   9 & 2.67 & 0.00 & 0.02 \\ 
  GO:0007154 & cell communication & 3168 &  94 & 71.77 & 0.00 & 0.02 \\ 
  GO:0007599 & hemostasis & 120 &   9 & 2.72 & 0.00 & 0.02 \\ 
  GO:0007006 & mitochondrial membrane organization &  76 &   7 & 1.72 & 0.00 & 0.02 \\ 
  GO:0023052 & signaling & 3091 &  92 & 70.02 & 0.00 & 0.02 \\ 
  GO:0070884 & regulation of calcineurin-NFAT signaling... &  11 &   3 & 0.25 & 0.00 & 0.02 \\ 
  GO:0006120 & mitochondrial electron transport, NADH t... &  12 &   3 & 0.27 & 0.00 & 0.02 \\ 
  GO:0043588 & skin development & 150 &  10 & 3.40 & 0.00 & 0.02 \\ 
  GO:1903035 & negative regulation of response to wound... & 103 &   8 & 2.33 & 0.00 & 0.02 \\ 
  GO:0044700 & single organism signaling & 3084 &  91 & 69.86 & 0.00 & 0.02 \\ 
  GO:0007033 & vacuole organization & 130 &   9 & 2.94 & 0.00 & 0.02 \\ 
  GO:0035898 & parathyroid hormone secretion &   4 &   2 & 0.09 & 0.00 & 0.02 \\ 
  GO:0060684 & epithelial-mesenchymal cell signaling &   4 &   2 & 0.09 & 0.00 & 0.02 \\ 
  GO:0050701 & interleukin-1 secretion &  27 &   4 & 0.61 & 0.00 & 0.02 \\ 
  GO:0006914 & autophagy & 358 &  17 & 8.11 & 0.00 & 0.02 \\ 
  GO:0006937 & regulation of muscle contraction &  67 &   6 & 1.52 & 0.00 & 0.02 \\ 
  GO:0030168 & platelet activation &  67 &   6 & 1.52 & 0.00 & 0.02 \\ 
  GO:1902188 & positive regulation of viral release fro... &  15 &   3 & 0.34 & 0.00 & 0.02 \\ 
  GO:0071496 & cellular response to external stimulus & 139 &   9 & 3.15 & 0.00 & 0.02 \\ 
  GO:0032905 & transforming growth factor beta1 product... &   5 &   2 & 0.11 & 0.00 & 0.02 \\ 
  GO:0032908 & regulation of transforming growth factor... &   5 &   2 & 0.11 & 0.00 & 0.02 \\ 
  GO:0045991 & carbon catabolite activation of transcri... &   5 &   2 & 0.11 & 0.00 & 0.02 \\ 
  GO:1900165 & negative regulation of interleukin-6 sec... &   5 &   2 & 0.11 & 0.00 & 0.02 \\ 
  GO:0032692 & negative regulation of interleukin-1 pro... &  16 &   3 & 0.36 & 0.01 & 0.02 \\ 
  GO:0050878 & regulation of body fluid levels & 197 &  11 & 4.46 & 0.01 & 0.02 \\ 
  GO:0007165 & signal transduction & 2897 &  84 & 65.63 & 0.01 & 0.02 \\ 
  GO:0007176 & regulation of epidermal growth factor-ac... &  17 &   3 & 0.39 & 0.01 & 0.02 \\ 
  GO:0033173 & calcineurin-NFAT signaling cascade &  17 &   3 & 0.39 & 0.01 & 0.02 \\ 
  GO:0050663 & cytokine secretion & 121 &   8 & 2.74 & 0.01 & 0.02 \\ 
  GO:0019076 & viral release from host cell &  33 &   4 & 0.75 & 0.01 & 0.02 \\ 
  GO:0006936 & muscle contraction & 123 &   8 & 2.79 & 0.01 & 0.02 \\ 
  GO:0050728 & negative regulation of inflammatory resp... &  75 &   6 & 1.70 & 0.01 & 0.02 \\ 
  GO:0007220 & Notch receptor processing &   6 &   2 & 0.14 & 0.01 & 0.02 \\ 
  GO:0044828 & negative regulation by host of viral gen... &   6 &   2 & 0.14 & 0.01 & 0.02 \\ 
  GO:0045990 & carbon catabolite regulation of transcri... &   6 &   2 & 0.14 & 0.01 & 0.02 \\ 
  GO:0050711 & negative regulation of interleukin-1 sec... &   6 &   2 & 0.14 & 0.01 & 0.02 \\ 
  GO:0031670 & cellular response to nutrient &  18 &   3 & 0.41 & 0.01 & 0.02 \\ 
  GO:0050919 & negative chemotaxis &  18 &   3 & 0.41 & 0.01 & 0.02 \\ 
  GO:0032612 & interleukin-1 production &  54 &   5 & 1.22 & 0.01 & 0.02 \\ 
  GO:0050818 & regulation of coagulation &  54 &   5 & 1.22 & 0.01 & 0.02 \\ 
  GO:0060048 & cardiac muscle contraction &  54 &   5 & 1.22 & 0.01 & 0.02 \\ 
  GO:0042493 & response to drug & 151 &   9 & 3.42 & 0.01 & 0.02 \\ 
  GO:0031668 & cellular response to extracellular stimu... & 126 &   8 & 2.85 & 0.01 & 0.02 \\ 
  GO:0008053 & mitochondrial fusion &  19 &   3 & 0.43 & 0.01 & 0.02 \\ 
  GO:0044803 & multi-organism membrane organization &  19 &   3 & 0.43 & 0.01 & 0.02 \\ 
  GO:0050704 & regulation of interleukin-1 secretion &  19 &   3 & 0.43 & 0.01 & 0.02 \\ 
  GO:0050707 & regulation of cytokine secretion & 104 &   7 & 2.36 & 0.01 & 0.02 \\ 
  GO:0009991 & response to extracellular stimulus & 274 &  13 & 6.21 & 0.01 & 0.02 \\ 
  GO:0007618 & mating &  20 &   3 & 0.45 & 0.01 & 0.02 \\ 
  GO:0009612 & response to mechanical stimulus &  58 &   5 & 1.31 & 0.01 & 0.02 \\ 
  GO:0007620 & copulation &   7 &   2 & 0.16 & 0.01 & 0.02 \\ 
  GO:0044793 & negative regulation by host of viral pro... &   7 &   2 & 0.16 & 0.01 & 0.02 \\ 
  GO:1904467 & regulation of tumor necrosis factor secr... &   7 &   2 & 0.16 & 0.01 & 0.02 \\ 
  GO:1990774 & tumor necrosis factor secretion &   7 &   2 & 0.16 & 0.01 & 0.02 \\ 
   \hline
\end{tabular}
\end{table}