### 2016-08-17 Supplement/tex/cluster4_MF_Mm_CompInfDown.tex

% latex table generated in R 3.3.1 by xtable 1.8-2 package
% Mon Aug 15 15:27:20 2016
\begin{table}[ht]
\centering
\begin{tabular}{llrrrrr}
  \hline
GO.ID & Term & Annotated & Significant & Expected & p.value & adj.p \\ 
  \hline
GO:0019166 & trans-2-enoyl-CoA reductase (NADPH) acti... &   3 &   2 & 0.07 & 0.00 & 0.03 \\ 
  GO:0005102 & receptor binding & 889 &  34 & 20.21 & 0.00 & 0.03 \\ 
  GO:0047391 & alkylglycerophosphoethanolamine phosphod... &   4 &   2 & 0.09 & 0.00 & 0.03 \\ 
  GO:0004190 & aspartic-type endopeptidase activity &  15 &   3 & 0.34 & 0.00 & 0.03 \\ 
  GO:0031683 & G-protein beta/gamma-subunit complex bin... &  15 &   3 & 0.34 & 0.00 & 0.03 \\ 
  GO:0070001 & aspartic-type peptidase activity &  15 &   3 & 0.34 & 0.00 & 0.03 \\ 
  GO:0004197 & cysteine-type endopeptidase activity &  57 &   5 & 1.30 & 0.01 & 0.03 \\ 
  GO:0016922 & ligand-dependent nuclear receptor bindin... &  20 &   3 & 0.45 & 0.01 & 0.03 \\ 
   \hline
\end{tabular}
\end{table}