### 2016-08-17 Supplement/tex/cluster5_BP_Mm_7dpiUp.tex

% latex table generated in R 3.3.1 by xtable 1.8-2 package
% Mon Aug 15 15:25:59 2016
\begin{table}[ht]
\centering
\begin{tabular}{llrrrrr}
  \hline
GO.ID & Term & Annotated & Significant & Expected & p.value & adj.p \\ 
  \hline
GO:0002376 & immune system process & 1490 &  62 & 25.03 & 0.00 & 0.00 \\ 
  GO:0006955 & immune response & 775 &  42 & 13.02 & 0.00 & 0.00 \\ 
  GO:0002250 & adaptive immune response & 241 &  21 & 4.05 & 0.00 & 0.00 \\ 
  GO:0002682 & regulation of immune system process & 821 &  38 & 13.79 & 0.00 & 0.00 \\ 
  GO:0042981 & regulation of apoptotic process & 1026 &  43 & 17.24 & 0.00 & 0.00 \\ 
  GO:0043067 & regulation of programmed cell death & 1040 &  43 & 17.47 & 0.00 & 0.00 \\ 
  GO:0001775 & cell activation & 620 &  31 & 10.42 & 0.00 & 0.00 \\ 
  GO:0010941 & regulation of cell death & 1120 &  44 & 18.82 & 0.00 & 0.00 \\ 
  GO:0006952 & defense response & 823 &  36 & 13.83 & 0.00 & 0.00 \\ 
  GO:0050776 & regulation of immune response & 410 &  24 & 6.89 & 0.00 & 0.00 \\ 
  GO:0045321 & leukocyte activation & 543 &  28 & 9.12 & 0.00 & 0.00 \\ 
  GO:0045087 & innate immune response & 395 &  23 & 6.64 & 0.00 & 0.00 \\ 
  GO:0006915 & apoptotic process & 1301 &  47 & 21.86 & 0.00 & 0.00 \\ 
  GO:0002684 & positive regulation of immune system pro... & 531 &  27 & 8.92 & 0.00 & 0.00 \\ 
  GO:0012501 & programmed cell death & 1323 &  47 & 22.23 & 0.00 & 0.00 \\ 
  GO:0002443 & leukocyte mediated immunity & 216 &  16 & 3.63 & 0.00 & 0.00 \\ 
  GO:0002460 & adaptive immune response based on somati... & 166 &  14 & 2.79 & 0.00 & 0.00 \\ 
  GO:0006954 & inflammatory response & 369 &  21 & 6.20 & 0.00 & 0.00 \\ 
  GO:0001816 & cytokine production & 434 &  23 & 7.29 & 0.00 & 0.00 \\ 
  GO:0050896 & response to stimulus & 4563 & 110 & 76.67 & 0.00 & 0.00 \\ 
  GO:0043065 & positive regulation of apoptotic process & 439 &  23 & 7.38 & 0.00 & 0.00 \\ 
  GO:0009617 & response to bacterium & 283 &  18 & 4.75 & 0.00 & 0.00 \\ 
  GO:0043068 & positive regulation of programmed cell d... & 443 &  23 & 7.44 & 0.00 & 0.00 \\ 
  GO:0002819 & regulation of adaptive immune response & 106 &  11 & 1.78 & 0.00 & 0.00 \\ 
  GO:0001817 & regulation of cytokine production & 385 &  21 & 6.47 & 0.00 & 0.00 \\ 
  GO:0006950 & response to stress & 2248 &  65 & 37.77 & 0.00 & 0.00 \\ 
  GO:0008219 & cell death & 1414 &  47 & 23.76 & 0.00 & 0.00 \\ 
  GO:0002703 & regulation of leukocyte mediated immunit... & 135 &  12 & 2.27 & 0.00 & 0.00 \\ 
  GO:0016265 & death & 1418 &  47 & 23.82 & 0.00 & 0.00 \\ 
  GO:0002366 & leukocyte activation involved in immune ... & 163 &  13 & 2.74 & 0.00 & 0.00 \\ 
  GO:0002263 & cell activation involved in immune respo... & 165 &  13 & 2.77 & 0.00 & 0.00 \\ 
  GO:0002822 & regulation of adaptive immune response b... &  95 &  10 & 1.60 & 0.00 & 0.00 \\ 
  GO:0010942 & positive regulation of cell death & 472 &  23 & 7.93 & 0.00 & 0.00 \\ 
  GO:0016337 & single organismal cell-cell adhesion & 508 &  24 & 8.54 & 0.00 & 0.00 \\ 
  GO:0050778 & positive regulation of immune response & 318 &  18 & 5.34 & 0.00 & 0.00 \\ 
  GO:0002252 & immune effector process & 485 &  23 & 8.15 & 0.00 & 0.00 \\ 
  GO:0007159 & leukocyte cell-cell adhesion & 357 &  19 & 6.00 & 0.00 & 0.00 \\ 
  GO:0001819 & positive regulation of cytokine producti... & 263 &  16 & 4.42 & 0.00 & 0.00 \\ 
  GO:0046649 & lymphocyte activation & 467 &  22 & 7.85 & 0.00 & 0.00 \\ 
  GO:0098609 & cell-cell adhesion & 545 &  24 & 9.16 & 0.00 & 0.00 \\ 
  GO:0002449 & lymphocyte mediated immunity & 160 &  12 & 2.69 & 0.00 & 0.00 \\ 
  GO:0002285 & lymphocyte activation involved in immune... & 111 &  10 & 1.87 & 0.00 & 0.00 \\ 
  GO:0098602 & single organism cell adhesion & 551 &  24 & 9.26 & 0.00 & 0.00 \\ 
  GO:0002683 & negative regulation of immune system pro... & 278 &  16 & 4.67 & 0.00 & 0.00 \\ 
  GO:0032623 & interleukin-2 production &  50 &   7 & 0.84 & 0.00 & 0.00 \\ 
  GO:0034109 & homotypic cell-cell adhesion & 386 &  19 & 6.49 & 0.00 & 0.00 \\ 
  GO:0002694 & regulation of leukocyte activation & 324 &  17 & 5.44 & 0.00 & 0.00 \\ 
  GO:0002706 & regulation of lymphocyte mediated immuni... &  98 &   9 & 1.65 & 0.00 & 0.00 \\ 
  GO:0042110 & T cell activation & 330 &  17 & 5.54 & 0.00 & 0.00 \\ 
  GO:0070489 & T cell aggregation & 330 &  17 & 5.54 & 0.00 & 0.00 \\ 
  GO:0007259 & JAK-STAT cascade &  99 &   9 & 1.66 & 0.00 & 0.00 \\ 
  GO:0097696 & STAT cascade &  99 &   9 & 1.66 & 0.00 & 0.00 \\ 
  GO:0071593 & lymphocyte aggregation & 331 &  17 & 5.56 & 0.00 & 0.00 \\ 
  GO:0002286 & T cell activation involved in immune res... &  57 &   7 & 0.96 & 0.00 & 0.00 \\ 
  GO:0097285 & cell-type specific apoptotic process & 367 &  18 & 6.17 & 0.00 & 0.00 \\ 
  GO:0002274 & myeloid leukocyte activation & 126 &  10 & 2.12 & 0.00 & 0.00 \\ 
  GO:0070486 & leukocyte aggregation & 337 &  17 & 5.66 & 0.00 & 0.00 \\ 
  GO:0042267 & natural killer cell mediated cytotoxicit... &  41 &   6 & 0.69 & 0.00 & 0.00 \\ 
  GO:0002705 & positive regulation of leukocyte mediate... &  81 &   8 & 1.36 & 0.00 & 0.00 \\ 
  GO:0007165 & signal transduction & 2897 &  73 & 48.67 & 0.00 & 0.00 \\ 
  GO:0046425 & regulation of JAK-STAT cascade &  83 &   8 & 1.39 & 0.00 & 0.00 \\ 
  GO:1904892 & regulation of STAT cascade &  83 &   8 & 1.39 & 0.00 & 0.00 \\ 
  GO:0002228 & natural killer cell mediated immunity &  43 &   6 & 0.72 & 0.00 & 0.00 \\ 
  GO:0050865 & regulation of cell activation & 348 &  17 & 5.85 & 0.00 & 0.00 \\ 
  GO:0046427 & positive regulation of JAK-STAT cascade &  44 &   6 & 0.74 & 0.00 & 0.00 \\ 
  GO:1904894 & positive regulation of STAT cascade &  44 &   6 & 0.74 & 0.00 & 0.00 \\ 
  GO:0002237 & response to molecule of bacterial origin & 190 &  12 & 3.19 & 0.00 & 0.00 \\ 
  GO:0051716 & cellular response to stimulus & 3830 &  90 & 64.35 & 0.00 & 0.00 \\ 
  GO:0042327 & positive regulation of phosphorylation & 613 &  24 & 10.30 & 0.00 & 0.00 \\ 
  GO:0002697 & regulation of immune effector process & 288 &  15 & 4.84 & 0.00 & 0.00 \\ 
  GO:0007155 & cell adhesion & 826 &  29 & 13.88 & 0.00 & 0.00 \\ 
  GO:0001934 & positive regulation of protein phosphory... & 583 &  23 & 9.80 & 0.00 & 0.00 \\ 
  GO:0042088 & T-helper 1 type immune response &  30 &   5 & 0.50 & 0.00 & 0.00 \\ 
  GO:0042269 & regulation of natural killer cell mediat... &  30 &   5 & 0.50 & 0.00 & 0.00 \\ 
  GO:0048518 & positive regulation of biological proces... & 3592 &  85 & 60.35 & 0.00 & 0.00 \\ 
  GO:0022610 & biological adhesion & 835 &  29 & 14.03 & 0.00 & 0.00 \\ 
  GO:0002715 & regulation of natural killer cell mediat... &  31 &   5 & 0.52 & 0.00 & 0.00 \\ 
  GO:0010562 & positive regulation of phosphorus metabo... & 672 &  25 & 11.29 & 0.00 & 0.00 \\ 
  GO:0045937 & positive regulation of phosphate metabol... & 672 &  25 & 11.29 & 0.00 & 0.00 \\ 
  GO:0001909 & leukocyte mediated cytotoxicity &  70 &   7 & 1.18 & 0.00 & 0.00 \\ 
  GO:0002707 & negative regulation of lymphocyte mediat... &  32 &   5 & 0.54 & 0.00 & 0.00 \\ 
  GO:0001910 & regulation of leukocyte mediated cytotox... &  50 &   6 & 0.84 & 0.00 & 0.00 \\ 
  GO:0044700 & single organism signaling & 3084 &  75 & 51.82 & 0.00 & 0.00 \\ 
  GO:0019221 & cytokine-mediated signaling pathway & 206 &  12 & 3.46 & 0.00 & 0.00 \\ 
  GO:0043403 & skeletal muscle tissue regeneration &  18 &   4 & 0.30 & 0.00 & 0.00 \\ 
  GO:0023052 & signaling & 3091 &  75 & 51.93 & 0.00 & 0.00 \\ 
  GO:0050777 & negative regulation of immune response &  96 &   8 & 1.61 & 0.00 & 0.00 \\ 
  GO:0007166 & cell surface receptor signaling pathway & 1389 &  41 & 23.34 & 0.00 & 0.00 \\ 
  GO:0048584 & positive regulation of response to stimu... & 1207 &  37 & 20.28 & 0.00 & 0.00 \\ 
  GO:0009605 & response to external stimulus & 1212 &  37 & 20.36 & 0.00 & 0.00 \\ 
  GO:0043312 & neutrophil degranulation &   8 &   3 & 0.13 & 0.00 & 0.00 \\ 
  GO:0007154 & cell communication & 3168 &  76 & 53.23 & 0.00 & 0.00 \\ 
  GO:0032612 & interleukin-1 production &  54 &   6 & 0.91 & 0.00 & 0.00 \\ 
  GO:0009611 & response to wounding & 424 &  18 & 7.12 & 0.00 & 0.00 \\ 
  GO:0048583 & regulation of response to stimulus & 2241 &  58 & 37.65 & 0.00 & 0.00 \\ 
  GO:0001906 & cell killing &  77 &   7 & 1.29 & 0.00 & 0.00 \\ 
  GO:0002685 & regulation of leukocyte migration & 102 &   8 & 1.71 & 0.00 & 0.00 \\ 
  GO:0035556 & intracellular signal transduction & 1613 &  45 & 27.10 & 0.00 & 0.00 \\ 
  GO:0050707 & regulation of cytokine secretion & 104 &   8 & 1.75 & 0.00 & 0.00 \\ 
  GO:0002699 & positive regulation of immune effector p... & 131 &   9 & 2.20 & 0.00 & 0.00 \\ 
   \hline
\end{tabular}
\end{table}