### 2016-09-05 manuscript/bmc_article.tex
manuscript: changes in results (adjustments) and addition of heatmap introduction. GO-discussion not included

%% BioMed_Central_Tex_Template_v1.06
%%                                      %
%  bmc_article.tex            ver: 1.06 %
%                                       %

%%IMPORTANT: do not delete the first line of this template
%%It must be present to enable the BMC Submission system to
%%recognise this template!!

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                     %%
%%  LaTeX template for BioMed Central  %%
%%     journal article submissions     %%
%%                                     %%
%%          <8 June 2012>              %%
%%                                     %%
%%                                     %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                                                 %%
%% For instructions on how to fill out this Tex template           %%
%% document please refer to Readme.html and the instructions for   %%
%% authors page on the biomed central website                      %%
%% http://www.biomedcentral.com/info/authors/                      %%
%%                                                                 %%
%% Please do not use \input{...} to include other tex files.       %%
%% Submit your LaTeX manuscript as one .tex document.              %%
%%                                                                 %%
%% All additional figures and files should be attached             %%
%% separately and not embedded in the \TeX\ document itself.       %%
%%                                                                 %%
%% BioMed Central currently use the MikTex distribution of         %%
%% TeX for Windows) of TeX and LaTeX.  This is available from      %%
%% http://www.miktex.org                                           %%
%%                                                                 %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%% additional documentclass options:
%  [doublespacing]
%  [linenumbers]   - put the line numbers on margins

%%% loading packages, author definitions

%\documentclass[twocolumn]{bmcart}% uncomment this for twocolumn layout and comment line below
\documentclass{bmcart}

%%% Load packages
\usepackage{amsthm,amsmath}
%\usepackage{asmath}
\RequirePackage{natbib}
%\RequirePackage[authoryear]{natbib}% uncomment this for author-year bibliography
\RequirePackage{hyperref}
\usepackage[utf8]{inputenc} %unicode support
%\usepackage[applemac]{inputenc} %applemac support if unicode package fails
%\usepackage[latin1]{inputenc} %UNIX support if unicode package fails
%\usepackage{siunitx}
\usepackage[colorlinks]{hyperref} % To color blocks of text 
%\usepackage[dvipsnames]{xcolor} % More color options (see Wikibooks)  
\usepackage{changepage}
\usepackage{booktabs} % For prettier tables
\usepackage{float} % To control positions of e.g. tables or figures
%\usepackage[section]{placeins}
\usepackage{graphicx}
\usepackage[flushleft]{threeparttable}

% to color table rows
\usepackage{color, colortbl}
\graphicspath{ {../} }

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                             %%
%%  If you wish to display your graphics for   %%
%%  your own use using includegraphic or       %%
%%  includegraphics, then comment out the      %%
%%  following two lines of code.               %%
%%  NB: These line *must* be included when     %%
%%  submitting to BMC.                         %%
%%  All figure files must be submitted as      %%
%%  separate graphics through the BMC          %%
%%  submission process, not included in the    %%
%%  submitted article.                         %%
%%                                             %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\def\includegraphic{}
%\def\includegraphics{}

%%% Put your definitions there:
\startlocaldefs
\endlocaldefs

%%% Begin ...
\begin{document}

%%% Start of article front matter
\begin{frontmatter}

\begin{fmbox}
\dochead{Research}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter the title of your article here     %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{Dual host-parasite transcriptomes of apicomplexan Eimeria
  falciformis and its natural mouse host}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter the authors here                   %%
%%                                          %%
%% Specify information, if available,       %%
%% in the form:                             %%
%%   <key>={<id1>,<id2>}                    %%
%%   <key>=                                 %%
%% Comment or delete the keys which are     %%
%% not used. Repeat \author command as much %%
%% as required.                             %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\author[ addressref={aff1},] {\inits{TK}\fnm{Totta} \snm{Kasemo}}
\author[ addressref={aff1},] {\inits{SS}\fnm{Simone} \snm{Spork}}
\author[ addressref={aff2},] {\inits{CD}\fnm{Christoph} \snm{Dieterich}}
\author[ addressref={aff1}, ]{\inits{RL}\fnm{Richard} \snm{Lucius}}
\author[ addressref={aff1, aff3},              
  corref={aff1},                   
  email={emanuel.heitlinger@hu-berlin.de}]{\inits{EH}\fnm{Emanuel} \snm{Heitlinger}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter the authors' addresses here        %%
%%                                          %%
%% Repeat \address commands as much as      %%
%% required.                                %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\address[id=aff1]{
	\orgname{Institute of Biology, Humboldt-Universitat zu Berlin}, 
	\street{Philippstr. 13, Haus 14},
	\postcode{10115}                 
	\city{Berlin},                   
	\cny{Germany}                    
}
\address[id=aff2]{  
	\orgname{Max Planck Institute for Biology of Ageing},
	\street{Joseph-Stelzmann-Strasse 9B},
	\postcode{50931}
	\city{Cologne},
	\cny{Germany}
}
\address[id=aff3]{  
	\orgname{Leibniz Institute of Zoo and Wildlife Research},
	\street{Alfred-Kowalke-Str. 17},
	\postcode{10315}
	\city{Berlin},
	\cny{Germany}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter short notes here                   %%
%%                                          %%
%% Short notes will be after addresses      %%
%% on first page.                           %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{artnotes}
%\note{Sample of title note}     % note to the article
%\note[id=n1]{Equal contributor} % note, connected to author
%\note[id=n2]
\end{artnotes}

%\end{fmbox}% comment this for two column layout

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% The Abstract begins here                 %%
%%                                          %%
%% Please refer to the Instructions for     %%
%% authors on http://www.biomedcentral.com  %%
%% and include the section headings         %%
%% accordingly for your article type.       %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{abstractbox}

\begin{abstract} % abstract
\parttitle{First part title} %if any
Text for this section.

\parttitle{Second part title} %if any
Text for this section.
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% The keywords begin here                  %%
%%                                          %%
%% Put each keyword in separate \kwd{}.     %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{keyword}
\kwd{Parasite, apicomplexa, RNA-seq, transcriptome, life-cycle, interaction}
%\kwd{article}
%\kwd{author}
\end{keyword}

% MSC classifications codes, if any
%\begin{keyword}[class=AMS]
%\kwd[Primary ]{}
%\kwd{}
%\kwd[; secondary ]{}
%\end{keyword}

\end{abstractbox}
\end{fmbox}% uncomment this for twcolumn layout
\end{frontmatter}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% The Main Body begins here                %%
%%                                          %%
%% Please refer to the instructions for     %%
%% authors on:                              %%
%% http://www.biomedcentral.com/info/authors%%
%% and include the section headings         %%
%% accordingly for your article type.       %%
%%                                          %%
%% See the Results and Discussion section   %%
%% for details on how to create sub-sections%%
%%                                          %%
%% use \cite{...} to cite references        %%
%%  \cite{koon} and                         %%
%%  \cite{oreg,khar,zvai,xjon,schn,pond}    %%
%%  \nocite{smith,marg,hunn,advi,koha,mouse}%%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%% start of article main body
% <put your article body there>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%	INTRODUCTION
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section*{Introduction}
Text and results for this section, as per the individual journal's instructions for authors. 
%\cite{koon,oreg,khar,zvai,xjon,schn,pond,smith,marg,hunn,advi,koha,mouse}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%	RESULTS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

  \section*{Results}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% 	OVERVIEW READS PER SAMPLE, LAYOT AS EXPERIMENTAL OVERVIEW
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% 	OVERVIEW READS PER SAMPLE, LAYOT AS EXPERIMENTAL OVERVIEW
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

  \section*{A dual transcriptomics experiment}

We performed mRNA sequencing of caecum tissue from mice infected with
the apicomplexan parasite \textit{E. falciformis}. Oocysts and
sporozoites were included as ``environmental'' stages and processed
``\textit{in vitro}''. To follow the life cycle of the parasite, we
compare different time-points after infection.

We additionally used different mouse strains which display different
immunocompetence in infection trials measured by oocyst output to assess the influence of host
immunocompetence on parasite development (figure 1a,b). In these
experiments we analysed mice lacking the Recombination activating gene 1, Rag1, and compared them to the
parental C57BL/6 mouse line. Rag1 knock-out mice lack adaptive immune
response due to the absence of mature B and T lymphocytes.  All three
mouse strains were used for infections of both naive mice and previously
infected animals (onward referred to as "challenged").

Basic phenotyping showed differences in oocyst numbers
between the immunocompetent (C57BL/6) and
immunodeficient (Rag1-/-) host strain in infection of naive mice and 
in challenge infections. Immunocompetent NMRI mice were infected with a higher dosis  
of sporulated oocysts, and a drastic reduction of oocysts in faeces was seen in challenged mice
compared to naive mice (Figure 1a). Oocyst numbers in faeces peaked on days 8-9 and 
all mice had cleared the infection by day 14. Development of \textit{E. falciformis} intestinal stages is reflected by 
the percentage of parasite reads sequenced per time-point post infection (Figure 1b) and this was confirmed by quantitative reverse transcription PCR (RT qPCR) of parasitic 18S  (Figure 1c).

We thus use dual RNA-seq to analyse the life cycle of \textit{E. falciformis} under
the influence of host immune responses at early stages of infection. We used an experimental
design which allows to compare infections at 5 days post infection (dpi) for all experimental
conditions (NMRI, C57BL/6 and Rag1-/- mouse strains in naive and challenge infection). In addition,
additional time-points 3 dpi and 7 dpi were analysed from NMRI mice (Figure 1d).

\section*{Parasite and host dual transcriptomes can be assessed in parallel}
In this RNA-seq experiment, each replicate sample was enriched for caecum epithelial tissue which was pooled 
from three mice. mRNA was extracted and
sequencing libraries prepared. Two biological replicates were used for
all but two conditions (with one and three replicates, respectively). These libraries were sequenced on
several lanes of Illumina GAIIX (13 samples) and HiSeq machines (14 samples) and mapped to both
mouse and parasite genomes simultaneously to avoid spurious assignments
of reads in ultra conserved genomic regions. 
As samples and individual replicates were sequenced in batches to different depth and using different
instrumentation (Table 1) we performed quality controls (additional files xyz). These confirm
the absence of batch effects influencing analysis and quality of 
results.

Total numbers of sequenced reads as well as reads
mapped to either the \textit{E. falciformis} genome or the
mouse genome are indicated in Table 1 for all replicates. The number of total read mappings 
for individual replicates ranged from 11,479,260 (sample NMRI\_sporozoites\_rep2) to 145,033,326
(NMRI\_2ndInf\_5dpi\_rep1). At the latest time-point samples, 7 dpi, the overall mRNA
output of the sampled caecum tissue is dominated by parasite material with proportional mRNA
abundance of 77\% (in NMRI\_1stInf\_7dpi\_rep2) and 92\% (in
NMRI\_1stInf\_7dpi\_rep1). (Figure 1b).

\section*{Exclusion of samples with uncertain infection status}
A maximum of 86\% (sample NMRI\_1stInf\_7dpi\_rep1) and a minimum of
0.0326\% (sample NMRI\_2ndInf\_3dpi\_rep2) of mapped reads could be
assigned to the \textit{E. falciformis} genome in samples considered
infected (Table 1). We excluded samples NMRI\_2nd\_3dpi\_rep1 and NMRI\_2nd\_5dpi\_rep2
due to low parasite contribution to the overall transcriptome. 
Technically, this exclusion made it possible to obtain read counts in agreement with a 
negative binomial distribution (see additional file x). It is also likely that the number of reads in the excluded samples would have been
insufficient to fully normalise these datasets to those with the highest parasite contributions.
From a biological point of view, both excluded are samples from challenge
infection and we assume that the infection was cleared or reduced to a
non-detectable level. One sample (NMRI\_1stInf\_0dpi\_rep1) was excluded because
the uninfected control showed unexpected mapping of reads to the
\textit{E. falciformis} genome. We consider the three excluded samples to display an uncertain 
state of infection.

\section*{The mouse transcriptome changes in the course of infection}
Statistical testing for differential expression between infected and
uninfected mice revealed changes in mRNA abundance becoming more
pronounced (involving more genes) at later time-points post infection (especially
at 7 dpi; Table 2).

The observed changes between controls and 7 dpi were in agreement with
previously published microarry data. Fold-Change data obtained from
\textit{E. falciformis} infected mice at 6 dpi on Agilent microarrays 
(Schmidt et al 2012) or 7 dpi (our RNAseq data) against uninfected controls
show a strong correlation (Spearman's \rho = 0.74; Figure 2a). Considering both biological 
(e.g. exact time-points) and technical differences between the two methods this comparison 
confirms the adequacy of using dual RNA-seq for assessing the host transcriptome.

For mouse, 325 mRNAs were considered differently abundant (DA; FDR<1\%)
between controls and 3 dpi, 1,804 mRNAs between controls and 5 dpi and 2,711
mRNAs between controls and 7 dpi. This lead to a combined set of 3,453 genes
responding to infection (Figure 2bi). DA mRNAs early in infection (3 dpi
and 5 dpi) were not simply a subset of genes DA later in
infection (7 dpi), but instead, the transcriptional profile of the mouse 
changes throughout the time-points tested here. To further validate this pattern of a distinct
response in early infection we performed hierarchical clustering on
the (union of) mouse genes DA between different time-points post infection (Figure 2c).

By hierarchical clustering of mouse samples, three main sample clusters formed, “main clusters” 
(dendrogram of columns at top of Figure 2c). Rag1-/- mice cluster with control samples, and there is no clear distinction between infected and non-infected Rag1-/- samples, confirming the immune deficiency phenotype in Rag1-/-. Control samples are defined by high abundance of transcripts in gene cluster 3 (yellow, clusters of rows on the left) and partly high abundance in cluster 4 (red). Immunodeficient Rag1-/- samples share this profile with control samples in gene cluster 4, and this cluster has low abundance in (most) other infection samples. This suggests that these are genes which become down-regulated in immune competent mice upon infection. 
%%Two naive 5 dpi samples to the left cluster separately and diverge from this and most other patterns. High in infection, but does not fit with third NMRI replicate....
Three samples of challenged mice from 3, 5, and 7 dpi display high abundance of transcripts in gene cluster 6 (blue). The same genes have a low abundance in two naive, immune competent mouse samples from 5 dpi. Genes in cluster 6 are therefore particularly interesting for investigating the role of adaptive immunity in this infection (although sample NMRI\_2ndInf\_5dpi\_rep1 has a different profile). 
%% GO terms in cluster 6 discuss

Transcripts in gene clusters 3 and 4 are lowly abundant among all samples in the middle main cluster. Among those samples, the NMRI mouse samples display a trend towards high abundance in clusters 1, 2, 5 (green) and 7. Especially in the two late infection samples (7 dpi) a high transcript abundance is seen in clusters 1 and 5. The leftmost main sample cluster containing immunocompetent 5 dpi samples is characterized by low transcript abundance in cluster 6, and high abundance in cluster 7.

The pronounced changes later in infection reflect the expected
onset of an adaptive immune response. This is underlined by enriched Gene Ontology, GO, terms
in gene clusters with pronounced profiles late in infection (Figure 2c). Terms such as "antigen binding" and "immunoglobulin receptor binding" (molecular function, MF), and "immune system process" and "adaptive immune
response" (biological process, BP) are highly enriched in Cluster 5 and confirm an activated adaptive immune response
at this time-point.

Particular "regulation of cell death" and "regulation of apoptotic process" (BP)

"UDP-N-acetylglucosamine-lysosomal-enzyme" (MF)

%% This might be relevant: The DMAP interaction domain of
%% UDP-GlcNAc:lysosomal enzyme
%% N-acetylglucosamine-1-phosphotransferase is a substrate recognition
%% module. Catalysis of the reaction: UDP-N-acetyl-D-glucosamine +
%% lysosomal-enzyme D-mannose = UMP + lysosomal-enzyme
%% N-acetyl-D-glucosaminyl-phospho-D-mannose

General changes upon infection are a down-regulation of genes involved
"lipid modification" or more detailed "phosphatidylinositol
phosphorylation". These changes are similarly pronounce in the
immunocompetent strains and Rag1-/- immunodefficient animals (cluster 3
of fibure 2c) (see supplementary table X).  The presence of these
changes immunocompromised animals invite to the speculation that those
genes are independent of a potential later adaptive immune response.

Changes not detectable in Rag1-/- immunodefficient animals (cluster 3
of fibure 2c) but otherwise observed in general infection are
characterised by and enrichment to genes involved in "positive
regulation of calcineurin-NFAT", "blood coagulation" and "inositol
phosphate-mediated signaling" (see supplementary table X). The
absence of these changes in Rag1-/- mice allow us to speculate that
those genes might be involved in a early onset of adaptive
immune responses.

%% Go on here if you can...

We detected further differences between infections of naive and
challenge infected mouse hosts, especially at 7 dpi. These genes are
characterized by...
%% We should't limit ourselfs to testing the clusters only! But
%% discuss the cluster first if you got difficulties coding the
%% enrichment tests.

\section*{The life cycle of E. falciformis is characterized by major changes in mRNA abundance}

Table 2 lists pairwise sample comparisons and number of genes with
differently abundant mRNAs per comparison. We applied hierarchical clustering on 
\textit{E. falciformis} mRNAs which were differently abundant at different time-points 
post infection, including comparisons with oocysts and sporozoites. 
Upon hierarchical clustering (Figure 3c) changes throughout the parasite's intracellular life cycle 
can be categorised into an early (3 and 5 dpi) and late (7dpi) phase. 
Mouse immune status has no apparent influence on clustering of
parasite samples. Sporozoites show expression profiles similar to early infection stages
while oocysts show a distinct mRNA abundance profile (MDS plot...).

Parasite gene clusters shown on the left side (colour-coded and numbered 1 - 7) can be used 
to characterise clusters of samples. Late in infection, a profile of high mRNA abundance in clusters 2 and 7 (mint green and purple) is seen. Also parasite genes in cluster 6 (blue) appear characteristic  by displaying low mRNA abundance in late stages of infection. 
 One late challenge infection sample deviates from this pattern together with an early challenge infection sample. These two samples are, apart from controls and excluded samples, the ones with the least number of detected parasite reads (1836 and 1580, respectively), which taken together suggests that infections had been cleared before sampling in these challenge infected mice.

The early infection sample cluster (12 samples in center of heatmap) is characterised by a high abundance of transcripts in clusters 3 (yellow) and 6, as well as a below average abundance in clusters 1 (pink), 2, and 5 (green). 
Sporozoite samples have highly abundant transcripts in cluster 4 (red) and low abundance of 
transcripts in gene cluster 7. The latter is shared with oocyst samples. The oocyst samples, 
in addition, have low abundance of transcripts in gene clusters 2, 3, 4, 6 (and 7) and highly 
abundant transcripts in clusters 1 and 5.

Taken together, sample and gene clustering indicates that genes in clusters 2, 6 and 7 might be 
useful to characterise assumed gametocytes in 7 dpi samples. Genes in cluster 4 are candidates 
for sporozoite specific genes, and clusters 1 and 5  for oocyst specific genes in E. falciformis.

%% Mix it all well into the GO-term discussion! GO-term inferred
%% biiology needs to be reported in connection to clutering-inferred
%% ``sample biology''.

%% For these samples, the pattern is most pronounced in gene cluster 1
%% (up), and 4 and 6 (down). Within cluster 1, the second replicate from
%% the challenge infection has a deviant profile (compare mouse data in
%% Figure 2). Distinct groups of genes also define sporozoites (cluster 5
%% up) and oocysts (cluster 3 up, cluster 6 down). mRNA profiles on days
%% 3 and 5 p.i. from all three mouse strains cluster together. These
%% samples are distinct from oocysts, NMRI day 7 p.i., and
%% sporozoites. Sporozoites cluster closer with days 3 and 5 than with
%% oocysts or day 7.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection*{GO term enrichments in heatmap clusters}
The annotations referred to here are inferred from orthologs in other \textit{Eimeria} spp. or
in \textit{T. gondii}

\subsubsection*{Preparation for invasion in oocysts}
The mRNA profile in the oocyst stage is mainly determined by highly
abundant genes in cluster 4.  Overrepresented GO-terms in this cluster
are enriched by ortholog genes to peptidases, microneme localized
proteins reported to be involved in invasion, genes associated with
adhesion in protozoans (and with clotting in higher eukaryotes), and
genes that are annotated to be involved in amino acid biosynthesis.
Aminopeptidase N ('related' annotation) is the reported ortholog for
three genes with abundant mRNAs in oocysts. In humans, this enzyme has
been reported to cleave peptides bound to major histocompatibility
complex, MHC, II (UniProt reference if we want to keep this... but
does any secretion happen from oocysts...? Or is this too far-fetched
to be interesting?).

A Thrombospondin type 1 domain-containing protein ortholog is highly
abundant in cluster 4 (high abundance in oocysts). Thrombospondin type
1 domains have been reported in \textit{E. tenella} microneme
localizing proteins, MIC, e.g. MIC4 (Tomley01) In \textit{E. tenella}
MIC4 mRNA was reported in sporozoites where it localizes to the apical
end, and in late schizonts and late oocyst stages, when sporozoites
are forming. (Tomley01). For the same gene, the \textit{T. gondii}
annotation is Sushi domain-containing protein, which is also the
ortholog annotation of another gene in this cluster. In the related
malaria parasite \textit{P. falciparum} the apical sushi protein, ASP,
(which has a sushi domain) localizes to micronemes in merozoites but
not other stages (OKeeffe05).  Limulus clotting factor C, Coch-5b2
(Cochlin) and Lgl1, LCCL, (syn. F5/8 domain) domains are associated
discoidin lectin domains and thereby with adhesion.  (Pfam entries for
'LCCL domain' and 'Discoidin domain', May 2016).  Taken together, this
indicates that the thrombospondin and sushi-domain genes
(EfaB\_MINUS\_4114.g412 and EfaB\_PLUS\_1425.g183) are involved in
sporozoite invasion in \textit{E. falciformis} and that the mRNAs are
transcribed and available before excystation.  A role in merozoite
re-invasion in \textit{E. falciformis} is not indicated by our data.
The LCCL domain annotation and thrombospondins role in higher
eukaryotes also indicates that adhesion or preparation for adhesion is
important in oocysts.  We suggest (speculate...?) that the
thrombospondin annotated ortholog and the LCCL domain-containing
protein (EfaB\_MINUS\_11233.g986) are involved in cell adhesion in
\textit{E. falciformis}.
%One examples is found in the slime mold \textit{Dictyostelium
%discoideum)} in which the LCCL domain is part of the mold's discoidin
%adhesion protein. (Pfam entry for 'discoidin domain', May 2016)

\subsubsection*{Amino acid biosynthesis in oocysts}
High abundance of aminotransferase mRNAs indicate amino acid
biosynthesis or preparation for the same in oocysts (cluster 4).  We
identify D-3-phosphoglycerate dehydrogenase and alanine dehydrogenase
orthologs, which are enzymes contributing to L-serine and L-alanine
production, respectively. A putative \textit{Eimeria}
spp. cystathionine beta-synthase, CBS, in this cluster also indicates
de novo cysteine production. Alkyl sulfatase mRNA is also abundant in
oocysts. Generally, this enzyme enables an organism to exploit organic
sulfur to produce and incorporate inorganic sulfur into the amino
acids cysteine and methionine, when no inorganoc sulfur is available.

'Embryonic development' Nicalin 1, patched family protein (hedgehog)

\subsubsection*{Oocysts contain mRNA for fatty acid catabolism}
MmgE/PrpD is overrepresented in oocysts. The enzyme is important for
propionate catabolism in the 2-methylcitric acid cycle and has been
shown to be used by the intestinal intracellular bacterium
\textit{Salmonella typhimurium} to generate pyruvate (Horswill99).
Propionate is one of two most abundant small-chain fatty acids in the
gut along with butyrate. Both fatty acids are largely produced as
degradation products from food by commensal bacteria (Sun13). Sharing
the intestines as a niche with \textit{S. typhimurium} it is possible
that also \textit{E. falciformis} uses Mmg/PrpD to exploit available
propionate for pyruvate production.

\subsubsection*{Oocyst highly abundant mRNAs are downregulated in sporozoites}
Interestingly, the genes described above which are thought to be
involved in amino acid biosynthesis and invasion are highly abundant
in oocysts but are underrepresented in schizont stages (day 3 and day
5 samples) and even in sporozoites. An average abundance was detected
on day 7 for these genes, indicating a role in either gametes or early
oocyst formation.  This pattern supports the suggestion that these
specific mRNAs (cluster 4) for invasion and biosynthetic processes are
prepared (and possibly expressed) in the oocyst stage but are no
longer detectable in the cell at the timepont when the protein is
assumed to be in use (sporozoites and merozoite stages).  Therefore,
correlating mRNA prevalence with biological function at the timepoint
when mRNAs are detected must be done with care.

\subsubsection*{Down in sporozoites and oocysts -->  cluster 3..........}
Oocysts: profile for 3, 5, 6:

Sporozoites: 3 and 5

Day 7: 5 and 7

Specific genes....  Enolase 2, encoded by Eno2, is among the
downregulated genes in oocysts and sporozoites.  In \textit{T. gondii}
the paralog Eno1 is strongly associated with the cyst (bradyzoite)
stage and Eno2 is associated with tachyzoite stages (kibe05). It is
therefore expected that this mRNA is underrepresented in oocysts and
our data also show that the same is true in sporozoites for
\textit{E. falciformis}. (TK: If important we could look specifically
for Eno1 in cluster 4).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%% DAY xx   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection*{Down in sporozoites and oocysts -->  cluster 3..........}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%% DAY 7  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection*{Motility-related mRNAs indicate gamete development on day 7}
Two clusters contain genes with mRNAs highly abundant on day 7 p.i;
cluster 1 and 2.  Dynein, kinesin and tubulin are annotations highly
represented among orthologs of genes in both these clusters. The
annotations indicate an important role for motility at this timepoint,
probably reflecting development of microgametes.  In addition, in
cluster 2, there are two 'EF-hand domain containing proteins'
annotations as well as caltractin, centrin-1, and troponin
annotations. Caltractin and centrin-1 are associated with the
centrosome and structure and function of microtubuli in mammals, and
troponin is linked to muscle function (UniProt).  Also potentially
linked to motility is the occurence of growth arrest specific protein
8, Gas8, which in the mouse has been reported to be highly expressed
in the testes and important for mouse sperm function (Yeh02).

Other genes among the 38 indicate carbon fixation
(glycolysis/gluconeogenesis) or conversions of nucleoside
phosphates. In addition, a Ras family protein, RNA polymerase II
transcription initiation factor and Sec23 and Sec24 were among
orthologs identified in \textit{E. falciformis} cluster 2.

In cluster 1, carbon metabolism genes are represented by
6-phosphogluconate dehydrogenase and glycogen phosphorylase family
protein 1. UDP-glucose 4-epimerase and amiloride-sensitive amine
oxidase are reported as upregulated in gametocytes in
\textit{E. tenella} by RNA-seq (Walker15) and suggested by those
authors to play a role in cyst wall synthesis.

\subsubsection*{Microneme proteins highly expressed on day 7 p.i.}
Unintuitively for a protozoan organism, seven out of eight GO
biological process terms in cluster 1 are associated with wound
healing and blood coagulation.  An explanation is offered by some of
the orthologs to the three \textit{E. falciformis} genes responsible
for these terms. In protozoa, e.g., other \textit{Eimeria} spp. and
\textit{Toxoplasma gondii} orthologs are annotated as 'Micronemal
protein MIC4, related' (E. tenella) and more generally for several
other protozoa, 'PAN domain containing proteins'.  The PAN domain is
found in the plasminogen/hepatocyte growth factor family and in
coagulation factor XI family (REF), explaining why terms related to
blood coagulation are enriched by these genes.  Later publications on
\textit{T. gondii} (Marchant12) also associate PAN domains and
proteins in apicomplexan parasites with micronemes and therefore
invasion. In our case, this is peculiar, since the enrichment appears
on day 7 p.i.. A possible role at this timepoint is suggested by work
on the fungi \textit{Sclerotinia sclerotiorum} where Yu et
al. reported an important role for PAN domain proteins in cell wall
integrity (Yu12). This role for MIC proteins has to our knowledge not
been investigated in apicomplexan parasites.  The PAN domain domain
has also been reported to be common in nematodes such as
\textit{Caenorhabditis elegans}, however the function is not
understood. (Thordai99) The other two GO terms in the cluster of day
seven upregulated genes are DNA replication and DNA replication
initiation, which most likely reflects late stage schizogony or gamete
formation.  Six genes contribute to this enrichment and orthologs are
either annotated as DNA replcation licencing factors, DNA polymerases
or minichromosome maintenance proteins 2/3/5/7, Mcm2/3/5/7.

\subsubsection*{Gene and sample patterns by hierarchical clustering}
Samples (columns) cluster into two major clusters where day 7
p.i. samples form one group distinct from other samples. In the second
group, oocsts and sporozoites are distinct and sporozoites cluster
most closely with day 3 and 5 p.i. samples. Day 3 and 5 p.i. samples
also cluster into two groups, of which one contains all NMRI day 5
p.i. samples. Apart from this, the two day 3 and 5 p.i. sample
clusters have no obvious patters.

For gene clusters (rows), the two groups with high mRNA abundance on
day 7 p.i. (cluster 1 and 2) do not cluster most closely with each
other, but with the cluster for high mRNA abundance in oocysts
(cluster 1 association) and with the cluster for high mRNA abundance
in sporozoites (cluster 2 association).

\section*{Evolutionary conservation of expression changes in \textit{E. falciformis}}

For comparison of changes in the lifecycle of \textit{E. falciformis}
with the economically important chicken parasite \textit{E. tenella}
and the model apicompomplexan \textit{Toxoplasma gondii} we downloaded
expresson datasets from ...

Oocysts, Sproozoites

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%	DISCUSSION
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Discussion}
In our analysis we demonstrate which biological processes are dominant
in different life cycle stages of \textit{E. falciformis} in the
mouse. The RNAseq transcriptome provided here allows for detailed
analysis of genes involved in those processes, providing candidates
for life stage specific marker in \textit{Eimeria} spp. research.

\ldots
\ldots
\ldots

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%	METHODS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Methods}
\subsection{Mice and infection procedure}
Three strains of mice were used in our experiments: NMRI (Charles
River Laboratories, Sulzfeld, Germany), C57BL/6 (), and Rag1-/- on
C57BL/6 background (gift from Susanne Hartmann, FU?).  Animal
procedures were performed according to the German Animal Protection
Laws as directed and approved by the overseeing authority Landesamt
fuer Gesundheit und Soziales (Berlin, Germany). Animals where infected
as described by Schmid et al., (Schmid12), but tapwater was used
instead of PBS for administration of oocysts. Briefly, NMRI mice were
infected two times, which will be referred to as first and second
infection. For the first infection, 150 sporulated oocysts were
administered in 100 $μ$L by oral gavage. During the first infection of
60 mice, all animals were weighed every day. On day zero, before
infection, as well as on day three, five and seven days post
infection, dpi, caeca from 3-4 sacrificed mice per time point were
collected. Epithelial cells were isolated as described in Schmid et
al.(schmid12). For challenge infection, mice recovered for four weeks
before second infection.  Recovery was monitored by weighing and
visual inspection of fur. For the second infection, 1500 sporulated
oocysts were applied by oral gavage. Three mice were used as
non-second infection control, referred to as day 0, second infection.

\subsection{Oocyst purification for infection and sequencing}
Sporulated oocysts were purified by flotation from feces stored in
potassium dichromate and administered orally in 100 uL tapwater. One
\textit{E. falciformis} isolate, \textit{E. falciformis} Bayer
Haberkorn 1970, was used for all infections and parasite samples. The
strain is maintained through passage in NMRI mice in our facilities as
described elsewhere (schmid12).

\subsection{Sporozoite isolation}
Sporozoites were isolated from sporocysts by excystment. For this,
sporocysts were incubated at 37$°$C in DMEM containing 0.04\%
tauroglycocholate (MP Biomedicals) and 0.25\% trypsin (Applichem) for
30 min. Sporozoites were purified by the method of Schmatz et al
(schmatz--).

\subsection{RNA extraction}
Total RNA was isolated from infected epithelial cells, sporozoites and
sporulated oocysts using Trizol according to the manufacturer’s
protocol (Invitrogen). High quality \emph{what is the meaning of 'high
  quality' here?}  RNA was used to produce an mRNA library using the
Illumina’s TruSeq RNA Sample Preparation guide.  \emph{stolen from
  genome paper} Sporozoites were stored in 1 mL Trizol until
RNA-isolation.  Total RNA was isolated using the PureLink RNA Mini Kit
(Invitrogen) and reverse transcribed into cDNA.

\subsection{Sequencing, sequence quality assessment and alignment}
cDNA samples were sequenced by either GAIIX or Illumina Hiseq 2000 as
specified in SI xx (both unstranded). A fastq\_quality\_filter
(FASTQ-toolkit, version 0.0.14, available at
https://github.com/agordon/fastx\_toolkit.git) was applied to Illumina
Hiseq 2000 samples after replacing "N" bases by "." annotation.  A
phred score of 10 was applied. We further set q = 60. These settings
require that nine out of ten bases or more are correct in at least
60\% of the bases for each read.

\subsection{Alignment and reference genomes}
We used the published \textit{Mus musculus} mm10 assembly (Genome
Reference Consortium Mouse Build 38, GCA\_000001635.2) as reference
genome including annotations for mouse data. The
\textit{E. falciformis} genome (Heitlinger14) was downloaded from
ToxoDB (Gajria07). For the alignment, the mouse and parasite genome
files were merged into a dual reference genome, and files including
mRNA sequences from both species were aligned against the dual
reference genome using TopHat2 (version 2.0.14, Trapnell09) with -G
specified, and a Bowtie2 (version 1.1.2, Langmead12) index of the dual
genome. Single-end and pair-end sequence samples were aligned
separately with library type 'fr-unstranded' specified for pair-end
samples. Import into R was enabled by the R package Ballgown, which
requires bam files to be processed by Tablemaker (Frazee15), in our
case used with -qW -G specified. Tablemaker in turn makes use of
Cufflinks (version 2.1.1, Trapnell10).

\subsection{Differential mRNA abundance, data normalisation and sample exclusions}
Count data was normalized using the R-package edgeR (version 3.14.0;
cite) with the upperquartile normalisation method. Briefly, genes with
zero coverage in all samples (libraries) are removed and normalisation
factors are calculated for the 75\% quantile for each library. This
normalisation is suitable for read densities following a negative
binomial distribution. Two samples contradicted this assumption
(parasite data) for later modelling and both mouse and parasite data
from these samples were excluded from further analysis:
NMRI\_1st\_3dpi\_rep1 and NMRI\_2nd\_5dpi\_rep1 (SI ...). The method
then fits a generalized linear model (GLM with a negative binomial
link function) for each gene (glmFit) and then performs likelihood
ratio tests for models w or w/o focal factor (glmLRT).

%% ........... include outcome of MDS analysis here?
%% No, results never are presented in methods.

\subsection{Selection of differentially abundant mRNAs nad hierarchical clustering}
A selection of differently abundant mRNAs are used for hierarchical
clustering of \textit{E. falciformis} life cycle relevant genes. In
each comparison (see Table 3), the union of the at most 500 genes
differentially abundant with lowest FDR (<0.05) are selected. In the
next step, the 500 mRNAs from each comparison (or less) are
joined. For \textit{E. falciformis} time p.i. comparison, this
resulted in 1618 unique genes selected as differently expressed . The
22 genes in the NMRI vs C57BL/6 comparison are not included in the
\textit{E. falciformis} life cycle analysis. The same comparisons for
mouse yielded 8052 genes in total and 1313 unique ones. In heatmaps,
all samples, i.e. also samples which themselves did not have any
significantly different mRNAs according to our selection, were
included in hierarchical clustering. Scale bar in heatmaps show 0 as
mean mRNA abundance for each gene (row). Up (green) and
down-regulation (brown) denote number of standard deviations from 0,
i.e., row mean. Hierarchical clustering was performed using with
Euclidean distances, using complete linkage ('complete', R package
base).

%% the paragraph above has results. All those number

\subsection{Differentially abundant mRNAs used for hierarchical clustering}
All analyses were performed in R (cite R-core). Complete scripts are
available at https://github.com/derele/Ef\_RNAseq.git tagged as version
1.0.
%% We will make this available as a github webpage!!!

\setlength{\tabcolsep}{10pt}
\begin{table}[H]
\small
\begin{center}
\caption{Genes used for hierarchical clustering of mRNAs differently abundant depending on time p.i..}
\begin{tabular}{*3l}    \toprule
	\textit{Data description} & \emph{E. falciformis} genes & Mouse genes	\\ \midrule
	Sum of 1st infection NMRI sample differences	& 4935 & 8052 \\ 
	(including oocysts and sporozoites if appl.)	\\	
	Used in hierarchical clustering (heatmap)  	& 1618 & 1313 \\ 	\bottomrule	
\hline
\end{tabular}
\end{center}
%\end{adjustwidth}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Backmatter begins here                   %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{backmatter}

\section*{Competing interests}
  The authors declare that they have no competing interests.

\section*{Author's contributions}
    Text for this section \ldots

\section*{Acknowledgements}
  Text for this section \ldots
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                  The Bibliography                       %%
%%                                                         %%
%%  Bmc_mathpys.bst  will be used to                       %%
%%  create a .BBL file for submission.                     %%
%%  After submission of the .TEX file,                     %%
%%  you will be prompted to submit your .BBL file.         %%
%%                                                         %%
%%                                                         %%
%%  Note that the displayed Bibliography will not          %%
%%  necessarily be rendered by Latex exactly as specified  %%
%%  in the online Instructions for Authors.                %%
%%                                                         %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% if your bibliography is in bibtex format, use those commands:
\bibliographystyle{bmc-mathphys} % Style BST file (bmc-mathphys, vancouver, spbasic).
\bibliography{bmc_article}      % Bibliography file (usually '*.bib' )
% for author-year bibliography (bmc-mathphys or spbasic)
% a) write to bib file (bmc-mathphys only)
% @settings{label, options="nameyear"}
% b) uncomment next line
%\nocite{label}

% or include bibliography directly:
% \begin{thebibliography}
% \bibitem{b1}
% \end{thebibliography}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                               %%
%% Figures                       %%
%%                               %%
%% NB: this is for captions and  %%
%% Titles. All graphics must be  %%
%% submitted separately and NOT  %%
%% included in the Tex document  %%
%%                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%
%% Do not use \listoffigures as most will included as separate files

\section*{Figures}

\begin{figure}[h!]
%  \includegraphics[width=0.5\linewidth]{efmm-ratio}
  \caption{\csentence{Experimental design and outcome of a dual RNAseq experiment}
    read sequences per mouse read sequences for each experimental
    condition. The more advanced the infection is, the higher the
    ratio of parasite reads is, reflecting parasite replication
    between day zero and day seven. ......compare 1st and 2nd
    ....... patterns between Rag and NMRI/C57BL/6?  ........ Values
    are mean of replicates (n=2, if * n=1) on log10 scale. Each sample
    (replicate) consists of mRNA from three different mice.}
  \end{figure}

%%%%%
figure 1 label cotribution of mouse vs. parasite transcripts to the
overall transcriptional output of the analysed tissue for all samples.
%%%%

%% Selecting at most at most 500 genes differentially abundant with
%% lowest FDR (<0.05) are selected. In the next step, the 500 mRNAs from
%% each comparison (or less) are joined. For \textit{E. falciformis} time
%% p.i. comparison, this resulted in 1618 unique genes selected as
%% differently expressed . The 22 genes in the NMRI vs C57BL/6 comparison
%% are not included in the \textit{E. falciformis} life cycle
%% analysis. The same comparisons for mouse yielded 8052 genes in total
%% and 1313 unique ones. In heatmaps, all samples, i.e. also samples
%% which themselves did not have any significantly different mRNAs
%% according to our selection, were included in hierarchical
%% clustering.

%% This might be figure label
%% Scale bar in heatmaps show 0 as mean mRNA abundance for
%% each gene (row). Up (green) and down-regulation (brown) denote number
%% of standard deviations from 0, i.e., row mean.

\begin{figure}[h!]
%\begin{center}
%\centering
%%	\includegraphics[width=\linewidth]{EfLifecycleHeatmap.pdf}  
	\caption{\textit{E. falciformis} mRNAs with significantly
          different abundance at different times p.i. in NMRI mice
          (see Table 2). Samples cluster according to expected
          developmental state of the parasite, without distinguishing
          between first and challenge infection. Hierarchical
          clustering was set to apply seven gene clusters.}
%\end{center}
\end{figure}

\subsection{Parasite development reflected in mouse data}
\begin{figure}[h!]
%\begin{center}
%\centering
%% \includegraphics[width=\linewidth]{MmLifecycleHeatmap.pdf}  
\caption{Mouse mRNAs with significantly different abundance at
  different times p.i. in NMRI mice (see Table 2). Samples cluster
  according to expected developmental state of the parasite, without
  distinguishing between first and challenge infection. Rag1-/- mouse
  samples cluster together. Hierarchical clustering was set to apply
  four gene clusters.}
%\end{center}
\end{figure}

When clustering mouse mRNAs which were different in time
p.i. comparisons in Table 2, three out of four samples from day 7
p.i. (NMRI only) cluster together. These samples are characterized by
gene cluster 3 (up) as well as 2 and 4 (down). The fourth day 7-sample
(challenge infection, second replicate) clusters with day 3 and 5
samples. Upon visual inspection of this sample in gene cluster 4, it
diplays a profile similar to non-infected mice. The same sample is
abnormal in the parasite profile (Figure 1). NMRI and C57BL/6
uninfected samples (0dpi) cluster together, defined by gene clusters 4
(up), and 1 (down). Rag1-/- day 0 samples are however less pronounced
in gene cluster 1. Day 3 and 5 samples cluster together, with Rag1-/-
forming a separate group and Rag1-/- non-infected most distant in this
sample cluster.

% \begin{figure}[h!]
 % \includegraphics[width=0.7\linewidth]{ef-heatmap} \caption{\csentence{Parasite
 % genes with different mRNA abundance between samples, sorted into
 % five clusters by hierarchical clustering (method = complete,
 % distance = euclidean).}  \textit{E. falciformis} samples from seven
 % days p.i. cluster together (NMRI mice only). These samples have
 % distinct mRNA abundance patterns in all gene clusters, although
 % more pronounced in clusters 1 (up) and 4 (down). Distinct groups of
 % genes also define sporozoites (cluster 5, up) and oocysts (cluster
 % 2, up). mRNA profiles on days three and five p.i. from all three
 % mouse strains cluster together. These samples are distinct from
 % oocysts, NMRI day 7 p.i., and sporozoites, however closest to the
 % latter. On scale bar, 0 is mean mRNA abundance for each gene. Up
 % and downregulation is standard deviations from mean.}  \end{figure}

%\begin{figure}[h!]
%  \includegraphics[width=0.7\linewidth]{mm-heatmap} \caption{\csentence{Mouse
%  genes with different mRNA abundance between samples, sorted into
%  four clusters by hierarchical clustering (method = complete,
%  distance = euclidean).}  Three out of four samples from day 7
%  p.i. (NMRI only) cluster together. These samples are characterized
%  by genes in cluster 2 (up), 3, and 4 (down). NMRI and C57BL/6
%  uninfected samples cluster together (left), defined by clusters 3,
%  4 (up), and 1 (down). Rag1-/- samples cluster together. These
%  samples share a weak downregulation of most genes in cluster 2 as
%  well as upregulation of a small group genes in the same
%  cluster. Uninfected Rag1-/- samples are separated from infected
%  ones with distinct profiles in clusters 1, 3, and 4. On scale bar,
%  0 is mean mRNA abundance for each gene. Up and downregulation is
%  standard deviations from mean.}  \end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                               %%
%% Tables                        %%
%%                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% Use of \listoftables is discouraged.
%%

\newcommand{\bcell}[2][c]{%
  \begin{tabular}[#1]{@{}c@{}}#2\end{tabular}}

\definecolor{LightCyan}{rgb}{0.88,1,1}
\definecolor{LightRed}{rgb}{1,0.88,1}

% Fri Aug 12 11:05:43 2016
\begin{table}[ht]
\centering
\hspace*{-2.5cm}\begin{tabular}{lllllllll}
  \hline
Sample* & \bcell{Sequencing\\method} & batch & \bcell{total\\reads} & \bcell{reads\\mapping\\Mouse} & \bcell{reads\\mapping\\\textit{E. falciformis}} & \bcell{Percentage\\\textit{E. falciformis}}** & \bcell{detected\\ \textit{E. falciformis}\\genes} \\ 
  \hline
NMRI\_2ndInf\_0dpi\_rep1 & GAII & 2 & 108,937,797 & 70,489,674 & 247 & 0.0004 & 1 \\ 
  Rag\_1stInf\_0dpi\_rep1 & hiseq & 3 & 25,362,793 & 18,853,850 & 443 & 0.0023 & 2 \\ 
  C57BL6\_1stInf\_0dpi\_rep1 & hiseq & 3 & 35,731,249 & 25,119,348 & 457 & 0.0018 & 2 \\ 
  C57BL6\_1stInf\_0dpi\_rep2 & hiseq & 3 & 47,085,959 & 34,377,133 & 608 & 0.0018 & 2 \\ 
  Rag\_1stInf\_0dpi\_rep2 & hiseq & 3 & 46,556,156 & 35,233,327 & 676 & 0.0019 & 2 \\ 
  NMRI\_2ndInf\_0dpi\_rep2 & hiseq & 3 & 58,122,244 & 40,794,245 & 3,406 & 0.0083 & 51 \\ 
  \rowcolor{LightCyan}
  NMRI\_2ndInf\_3dpi\_rep1 & hiseq & 3 & 57,934,016 & 40,544,287 & 4,803 & 0.0118 & 95 \\ 
  \rowcolor{LightCyan}
  NMRI\_2ndInf\_5dpi\_rep2 & hiseq & 3 & 63,965,539 & 48,289,181 & 10,941 & 0.0227 & 407 \\ 
  \rowcolor{LightRed}
  NMRI\_1stInf\_0dpi\_rep1 & GAII & 1 & 82,364,585 & 55,176,243 & 17,954 & 0.0325 & 701 \\ 
  NMRI\_2ndInf\_3dpi\_rep2 & hiseq & 3 & 65,548,826 & 46,171,909 & 29,548 & 0.0640 & 1,580 \\ 
  NMRI\_2ndInf\_7dpi\_rep2 & hiseq & 3 & 67,487,466 & 51,722,265 & 40,091 & 0.0775 & 1,836 \\ 
  Rag\_1stInf\_5dpi\_rep1 & hiseq & 3 & 38,651,359 & 29,982,453 & 63,024 & 0.2098 & 2,548 \\ 
  Rag\_1stInf\_5dpi\_rep2 & hiseq & 3 & 34,779,832 & 25,297,803 & 99,000 & 0.3898 & 2,828 \\ 
  C57BL6\_1stInf\_5dpi\_rep1 & hiseq & 3 & 40,904,388 & 29,319,604 & 185,969 & 0.6303 & 4,173 \\ 
  Rag\_2ndInf\_5dpi\_rep1 & hiseq & 3 & 50,049,848 & 37,093,621 & 192,856 & 0.5172 & 4,167 \\ 
  C57BL6\_1stInf\_5dpi\_rep2 & hiseq & 3 & 29,511,368 & 18,062,349 & 215,696 & 1.1801 & 3,823 \\ 
  C57BL6\_2ndInf\_5dpi\_rep1 & hiseq & 3 & 35,148,432 & 25,660,184 & 262,909 & 1.0142 & 4,563 \\ 
  NMRI\_1stInf\_3dpi\_rep1 & GAII & 1 & 73,236,430 & 49,993,358 & 394,384 & 0.7827 & 5,220 \\ 
  NMRI\_1stInf\_3dpi\_rep2 & GAII & 2 & 160,709,694 & 117,791,044 & 413,051 & 0.3494 & 4,862 \\ 
  NMRI\_1stInf\_5dpi\_rep2 & GAII & 2 & 119,902,722 & 76,419,774 & 794,570 & 1.0290 & 5,333 \\ 
  NMRI\_2ndInf\_5dpi\_rep1 & GAII & 2 & 230,773,955 & 143,186,486 & 1,846,840 & 1.2734 & 5,533 \\ 
  NMRI\_2ndInf\_7dpi\_rep1 & hiseq & 3 & 70,366,762 & 41,467,146 & 8,634,201 & 17.2335 & 5,875 \\ 
  NMRI\_1stInf\_5dpi\_rep1 & GAII & 2 & 76,702,168 & 47,037,087 & 8,669,701 & 15.5631 & 5,700 \\ 
  NMRI\_sporozoites\_rep2 & GAII & 0 & 19,551,681 & 8,656 & 11,470,604 & 99.9246 & 5,513 \\ 
  NMRI\_1stInf\_5dpi\_rep3 & GAII & 0 & 191,099,180 & 83,735,624 & 27,839,458 & 24.9513 & 5,784 \\ 
  NMRI\_1stInf\_7dpi\_rep1 & GAII & 1 & 66,505,514 & 3,310,666 & 39,400,884 & 92.2488 & 5,932 \\ 
  NMRI\_sporozoites\_rep1 & GAII & 1 & 67,325,397 & 4,334 & 43,774,401 & 99.9901 & 5,825 \\ 
  NMRI\_oocysts\_rep1 & GAII & 1 & 68,859,802 & 3,805 & 49,653,065 & 99.9923 & 5,695 \\ 
  NMRI\_oocysts\_rep2 & GAII & 0 & 151,090,783 & 18,524 & 71,019,860 & 99.9739 & 5,777 \\ 
  NMRI\_1stInf\_7dpi\_rep2 & GAII & 1 & 139,749,046 & 21,699,324 & 73,539,445 & 77.2159 & 5,943 \\ 
   \hline
\end{tabular}
\begin{tablenotes}[flushleft]\footnotesize\singlespacing
\item{*} sample names are given as a) mouse strain b) first or challenge
  infection c) days post infection (dpi) and d) replicate number
  seperated by undersocre . \\
\item{**} percentag mapping \textit{E. falciformis} is given as percentage in total mapping reads
\end{tablenotes}
\end{table}
\hspace*{+2.5cm}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% 	NUMBER OF GENES DIFFERENT IN DIFFERENT COMPARISONS	
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% I like this table. Looks good and transprorts the right kind of
%% information

\clearpage
\section{mRNA abundance differences between different experimental groups}
\setlength{\tabcolsep}{8pt}
\begin{table}[H]
\small
\begin{center}
\caption{mRNA abundance differences between different experimental groups.}
%\label{tab:table3}
\begin{tabular}{*3l}    \toprule
\textit{Day post infection} & \textit{Ef} genes different & Mouse genes different \\ 
	\textit{comparisons} 	    & (FDR$\leq$1\%) &  (FDR$\leq$1\%/5\%) \\ \midrule
	NMRI 0 vs NMRI 3		& NA   & 274 \\
	NMRI 0 vs NMRI 5		& NA   & 1736 \\
	NMRI 0 vs NMRI 7		& NA   & 2802 \\
	NMRI 3 vs NMRI 5     		& 111  & 1 \\
	NMRI 3 vs NMRI 7  		& 1385 & 1407 \\ 
	NMRI 5 vs NMRI 7  		& 1895 & 873 \\ 
	C57BL/6 0 vs C57BL/6 5		& NA	& 914 \\
	Rag1-/- 0 vs Rag1-/- 5		& NA	& 45 \\ \midrule
\textit{Day post infection,} & 		 & 		 \\ 
\textit{parasite relevant comparisons} 	    & 		&  \\ \midrule
	Oocysts vs NMRI 3  	& 3310 & NA \\  
	Oocysts vs NMRI 5	& 3605 & NA \\ 
	Oocysts vs NMRI 7	& 3085 & NA \\ 
	Oocysts vs sporozoites  & 3421 & NA \\
	Sporozoites vs NMRI 3 	& 1663 & NA \\
	Sporozoites vs NMRI 5 	& 1605 & NA \\
	Sporozoites vs NMRI 7 	& 2473 & NA \\ \midrule
\textit{First and second infection} & 		 & 	 \\ 
\textit{comparisons} 	    & 		& 	\\ \midrule
	NMRI 3 1st vs NMRI 3 2nd  	& 0  & 5 \\
	NMRI 5 1st vs NMRI 5 2nd  	& 0  & 1 \\
	NMRI 7 1st vs NMRI 7 2nd  	& 0  & 902 \\
	C57BL/6 1st vs C57BL/6 2nd (day 5) & 0 &  mouse \\
	Rag1-/- 1st vs Rag1-/- 2nd (day 5) & 0 & mouse  \\ \midrule
\bottomrule
	\hline
\end{tabular}
\end{center}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                               %%
%% Additional Files              %%
%%                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section*{Additional Files}
  \subsection*{Additional file 1 --- Raw and normalized counts }
  Raw counts of reads mappins to the \textit{E. falciformis} and mouse
  genome for individual samples in our study. Normalized counts for
  seperately for the host and parasite mappings (three compressed csv
  files).

  \subsection*{Additional file 2 --- Results of statistical tests (edgeR) }
  Focal contrast, fold-changes, likelihood ratio in/excluding this
  difference in models, p-values , and false discovery rates (adjusted
  p-values) are given for all tested contrasts (one compressed csv
  file).

  \subsection*{Additional file 3 --- Additional methods and resuts }
  % this would be cool but requires special referencing in the text
  % (hard to keep track and hard to fit in the journals referencing
  % system). But we can try.
  Document containing additonal figures and summary tables (pdf).

\subsection*{Additional file 4 --- Results of enrichment analyses (topGO)}
  Tables listing all tested gene sets and resulting significatn GO
  terms.

\end{backmatter}
\end{document}