### 2016-09-08 manuscript/bmc_article.tex
commit manuscript before starting to mess with including bibliography

%% BioMed_Central_Tex_Template_v1.06
%%                                      %
%  bmc_article.tex            ver: 1.06 %
%                                       %

%%IMPORTANT: do not delete the first line of this template
%%It must be present to enable the BMC Submission system to
%%recognise this template!!

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                     %%
%%  LaTeX template for BioMed Central  %%
%%     journal article submissions     %%
%%                                     %%
%%          <8 June 2012>              %%
%%                                     %%
%%                                     %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                                                 %%
%% For instructions on how to fill out this Tex template           %%
%% document please refer to Readme.html and the instructions for   %%
%% authors page on the biomed central website                      %%
%% http://www.biomedcentral.com/info/authors/                      %%
%%                                                                 %%
%% Please do not use \input{...} to include other tex files.       %%
%% Submit your LaTeX manuscript as one .tex document.              %%
%%                                                                 %%
%% All additional figures and files should be attached             %%
%% separately and not embedded in the \TeX\ document itself.       %%
%%                                                                 %%
%% BioMed Central currently use the MikTex distribution of         %%
%% TeX for Windows) of TeX and LaTeX.  This is available from      %%
%% http://www.miktex.org                                           %%
%%                                                                 %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%% additional documentclass options:
%  [doublespacing]
%  [linenumbers]   - put the line numbers on margins

%%% loading packages, author definitions

%\documentclass[twocolumn]{bmcart}% uncomment this for twocolumn layout and comment line below
\documentclass{bmcart}

%%% Load packages
\usepackage{amsthm,amsmath}
\RequirePackage{natbib}
%\RequirePackage[authoryear]{natbib}% uncomment this for author-year bibliography
\RequirePackage{hyperref}
\usepackage[utf8]{inputenc} %unicode support
%\usepackage[applemac]{inputenc} %applemac support if unicode package fails
%\usepackage[latin1]{inputenc} %UNIX support if unicode package fails
%\usepackage{siunitx}
\usepackage[colorlinks]{hyperref} % To color blocks of text 
%\usepackage[dvipsnames]{xcolor} % More color options (see Wikibooks)  
\usepackage{changepage}
\usepackage{booktabs} % For prettier tables
\usepackage{float} % To control positions of e.g. tables or figures
%\usepackage[section]{placeins}
\usepackage{graphicx}
\usepackage[flushleft]{threeparttable}

% to color table rows
\usepackage{color, colortbl}
\graphicspath{ {../} }

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                             %%
%%  If you wish to display your graphics for   %%
%%  your own use using includegraphic or       %%
%%  includegraphics, then comment out the      %%
%%  following two lines of code.               %%
%%  NB: These line *must* be included when     %%
%%  submitting to BMC.                         %%
%%  All figure files must be submitted as      %%
%%  separate graphics through the BMC          %%
%%  submission process, not included in the    %%
%%  submitted article.                         %%
%%                                             %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\def\includegraphic{}
%\def\includegraphics{}

%%% Put your definitions there:
\startlocaldefs
\endlocaldefs

%%% Begin ...
\begin{document}

%%% Start of article front matter
\begin{frontmatter}

\begin{fmbox}
\dochead{Research}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter the title of your article here     %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{Dual host-parasite transcriptomes of apicomplexan Eimeria
  falciformis and its natural mouse host}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter the authors here                   %%
%%                                          %%
%% Specify information, if available,       %%
%% in the form:                             %%
%%   <key>={<id1>,<id2>}                    %%
%%   <key>=                                 %%
%% Comment or delete the keys which are     %%
%% not used. Repeat \author command as much %%
%% as required.                             %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\author[ addressref={aff1},] {\inits{TK}\fnm{Totta} \snm{Kasemo}}
\author[ addressref={aff1},] {\inits{SS}\fnm{Simone} \snm{Spork}}
\author[ addressref={aff2},] {\inits{CD}\fnm{Christoph} \snm{Dieterich}}
\author[ addressref={aff1}, ]{\inits{RL}\fnm{Richard} \snm{Lucius}}
\author[ addressref={aff1, aff3},              
  corref={aff1},                   
  email={emanuel.heitlinger@hu-berlin.de}]{\inits{EH}\fnm{Emanuel} \snm{Heitlinger}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter the authors' addresses here        %%
%%                                          %%
%% Repeat \address commands as much as      %%
%% required.                                %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\address[id=aff1]{
	\orgname{Institute of Biology, Humboldt-Universitat zu Berlin}, 
	\street{Philippstr. 13, Haus 14},
	\postcode{10115}                 
	\city{Berlin},                   
	\cny{Germany}                    
}
\address[id=aff2]{  
	\orgname{Max Planck Institute for Biology of Ageing},
	\street{Joseph-Stelzmann-Strasse 9B},
	\postcode{50931}
	\city{Cologne},
	\cny{Germany}
}
\address[id=aff3]{  
	\orgname{Leibniz Institute of Zoo and Wildlife Research},
	\street{Alfred-Kowalke-Str. 17},
	\postcode{10315}
	\city{Berlin},
	\cny{Germany}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter short notes here                   %%
%%                                          %%
%% Short notes will be after addresses      %%
%% on first page.                           %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{artnotes}
%\note{Sample of title note}     % note to the article
%\note[id=n1]{Equal contributor} % note, connected to author
%\note[id=n2]
\end{artnotes}

%\end{fmbox}% comment this for two column layout

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% The Abstract begins here                 %%
%%                                          %%
%% Please refer to the Instructions for     %%
%% authors on http://www.biomedcentral.com  %%
%% and include the section headings         %%
%% accordingly for your article type.       %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{abstractbox}

\begin{abstract} % abstract
\parttitle{First part title} %if any
Text for this section.

\parttitle{Second part title} %if any
Text for this section.
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% The keywords begin here                  %%
%%                                          %%
%% Put each keyword in separate \kwd{}.     %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{keyword}
\kwd{Parasite, apicomplexa, RNA-seq, transcriptome, life-cycle, interaction}
%\kwd{article}
%\kwd{author}
\end{keyword}

% MSC classifications codes, if any
%\begin{keyword}[class=AMS]
%\kwd[Primary ]{}
%\kwd{}
%\kwd[; secondary ]{}
%\end{keyword}

\end{abstractbox}
\end{fmbox}% uncomment this for twcolumn layout
\end{frontmatter}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% The Main Body begins here                %%
%%                                          %%
%% Please refer to the instructions for     %%
%% authors on:                              %%
%% http://www.biomedcentral.com/info/authors%%
%% and include the section headings         %%
%% accordingly for your article type.       %%
%%                                          %%
%% See the Results and Discussion section   %%
%% for details on how to create sub-sections%%
%%                                          %%
%% use \cite{...} to cite references        %%
%%  \cite{koon} and                         %%
%%  \cite{oreg,khar,zvai,xjon,schn,pond}    %%
%%  \nocite{smith,marg,hunn,advi,koha,mouse}%%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%% start of article main body
% <put your article body there>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%	INTRODUCTION
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section*{Introduction}
Text and results for this section, as per the individual journal's instructions for authors. 
%\cite{koon,oreg,khar,zvai,xjon,schn,pond,smith,marg,hunn,advi,koha,mouse}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%	RESULTS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

  \section*{Results}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% 	OVERVIEW READS PER SAMPLE, LAYOT AS EXPERIMENTAL OVERVIEW
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

 \section*{A dual transcriptomics experiment}

We performed mRNA sequencing of caecum epithelial tissue from mice infected with
the apicomplexan parasite \textit{E. falciformis}. Oocysts and
sporozoites were included as "environmental" stages and processed
in vitro. To follow the life cycle of the parasite, we
compare different time-points after infection. We additionally used different mouse strains which display different immunocompetence in infection trials, measured by oocyst output in order to assess the influence of host immunocompetence on parasite development (Figure 1a and 1b). Immune competent NMRI mice were infected and sampled at three time-points post infection, p.i. We also infected mice lacking the Recombination activating gene 1, Rag1, (Rag1-/-) and compared them to the parental C57BL/6 mouse line. Rag1-/- mice lack mature B and T lymphocytes, which is taken as a proxy for absent adaptive immunity. All three mouse strains were used for infections of both naive mice and previously infected (and recovered) animals (onward referred to as "challenged" or "challenge infection").
\newline Basic phenotyping (Figure 1a) showed differences in oocyst numbers between the immunocompetent (C57BL/6) and immunodeficient (Rag1-/-) host strains in naive and challenge infected mice. Immunocompetent NMRI mice were infected with a higher dosis of sporulated oocysts, and a drastic reduction of oocysts in faeces was seen in challenged mice compared to naive mice. Oocyst numbers in faeces peaked on days 8-9 andall mice had cleared the infection by day 14. Development and replication of \textit{E. falciformis} intestinal stages is reflected by the percentage of parasite reads sequenced per time-point post infection (Figure 1b) and this was confirmed by quantitative reverse transcription PCR (RT qPCR) of parasitic 18S (Figure 1c).
\newline We thus use dual RNA-seq to analyse the life cycle of \textit{E. falciformis} under
the influence of host immune responses at early and late stages of infection. We used an experimental
design which allows to compare infections at 5 days post infection (dpi) for all experimental
conditions (NMRI, C57BL/6 and Rag1-/- mouse strains in naive and challenge infection). Additional time-points 3 dpi and 7 dpi were analysed from NMRI mice (Figure 1d).

\section*{Parasite and host dual transcriptomes can be assessed in parallel}
In this RNA-seq experiment, each replicate sample was enriched for caecum epithelial tissue and pooled 
from three mice. mRNA was extracted and sequencing libraries prepared. Two biological replicates were used for all but two conditions (with one and three replicates, respectively). Libraries were sequenced on several lanes of Illumina GAIIX (13 samples) and HiSeq machines (14 samples) and mapped to both mouse and parasite genomes simultaneously to avoid spurious assignments of reads in ultra conserved genomic regions. As samples and individual replicates were sequenced in batches to different depth and using different instrumentation (Table 1) we performed quality controls (additional files xyz). These confirm the absence of batch effects influencing analysis and quality of results.

Total numbers of sequenced reads as well as reads mapped to either the \textit{E. falciformis} genome or the mouse genome are indicated in Table 1 for all replicates. The number of total read mappings for individual replicates ranged from 25,362,739 (sample Rag\_1stInf\_0dpi\_rep1) to 139,749,046 (NMRI\_1stInf\_7dpi\_rep2). At the latest time-point samples, 7 dpi, the overall mRNA output of the sampled caecum tissue is dominated by parasite material with proportional mRNA abundance of 77\% (in NMRI\_1stInf\_7dpi\_rep2) and 92\% (in NMRI\_1stInf\_7dpi\_rep1). (Figure 1b).

\section*{Exclusion of samples with uncertain infection status}
A maximum of 92\% (sample NMRI\_1stInf\_7dpi\_rep1) and a minimum of 0.064\% (sample NMRI\_2ndInf\_3dpi\_rep2) of mapped reads could be assigned to the \textit{E. falciformis} genome in samples considered infected (Table 1). We excluded samples NMRI\_2nd\_3dpi\_rep1 and NMRI\_2nd\_5dpi\_rep2 due to low parasite contribution to the overall transcriptome. Technically, this exclusion made it possible to obtain read counts in agreement with a negative binomial distribution (see additional file x). It is also likely that the number of reads in the excluded samples would have been insufficient to fully normalise these datasets to those with the highest parasite contributions. From a biological point of view, both excluded samples are samples from challenge infection and we assume that the infection was cleared or reduced to a non-detectable level. One sample (NMRI\_1stInf\_0dpi\_rep1) was excluded because the uninfected control showed unexpected mapping of reads to the \textit{E. falciformis} genome. We consider the three excluded samples to display an uncertain state of infection.

\section*{The mouse transcriptome changes upon \emph{E. falciformis} infection}
Statistical testing for differential expression between infected and uninfected mice revealed changes in mRNA abundance becoming more pronounced (containing more genes) at later time-points post infection (Table 2). For mouse, 325 mRNAs were considered differently abundant (DA; FDR < 0.01) between controls and 3 dpi, 1,804 mRNAs between controls and 5 dpi and 2,711 mRNAs between controls and 7 dpi. This lead to a combined set of 3,453 unique genes responding to infection (Figure 2bi). DA mRNAs early in infection (3 dpi and 5 dpi) were not a strict subset of genes DA later in infection (7 dpi). Instead, the transcriptional profile of the mouse changes throughout the infection. Differences between controls and 7 dpi were in agreement with previously published microarry data. Fold-change data obtained from \textit{E. falciformis} infected mice at 6 dpi on Agilent microarrays (\cite{schmidt12}) or 7 dpi (our RNA-seq data) against uninfected controls show a strong correlation (Spearman's $\rho$ = 0.74; Figure 2a). Considering both biological differences in the experiments (e.g. exact time-points for samples), and technical differences between the two methods this comparison confirms the adequacy of using dual RNA-seq for assessing the host transcriptome.

\section*{Epithelial responses vary with mouse immune status}
A distinct response in early infection is indicated by the number of DA mRNAs between different time-points. To further validate the pattern we performed hierarchical clustering on the (union of) mouse genes DA between different time-points post infection (Figure 2c). By hierarchical clustering of mouse samples, three main sample clusters formed (dendrogram of columns at top of Figure 2c). Immune deficient Rag1-/- mice cluster with control samples. There is no clear distinction between infected and non-infected Rag1-/- samples, which confirms the immune deficiency phenotype in these mice. 
We identify a group of genes (cluster 4, Figure 2c) which changes transcriptional profile upon infection (all time-points of infected samples) in immune competent mice only. In immune compromised Rag1-/-, another group of genes (cluster 3, figure 2c) display the "infection profile" of immune competent mice also in control samples, indicating that these are mRNAs kept at a low abundance in healthy animals in a T and B cell dependent manner. Hence, these genes in some way depend on functional T and B cells also in controls (cluster 3) and at the earliest time-point post infection (3 dpi, cluster 4). 
Gene Ontology, GO, terms enriched in cluster 3 are, e.g., "lipid modification" and "lipid metabolic process" and "protein intraciliary transport". Other enriched terms in cluster 3 are terms for regulation of other metabolic processes, for blood coagulation as well as terms containing "spinal cord", "axon" or "neuronal" regulation or development. (SI file x). We suggest that these are processes which are all intrinsically different in Rag1-/- mice compared to immune competent mice. The genes in cluster 4, which change only in immune competent mice, are highly abundant in controls including Rag1-/-. Infected immune competent mice have a lower abundance of these mRNAs.
 Therefore, these genes appear to be down-regulated upon infection by processes which depend on (mature) T and/or B cell activity, surprisingly also at 3 dpi. Enriched GO terms in this gene cluster include 13 terms for cytokines, particularly interleukin-1, IL-1, ("negative regulation of" and "secretion of"). Terms such as "negative regulation of viral (or inflammatory) response", "negative chemotaxis", "autophagy", "blood coagulation", "inositol phosphate-mediated signaling", and "positive regulation of calcineurin-NFAT" are also enriched in cluster 4. Inositol signaling can lead to release of calcium and calcineurin-dependent translocation of NFAT to the nucleus and activation of its target genes in T cells, but also many other cell types (reviewed by macian05). GO enrichment of cluster 4 also highlights particular regulation of transforming growth factor$\beta$, TGF$\beta$, epidermal growth factor, EGF, and tumor necrosis factor, TNF. TGF$\beta$ is important for wound healing in intestinal epithelium (beck03), and EGF regulates proliferation of epithelial cells and inhibits apoptosis (suzuki10). TNF appears to be dose-dependent and capable to suppress inflammatory responses (noti10) but is also reported to regulate proliferation of epithelial cells (kaiser97). Additionally, IL-1 and IL-6 are among the enriched GO terms in cluster 4. The IL-1 receptor (type I) is similar to Toll-like receptors and IL-1 induces innate immune responses in many cell types, and influences lymphocyte activity (dinarello09). IL-6 has been shown to support repair and inhibit apoptosis after epithelial wounding (kuhn14) and it is known to be important for development of Th17 responses (ref in kuhn intro). 
%Th17 .... Stange's work (discuss Richard). 
IL-6 induces the Janus kinase, JAK, and signal transducer and activator of transcription STAT3, which, among other functions, is involved in wound healing in intestinal epithelial cells (pickert09). This analysis suggests that TGF$\beta$, TNF, EGF, IL-1 and IL-6 are main actors in the epithelial response to \textit{E. falciformis} infection and that the response is T and B cell dependent, also at early time-points. Enrichment of these biological processes further indicates that hosts invest resources in intestinal healing also at early time-points of \textit{E. falciformis} infection.

\newline The pronounced changes late in infection (7 dpi) reflect the expected onset of an adaptive immune response. This is underlined by enriched GO terms in gene clusters 1 and 5. Terms such as "antigen binding" and "immunoglobulin receptor binding" (molecular function, MF), and "immune system process", "adaptive immune response"and "innate immune response" (biological process, BP) are highly enriched in cluster 5 and confirm an activated adaptive immune response at this time-point. Among the same genes, natural killer cell regulation, JAK-STAT signalling, and IL-1 and interleukin-2, IL-2 production are enriched biological processes. IL-2 is one target of NFAT signaling and as mentioned above, JAK-STAT signaling can be induced by IL-6. This indicates that the early responses in, e.g., NFAT and IL-6 regulation induce distinct mRNA abundace differences later in infection (7 dpi).

Taken together, clusters 3 and 4 highlight the difference between Rag1-/- mice and immune competent mice in this parasitic infection. Some genes are different also between control animals (Rag1-/- and NMRI, cluster 3). These do not change upon infection in mice lacking mature T and B cells. Other genes (cluster 4) for which the transcriptional profile is similar in all controls, mRNA abundance also changes upon infection only in animals with functional maturation of T and B cells. Using GO enrichment analysis, we show that regulation of IL-1, IL-6, TNF,TGF$\beta$ and EGF are regulated in both first and challenge infections and suggest wound healing as a probably function of this regulation, especially at 7 dpi.

\section*{RNA processing is enriched in challenge infected immune competent mice} 
Three challenge infected samples (3 dpi, 5 dpi and 7 dpi) from immune competent mice show a distinct profile in cluster 6. This cluster is highly enriched for GO terms for RNA processing and splicing, as well as terms for histone and chromatin modification. \ldots
\newline
Include or not?
- daunorubicin metabolic process > cancer drug which prevents DNA replication by stabilizing DNA
- doxorubicin metabolic process  > cancer drug
\newline

%%%% INPUT
"UDP-N-acetylglucosamine-lysosomal-enzyme" (MF)
%%%%% here %%%%%

%% This might be relevant: The DMAP interaction domain of
%% UDP-GlcNAc:lysosomal enzyme
%% N-acetylglucosamine-1-phosphotransferase is a substrate recognition
%% module. Catalysis of the reaction: UDP-N-acetyl-D-glucosamine +
%% lysosomal-enzyme D-mannose = UMP + lysosomal-enzyme
%% N-acetyl-D-glucosaminyl-phospho-D-mannose

%% Go on here if you can...

%We detected further differences between infections of naive and
%challenge infected mouse hosts, especially at 7 dpi. These genes are
%characterized by...
%% We should't limit ourselfs to testing the clusters only! But
%% discuss the cluster first if you got difficulties coding the
%% enrichment tests.

\section*{Transcriptional differences in the parasite life cycle are independent of mouse immune status}
We performed statistical tests to evaluate significant differences (FDR$<$0.01) in mRNA abundance between different parasite life cycle stages, approximated by time post infection (Table 2). Between early time-points, 3 dpi and 5 dpi, 103 mRNAs were different, whereas between 3 dpi and 7 dpi 1399 mRNAs were DA, and between 5 dpi and 7 dpi 2084 mRNAs were DA (Figure 3a). This indicates that the major changes take place between 5 dpi and 7 dpi, and that variation is smaller between 3 dpi and 5 dpi. This motivated us to define 3 dpi and 5 dpi as "early infection" and 7 dpi as "late infection". Early and late infection samples were tested for DA compared to sporozoites and sporulated oocysts, resulting in 1697 and 3919 DA mRNAs, respectively.
%% Check venn vs table2 and change in ms + code
To evaluate this outcome further, we performed hierarchical clustering of (the union of) the DA mRNAs in the comparisons described above and applied GO enrichment analysis of gene clusters (annotations are inferred from orthologs in other \textit{Eimeria} spp. or \textit{T. gondii}). In the parasite transcriptome, we see no difference between infection in immune competent mice, or T and B cell deficient Rag1-/- mice, or between naive and challenge infected mice. This is surprising considering the measured differences in oocyst output in the same comparisons (Figure 1a), and the fact that these differences are visible in the mouse transcriptomes. On the parasite side, major patters instead seem to be determined by life cycle stages, independent of the host immune status. Distinct clusters of genes define early infection (3 dpi and 5 dpi) in which schizogony (asexual reproduction) takes place, and separately, late infection (7 dpi) in which it is assumed that gametocytes are present. Extracellular samples of sporozoites and oocysts also cluster separately and are defined by distinct gene clusters.

\section*{Early infection transcriptomes reflect parasite expansion}
GO enrichment of the major sporozoite defining cluster with high mRNA abundance in this stage (cluster 4) proposes that different biosynthesis processes are important, as well as "maintenance of protein location in cell". Possibly, the latter is due to control of microneme protein localization as sporozoites prepare for invasion. Biosynthesis terms probably reflect preparation for asexual expansion and perhaps ATP production for motility in sporozoites, although no GO terms for motility are enriched in this cluster. In the following intracellular life cycle stages, measured at 3 dpi and 5 dpi, in which several rounds of schizogony (asexual replication) take place, mRNAs in cluster 6 are abundant in all early samples (except one, discussed below). Among these "early infection" genes in cluster 6, several GO terms (biological process) for biosynthetic acitivity are enriched, e.g., "ribosome biogenesis" and "cellular biosynthetic process", as well as terms for "gene expression" and RNA processing, including terms for tRNAs and ncRNAs. "Cellular amino acid catabolic process" is also enriched. In cluster 3, with a similar profile as cluster 6, no GO terms are enriched. Cluster 6 is distinct in early infection with a biosynthesis profile shared with invasive sporozoites. Replication and growth-related processes being enriched highlights the parasite's expansion in numbers on 3 dpi and 5 dpi, as supported both by previous knowledge about the life cycle and our increase in sequences from the parasite (Figure 1b).

\section*{Gametocytes likely determine transcriptome late in infection}
Two gene clusters have a distinct profile with high mRNA abundance on 7 dpi (clusters 2 and 7). Both clusters display low mRNA abundance in other life cycle stages, especially in oocysts and sporozoites. Enriched GO terms such as "movement of cell or subcellular component" and "microtubule-based movement" along with terms suggesting ATP production ("ATP generation from ADP") indicate the presence of motile and energy demanding gametocytes in these samples. Other cluster 2 terms such as "chitin metabolic process" along with enriched metabolic and biosynthetic processes and "gene expression" in cluster 7 suggests ongoing encystation at 7 dpi, fitting the timing of oocyst output which peaks at 8-9 dpi. In addition, cluster 2 is enriched for a number of GO terms for "blood coagulation" and reflect 
%%%(double check) 
the presence of Thrombospondin type I domains in the protein products of cluster 2 mRNAs. Thrombospondin type 1 domains have been reported in \textit{E. tenella} microneme localizing proteins, MIC, e.g. MIC4 (Tomley01). MIC4 mRNA was reported in sporozoites where it localizes to the apical end, and in late schizonts and late oocyst stages, when sporozoites are forming. This suggests that \textit{E. falciformis} prepares for invasion already in gametocytes or during oocyst formation, however this is speculative.

\section*{Oocysts are characterized by stress responses and differentiation}
Clusters 1 and 5 contain mRNAs with high abundance in oocysts and all other clusters have low abundance in this stage. GO enrichment in cluster 1 contains only one term (adj. p-value 0.11) for "DNA-templated transcription, initiation". Cluster 5 is enriched for terms related to stress responses, "DNA repair", "protein modification process" and for "cell differentiation". Stress responses and DNA repair can be a result of storage in potassium dichromate of mouse faeces with oocysts before purification. Initiation of transcription and cell differentiation probably reflects (slow) preparation for invasion when oocysts are taken out of storage and purified for RNA extraction. Overall, the oocyst profile in five out of seven gene clusters is characterized by below average abundance of mRNAs, as can be expected in this life cycle stage which endures long-term survival outside the host. Taken together, sample and gene clustering indicates that genes in clusters 6 are abundant only early in infection and are candidates for merozoite specific genes, whereas clusters 2 and 7 might be useful to characterise gametocytes. Genes in cluster 4 can be considered sporozoite specific, and clusters 1 and 5 as oocyst specific genes in \textit{E. falciformis}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% comment on weird day 7 day 3 clustering
\section*{Imperfect clustering might reflect true biological differences}
In parasite data, one late challenge infection sample deviates from the "late infection pattern" and clusters with an early challenge infection sample. These two samples are, apart from controls and excluded samples (see above), the ones with the least number of detected parasite genes (1836 and 1580, respectively). This, along with the unclear transcriptional profile and the fact that they are both challenge infections in immune competent hosts, suggest that infections had been cleared before sampling. If this is the case, it raises the question of why it was cleared in these samples but not in other challenge infected immune competent mice. The answer can be technical problems in the experiment. However, other patterns suggest a different interpretation. Hierarchical clustering analysis in most cases does not cluster replicates together, which at a first glance suggests such problems. On the other hand, very distinct samples such as oocysts and sporozoites do cluster as replicates. Also, considering the early/late infection patterns in the parasite data and that these fit well with previous knowledge about this infection (asynchronous schizogony and gametocyte formation at 7 dpi), it is worthwhile to consider whether replicate separation reflects true biological variation between hosts and possibly also parasites (oocysts used for infection). It is perceivable that the parasite accommodates to host variation with small differences in, e.g., host stress levels due to litter mates, draught, differences in ght exposure or other factors which may vary also in a controlled animal facility. Parasite adjustment to such factors could explain the transcriptional profiles we see, with overall patterns but replicate separation. We suggest that considering such possibilities is useful for interpreting results and understanding basic biology of the parasite, and, in extension better understand infections in less homogeneous hosts than laboratory mice.

\section*{Evolutionary conservation in life cycle-characteristic gene groups}
To gain insight about the conservation status of life stage-specific gene groups in 
\textit{E. falciformis}, we tested gene clusters in Figure 3c for enrichment of gene orthologs which 
are present in other species. Groups analyzed for orthologs are: i) one category with \textit{E. falciformis} only ("Efalci"), ii) one category of 10 apicomplexan parasites ("Api") (see xxx), iii) one category of Api but excluding \textit{Cryprosporidium hominis} ("Api minus C.h."), iv) one catogory with three \textit{Eimeria} species: \textit{E. falciformis, E. maxima} and \textit{E. tenella} ("Eimeria"), and v) one conserved group containing a broad range of species with, e.g., \textit{Saccharomyces cerevisiae} and \textit{Arabidopsis thaliana} ("conserved"). 
Sporozoites are enriched for \textit{E. falciformis} specific genes, and the group "conserved" is underrepresented, indicating a strong species specificity in the sporozoite stage. Gene clusters characteristic for early infection are enriched for "Api" (cluster 6), probably reflecting that similar processes and genes are involved in asexual reproduction of the selected apicomplexan species. In late infection (cluster 2), characteristic genes are underrepresented for conserved and "Api minus C.h", whereas the other characteristic cluster for late infection is enriched for "conserved" orthologs. This clustering and difference in enrichment of conserved versus apicomplexan (minus \textit{C. hominis}) genes probably reflects activity of generic and phylum specific processes/genes in oocysts.
One oocyst cluster (cluster 1) is underrepresented for "Eimeria", whereas the other oocyst genes (cluster 5) are enriched for "Api". 
\ldots

In addition to analysing conservation by ortholog enrichment, we also performed Spearman correlation analysis between our RNA-seq transcriptomes and RNA-seq data from related parasites. Two datasets for the economically important chicken parasite \textit{E. tenella} (walker15 and reid14) and one dataset of the model apicompomplexan parasite \textit{Toxoplasma gondii} (hehl15) were included in the comparison. Interestingly, this analysis confirms the species specificity for the sporozoite transcriptome, by clustering \textit{E. tenella} sporozoite samples together, but \textit{E. falciformis} sporozoites with \textit{E. falciformis} early infection samples. \textit{E. falciformis} late infection samples correlate the most with \textit{E. tenella} gametocytes, indicating similarity also between the species in this stage and supporting the presence of gametocytes in our samples. \textit{E. tenella} merozoites from both independent studies are most similar to early \textit{E. falciformis} samples, indicating similarity also during asexual reproduction which is shown by the conservation analysis as well (Figure?/Table?). \textit{E. falciformis} oocysts cluster with unsporulated \textit{E. tenella} oocysts, whereas \textit{E. tenella} sporulated oocysts are most similar to \textit{E. tenella} sporozoites. 
%% SUM UP

In summary, we have performed a dual RNA-seq transcriptome of an apicomplexan parasite in its natural host. Our analysis of differentially abundant mRNAs at different time-points post infection highlights large groups of genes which characterize the different life stages of the parasite. The dual approach also allows insight into the host responses to this intracellular infection. 
Using GO enrichment tests and by analysing how species specific these life stage defining gene groups are, we show that the transcriptional profile of sporozoites is the most \textit{E. falciformis} specific stage of the time-points we have analyzed. The sporozoite defining gene group is characterized by ATP production, regulation of protein localization and biosynthetic processes. The analysis further demonstrates that early infection, 3 and 5 dpi, is not homogenous between samples, and that the conformity seen in sporozoites (and oocysts) is regained late in infection. We speculate that parasites during asexual replication are more asynchronous than late stage merozoites or gametocytes on 7 dpi, possibly due to an increasing influence from the host adaptive immune response. The asynchrone early parasite asexual replication stages are enriched for growth related processes, and a large expansion in parasite numbers is reflected in the percentage of parasite genes, read numbers as well asin RT-qPCR data. Gametocyte formation on 7 dpi is supported by GO enrichments of terms for motility and ATP production, as well as 7 dpi samples correlating strongly with \textit{E. tenella} gametocyte samples.  
Our analysis interestingly does not detect any influence of mouse immune status on the parasite transcriptome, even though a difference in oocyst output between Rag1-/- and its C57BL/6 control is clear. We sequenced enriched mRNA and it is possible that non-coding RNAs or other post transcriptional regulatory mechanisms in the parasite could explain differences in oocyst output. Another possibility is that immune competent mice mainly interfere with the parasite at a very late time-point before oocyst formation, which a different experimental design could elucidate.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%	METHODS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Methods}
\subsection{Mice and infection procedure}
Three strains of mice were used in our experiments: NMRI (Charles
River Laboratories, Sulzfeld, Germany), C57BL/6 (), and Rag1-/- on
C57BL/6 background (gift from Susanne Hartmann, FU?).  Animal
procedures were performed according to the German Animal Protection
Laws as directed and approved by the overseeing authority Landesamt
fuer Gesundheit und Soziales (Berlin, Germany). Animals where infected
as described by Schmid et al., (schmid12), but tapwater was used
instead of PBS for administration of oocysts. Briefly, NMRI mice were
infected two times, which will be referred to as first and second
infection. For the first infection, 150 sporulated oocysts were
administered in 100 $\mu$L by oral gavage. During the first infection of
60 mice, all animals were weighed every day. On day zero, before
infection, as well as on 3 dpi, 5 dpi and 7 dpi, caeca from 3-4 sacrificed mice per time point were
collected. Epithelial cells were isolated as described in Schmid et
al.(schmid12). For challenge infection, mice recovered for four weeks
before second infection.  Recovery was monitored by weighing and
visual inspection of fur. For the second infection, 1500 sporulated
oocysts were applied by oral gavage. Three mice were used as
non-second infection control, referred to as day 0, second infection.

\subsection{Oocyst purification for infection and sequencing}
Sporulated oocysts were purified by flotation from feces stored in
potassium dichromate and administered orally in 100 $\mu$L tapwater. One
\textit{E. falciformis} isolate, \textit{E. falciformis} Bayer
Haberkorn 1970, was used for all infections and parasite samples. The
strain is maintained through passage in NMRI mice in our facilities as
described elsewhere (schmid12).

\subsection{Sporozoite isolation}
Sporozoites were isolated from sporocysts by excystment. For this,
sporocysts were incubated at 37\newcommand{\degree}C in DMEM containing 0.04\%
tauroglycocholate (MP Biomedicals) and 0.25\% trypsin (Applichem) for
30 min. Sporozoites were purified by the method of Schmatz et al
(schmatz--).

\subsection{RNA extraction}
Total RNA was isolated from infected epithelial cells, sporozoites and
sporulated oocysts using Trizol according to the manufacturer’s
protocol (Invitrogen). High quality \emph{what is the meaning of 'high
  quality' here?}  RNA was used to produce an mRNA library using the
Illumina’s TruSeq RNA Sample Preparation guide. Sporozoites were stored in 1 mL Trizol until
RNA-isolation. Total RNA was isolated using the PureLink RNA Mini Kit
(Invitrogen) and reverse transcribed into cDNA.

\subsection{Sequencing, sequence quality assessment and alignment}
cDNA samples were sequenced by either GAIIX or Illumina Hiseq 2000 as
specified in SI xx (both unstranded). A fastq\_quality\_filter
(FASTQ-toolkit, version 0.0.14, available at
https://github.com/agordon/fastx\_toolkit.git) was applied to Illumina
Hiseq 2000 samples after replacing "N" s by "." annotation.  A
phred score of 10 was applied. We further set q = 60. These settings
require that nine out of ten bases or more are correct in at least
60\% of the bases for each read.

\subsection{Alignment and reference genomes}
We used the published \textit{Mus musculus} mm10 assembly (Genome
Reference Consortium Mouse Build 38, GCA\_000001635.2) as reference
genome including annotations for mouse data. The
\textit{E. falciformis} genome (Heitlinger14) was downloaded from
ToxoDB (Gajria07). For the alignment, the mouse and parasite genome
files were merged into a dual reference genome, and files including
mRNA sequences from both species were aligned against the dual
reference genome using TopHat2 (version 2.0.14, Trapnell09) with -G
specified, and a Bowtie2 (version 1.1.2, Langmead12) index of the dual
genome. Single-end and pair-end sequence samples were aligned
separately with library type 'fr-unstranded' specified for pair-end
samples. Import into R was enabled by the R package Ballgown, which
requires bam files to be processed by Tablemaker (Frazee15), in our
case used with -qW -G specified. Tablemaker in turn makes use of
Cufflinks (version 2.1.1, Trapnell10).

\subsection{Differential mRNA abundance, data normalisation and sample exclusions}
Count data was normalized using the R-package edgeR (version 3.14.0;
cite) with the upperquartile normalisation method. Briefly, genes with
zero coverage in all samples (libraries) are removed and normalisation
factors are calculated for the 75\% quantile for each library. This
normalisation is suitable for read densities following a negative
binomial distribution. Two samples contradicted this assumption
(parasite data) for later modelling and both mouse and parasite data
from these samples were excluded from further analysis:
NMRI\_1st\_3dpi\_rep1 and NMRI\_2nd\_5dpi\_rep1 (SI ...). The method
then fits a generalized linear model (GLM with a negative binomial
link function) for each gene (glmFit) and then performs likelihood
ratio tests for models w or w/o focal factor (glmLRT).

\subsection{Selection of differentially abundant mRNAs and hierarchical clustering}
A selection of differently abundant mRNAs are used for hierarchical
clustering of \textit{E. falciformis} life cycle relevant genes. In
each comparison (see Table 3), the union of the at most 500 genes
differentially abundant with lowest FDR ($<$0.05) are selected. In the
next step, the 500 mRNAs from each comparison (or less) are
joined. In heatmaps, all samples, i.e., also samples which did not have any
significantly different mRNAs according to our selection, were
included in hierarchical clustering. Scale bar in heatmaps shows 0 as
mean mRNA abundance for each gene (row). Up- (green) and
down-regulation (brown) denote number of standard deviations from 0,
i.e., row mean. Hierarchical clustering was performed using with
Euclidean distances, by the complete linkage method ('complete', R package
base).

All analyses were performed in R (cite R-core). Complete scripts are
available at https://github.com/derele/Ef\_RNAseq.git tagged as version
1.0.
%% We will make this available as a github webpage!!!

\setlength{\tabcolsep}{10pt}
\begin{table}[H]
\small
\begin{center}
\caption{Genes used for hierarchical clustering of mRNAs differently abundant depending on time p.i..}
\begin{tabular}{*3l}    \toprule
	\textit{Data description} & \emph{E. falciformis} genes & Mouse genes	\\ \midrule
	Sum of 1st infection NMRI sample differences	& 4 & 8052 \\ 
	(including oocysts and sporozoites if appl.)	\\	
	Used in hierarchical clustering (heatmap)  	& 1618 & 1313 \\ 	\bottomrule	
\hline
\end{tabular}
\end{center}
%\end{adjustwidth}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Backmatter begins here                   %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{backmatter}

\section*{Competing interests}
  The authors declare that they have no competing interests.

\section*{Author's contributions}
Animal and parasite experiments: Simone Spork, experimental design: Richard Lucius, Simone Spork, Emanuel Heitlinger, RNA sequecing: Christoph Dieterich, data anylsis: Emanuel Heitlinger, Totta Kasemo, writing: Totta Kasemo, Emanuel Heilinger, Simone Spork (all?).

\section*{Acknowledgements}
This project was funded by..... Totta Kasemo is funded by the German Research Foundation (DFG) program GRK4026: Parasite Infections: From Experimental Models to Natural Systems. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                  The Bibliography                       %%
%%                                                         %%
%%  Bmc_mathpys.bst  will be used to                       %%
%%  create a .BBL file for submission.                     %%
%%  After submission of the .TEX file,                     %%
%%  you will be prompted to submit your .BBL file.         %%
%%                                                         %%
%%                                                         %%
%%  Note that the displayed Bibliography will not          %%
%%  necessarily be rendered by Latex exactly as specified  %%
%%  in the online Instructions for Authors.                %%
%%                                                         %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% if your bibliography is in bibtex format, use those commands:
\bibliographystyle{spbasic} % Style BST file (bmc-mathphys, vancouver, spbasic).
\bibliographyi{ef-rnaseq-refs.bib} %{bmc_article}      % Bibliography file (usually '*.bib' )
% for author-year bibliography (bmc-mathphys or spbasic)
% a) write to bib file (bmc-mathphys only)
% @settings{label, options="nameyear"}
% b) uncomment next line
%\nocite{label}

% or include bibliography directly:
% \begin{thebibliography}
% \bibitem{b1}
% \end{thebibliography}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                               %%
%% Figures                       %%
%%                               %%
%% NB: this is for captions and  %%
%% Titles. All graphics must be  %%
%% submitted separately and NOT  %%
%% included in the Tex document  %%
%%                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%
%% Do not use \listoffigures as most will included as separate files

\section*{Figures}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                               %%
%% Tables                        %%
%%                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% Use of \listoftables is discouraged.
%%

\newcommand{\bcell}[2][c]{%
  \begin{tabular}[#1]{@{}c@{}}#2\end{tabular}}

\definecolor{LightCyan}{rgb}{0.88,1,1}
\definecolor{LightRed}{rgb}{1,0.88,1}

% Fri Aug 12 11:05:43 2016
\begin{table}[ht]
\centering
\hspace*{-2.5cm}\begin{tabular}{lllllllll}
  \hline
Sample* & \bcell{Sequencing\\method} & batch & \bcell{total\\reads} & \bcell{reads\\mapping\\Mouse} & \bcell{reads\\mapping\\\textit{E. falciformis}} & \bcell{Percentage\\\textit{E. falciformis}}** & \bcell{detected\\ \textit{E. falciformis}\\genes} \\ 
  \hline
NMRI\_2ndInf\_0dpi\_rep1 & GAII & 2 & 108,937,797 & 70,489,674 & 247 & 0.0004 & 1 \\ 
  Rag\_1stInf\_0dpi\_rep1 & hiseq & 3 & 25,362,793 & 18,853,850 & 443 & 0.0023 & 2 \\ 
  C57BL6\_1stInf\_0dpi\_rep1 & hiseq & 3 & 35,731,249 & 25,119,348 & 457 & 0.0018 & 2 \\ 
  C57BL6\_1stInf\_0dpi\_rep2 & hiseq & 3 & 47,085,959 & 34,377,133 & 608 & 0.0018 & 2 \\ 
  Rag\_1stInf\_0dpi\_rep2 & hiseq & 3 & 46,556,156 & 35,233,327 & 676 & 0.0019 & 2 \\ 
  NMRI\_2ndInf\_0dpi\_rep2 & hiseq & 3 & 58,122,244 & 40,794,245 & 3,406 & 0.0083 & 51 \\ 
  \rowcolor{LightCyan}
  NMRI\_2ndInf\_3dpi\_rep1 & hiseq & 3 & 57,934,016 & 40,544,287 & 4,803 & 0.0118 & 95 \\ 
  \rowcolor{LightCyan}
  NMRI\_2ndInf\_5dpi\_rep2 & hiseq &  63,965,539 & 48,289,181 & 10,941 & 0.0227 & 407 \\ 
  \rowcolor{LightRed}
  NMRI\_1stInf\_0dpi\_rep1 & GAII & 1 & 82,364,585 & 55,176,243 & 17,954 & 0.0325 & 701 \\ 
  NMRI\_2ndInf\_3dpi\_rep2 & hiseq & 3 & 65,548,826 & 46,171,909 & 29,548 & 0.0640 & 1,580 \\ 
  NMRI\_2ndInf\_7dpi\_rep2 & hiseq & 3 & 67,487,466 & 51,722,265 & 40,091 & 0.0775 & 1,836 \\ 
  Rag\_1stInf\_5dpi\_rep1 & hiseq & 3 & 38,651,359 & 29,982,453 & 63,024 & 0.2098 & 2,548 \\ 
  Rag\_1stInf\_5dpi\_rep2 & hiseq & 3 & 34,779,832 & 25,297,803 & 99,000 & 0.3898 & 2,828 \\ 
  C57BL6\_1stInf\_5dpi\_rep1 & hiseq & 3 & 40,904,388 & 29,319,604 & 185,969 & 0.6303 & 4,173 \\ 
  Rag\_2ndInf\_5dpi\_rep1 & hiseq & 3 & 50,049,848 & 37,093,621 & 192,856 & 0.5172 & 4,167 \\ 
  C57BL6\_1stInf\_5dpi\_rep2 & hiseq & 3 & 29,511,368 & 18,062,349 & 215,696 & 1.1801 & 3,823 \\ 
  C57BL6\_2ndInf\_5dpi\_rep1 & hiseq & 3 & 35,148,432 & 25,660,184 & 262,909 & 1.0142 & 4,563 \\ 
  NMRI\_1stInf\_3dpi\_rep1 & GAII & 1 & 73,236,430 & 49,993,358 & 394,384 & 0.7827 & 5,220 \\ 
  NMRI\_1stInf\_3dpi\_rep2 & GAII & 2 & 160,709,694 & 117,791,044 & 413,051 & 0.3494 & 4,862 \\ 
  NMRI\_1stInf\_5dpi\_rep2 & GAII & 2 & 119,902,722 & 76,419,774 & 794,570 & 1.0290 & 5,333 \\ 
  NMRI\_2ndInf\_5dpi\_rep1 & GAII & 2 & 230,773,955 & 143,186,486 & 1,846,840 & 1.2734 & 5,533 \\ 
  NMRI\_2ndInf\_7dpi\_rep1 & hiseq & 3 & 70,366,762 & 41,467,146 & 8,634,201 & 17.2335 & 5,875 \\ 
  NMRI\_1stInf\_5dpi\_rep1 & GAII & 2 & 76,702,168 & 47,037,087 & 8,669,701 & 15.5631 & 5,700 \\ 
  NMRI\_sporozoites\_rep2 & GAII & 0 & 19,551,681 & 8,656 & 11,470,604 & 99.9246 & 5,513 \\ 
  NMRI\_1stInf\_5dpi\_rep3 & GAII & 0 & 191,099,180 & 83,735,624 & 27,839,458 & 24.9513 & 5,784 \\ 
  NMRI\_1stInf\_7dpi\_rep1 & GAII & 1 & 66,505,514 & 3,310,666 & 39,400,884 & 92.2488 & 5,932 \\ 
  NMRI\_sporozoites\_rep1 & GAII & 1 & 67,325,397 & 4,334 & 43,774,401 & 99.9901 & 5,825 \\ 
  NMRI\_oocysts\_rep1 & GAII & 1 & 68,859,802 & 3,805 & 49,653,065 & 99.9923 & 5,695 \\ 
  NMRI\_oocysts\_rep2 & GAII & 0 & 151,090,783 & 18,524 & 71,019,860 & 99.9739 & 5,777 \\ 
  NMRI\_1stInf\_7dpi\_rep2 & GAII & 1 & 139,749,046 & 21,699,324 & 73,539,445 & 77.2159 & 5,943 \\ 
   \hline
\end{tabular}
\begin{tablenotes}[flushleft]\footnotesize \singlespacing
\item{*} sample names are given as a$)$ mouse strain b$)$ first or challenge
  infection c$)$ days post infection (dpi) and d$)$ replicate number
  seperated by undersocre . \\
\item{**} percentag mapping \textit{E. falciformis} is given as percentage in total mapping reads
\end{tablenotes}
\end{table}
\hspace*{+2.5cm}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% 	NUMBER OF GENES DIFFERENT IN DIFFERENT COMPARISONS	
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% I like this table. Looks good and transprorts the right kind of
%% information

\clearpage
\section{mRNA abundance differences between different experimental groups}
\setlength{\tabcolsep}{8pt}
\begin{table}[H]
\small
\begin{center}
\caption{mRNA abundance differences between different experimental groups.}
%\label{tab:table3}
\begin{tabular}{*3l}    \toprule
\textit{Day post infection} & \textit{Ef} genes different & Mouse genes different \\ 
	\textit{comparisons} 	    & (FDR$\leq$1\%) &  (FDR$\leq$1\%/5\%) \\ \midrule
	NMRI 0 vs NMRI 3		& NA   & 274 \\
	NMRI 0 vs NMRI 5		& NA   & 1736 \\
	NMRI 0 vs NMRI 7		& NA   & 2802 \\
	NMRI 3 vs NMRI 5     		& 111  & 1 \\
	NMRI 3 vs NMRI 7  		& 1385 & 1407 \\ 
	NMRI 5 vs NMRI 7  		& 1895 & 873 \\ 
	C57BL/6 0 vs C57BL/6 5		& NA	& 914 \\
	Rag1-/vs Rag1-/- 5		& NA	& 45 \\ \midrule
\textit{Day post infection,} & 		 & 		 \\ 
\textit{parasite relevant comparisons} 	    & 		&  \\ \midrule
	Oocysts vs NMRI 3  	& 3310 & NA \\  
	Oocysts vs NMRI 5	& 3605 & NA \\ 
	Oocysts vs NMRI 7	& 3085 & NA \\ 
	Oocysts vs sporozoites  & 3421 & NA \\
	Sporozoites vs NMRI 3 	& 1663 & NA \\
	Sporozoites vs NMRI 5 	& 1605 & NA \\
	Sporozoites vs NMRI 7 	& 2473 & NA \\ \midrule
\textit{First and second infection} & 		 & 	 \\ 
\textit{comparisons} 	    & 		& 	\\ \midrule
	NMRI 3 1st vs NMRI 3 2nd  	& 0  & 5 \\
	NMRI 5 1st vs NMRI 5 2nd  	& 0  & 1 \\
	NMRI 7 1st vs NMRI 7 2nd  	& 0  & 902 \\
	C57BL/6 1st vs C57BL/6 2nd (day 5) & 0 &  mouse \\
	Rag1-/- 1st vs Rag1-/- 2nd (day 5) & 0 & mouse  \\ \midrule
\bottomrule
	\hline
\end{tabular}
\end{center}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                               %%
%% Additional Files              %%
%%                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section*{Additional Files}
  \subsection*{Additional file 1 --- Raw and normalized counts }
  Raw counts of reads mappins to the \textit{E. falciformis} and mouse
  genome for individual samples in our study. Normalized counts for
  seperately for the host and parasite mappings (three compressed csv
  files).

  \subsection*{Additional file 2 --- Results of statistical tests (edgeR) }
  Focal contrast, fold-changes, likelihood ratio in/excluding this
  difference in models, p-values , and false discovery rates (adjusted
  p-values) are given for all tested contrasts (one compressed csv
  file).

  \subsection*{Additional file 3 --- Additional methods and resuts }
  % this would be cool but requires special referencing in the text
  % (hard to keep track and hard to fit in the journals referencing
  % system). But we can try.
  Document containing additonal figures and summary tables (pdf).

\subsection*{Additional file 4 --- Results of enrichment analyses (topGO)}
  Tables listing all tested gene sets and resulting significatn GO
  terms.

\end{backmatter}
\end{document}